### 2020-10-28 Manuscript/Manuscript_Paper.Rmd
Create Manuscript_Paper.Rmd

---
title: "Untitled"
author: "Kimberly Hinz"
date: "2020/12/08"
output:
  word_document:
    reference_docx: Styles1.docx
    fig_caption: true
    # toc: true
    # toc_depth: 5
# bibliography: References.bib
csl: nature-genetics.csl
---

```{=openxml}
<w:p><w:r><w:br w:type="page"/></w:r></w:p>
```
# Abstract
*Mixta* is a recently described genus within the bacterial family *Erwiniaceae* that is closely related to *Pantoea*, *Erwinia*, and *Tatumella*. Depending on the genes and method used for phylogenetic analysis, the genus with which *Mixta* shares the most recent ancestor differs. This study aimed to determine the cause behind these contentious results, and identified genes that resulted from horizontal gene transfer. Ten complete or partial genomes – two *Mixta* species, six species from other *Erwiniaceae* genera, and two outgroups – were retrieved from NCBI and annotated. Homologous genes were extracted yielding a dataset of 1035 genes. The genes were aligned With the ClustalW algorithm and MEGA-CC was used for model testing and computing the distance matrix for each gene. *Pantoea* was the closest relative to the *Mixta* species in most of the analyses; however, results varied for some genes and gene datasets as they have likely resulted from horizontal gene transfer. Diligence must be given to the selection of genes for phylogenetic analysis in order to prevent xenologous signal from distorting the true relationships.

# Introduction
The bacterial family Erwiniaceae includes diverse Gram-negative 1 2 organisms including the genera Mixta, Pantoea, Erwinia, Tatumella, Buchnera, Wigglesworthia, Phaseolibacter, Kalamiella (LPSN, http://www.bacterio.net/). Three of these, Mixta, Pantoea, and Erwinia, have been the focus of numerous studies given their relevance to plant and human pathogens. For example, the genus Erwinia contains plant pathogens (e.g. E. amylovora) 3 4 5, epiphytes (e.g. E. tasmaniensis) 6 7, species that may contribute to plant disease 8, and others that are insect-associated 9 10 11. Similarly, Pantoea, a genus reported by Gavini et al. 12, includes species that inhabit a variety of environments 12 13 14, including human and plant pathogens 2 15 16 17 18 19 20 as well as species associated with plant growth 21 22, insects 23, and fungi 24 25. Mixta, originally Pantoea, is a novel genus first suggested by Palmer et al. 26 and then later proposed by the same researchers in 2018 27. Two species, M. calida and M. gaviniae (previously species of Pantoea), were originally found in infant formula and its production environment 13 and Fritz et al. 28 demonstrated that some strains of M. calida were linked with post-surgical meningitis in humans. Currently, there are only five Mixta species validly published (LPSN, http://www.bacterio.net/).

Understanding the phylogeny of bacterial species and genes allows researchers to investigate the evolutionary history of the species, their relatedness, and the individual genes that allow the species to survive under various conditions 26 29 30 31 32 33. The most common methods in phylogenetic analysis are those including 16S ribosomal RNA (rRNA) sequences 34, multi-locus sequence analysis (MLSA) with several housekeeping genes 34 35 36, and ribosomal MLSA with 53 structural ribosomal proteins 26 29 37. However, the phylogenetic trees generated from these methods may not be very robust 34 36 38 because of the lower phylogenetic signal from the highly conserved housekeeping genes 34 35 39 40 41 42. This is particularly seen within Erwiniaceae wherein most of our understanding of the phylogeny and interrelationships are based on the 16S rRNA genes 43 44 45 46 47 because of its ubiquity. However, it has insufficient resolution at the species level 36 and the genera Pantoea, Erwinia, Tatumella, and Mixta are often not monophyletic 48. Another issue that may arise with MLSA is unknowingly including paralogues (i.e. genes originating from an intragenomic duplication event) or xenologues (i.e. genes originating from horizontal gene transfer) 26 *49.

The DNA topoisomerase (gyrase) subunit (gyrB), DNA-directed RNA polymerase subunit (rpoB), ATP synthase subunit (atpD), and initiation translation factor (infB) have been shown to be more reliable than 16S rRNA gene sequences for determining intra- and intergeneric relationships 50 51 52 53 and a robust method for the Erwiniaceae family 54. The genera within Erwiniaceae are monophyletic when concatenated partial gyrB, rpoB, atpD, and infB gene sequences are used 38 55 56.

Nevertheless, despite efforts to find a suitable method for identifying the species tree for Erwiniaceae genera, Mixta’s position with relation to Pantoea, Erwinia, and Tatumella vastly differs between studies 2 26 27 47 57 48 58 59. Not only do different gene combinations provide different results, but different algorithms and models for alignments and phylogenetic trees also contribute the contentious results 45 60. For example, Prakash et al. 2, Palmer et al. 26, Rezzonico et al. 48, and Gueule et al. 59 all used gyrB, rpoB, atpD, and infB nucleotide gene sequences in their phylogenetic analyses. Prakash et al. conducted a neighbour-joining statistical method while the others used a maximum-likelihood (ML) method. Three of the studies – Rezzonico et al., Prakash et al., and Gueule et al. – placed the Mixta species within the Pantoea genera whereas the analysis in Palmer et al. places Mixta outside of the other three genera. Palmer et al. 27, Rezzonico et al. 48, and Brady et al. 58 also ran analyses with the four MLSA genes as amino acid sequences. These three studies show contentious results as Rezzonico et al. shows that Mixta is related closest with Tatumella, Palmer et al. shows that Mixta is related closest with Pantoea, and Brady et al. shows Mixta within the Pantoea genus. In a separate study conducted by Palmer et al. 26, two ML trees were computed, one with 1039 core amino acid sequence and the other with 52 ribosomal MLSA nucleotide genes. Both ML trees contained species from Mixta, Pantoea, Erwinia, and Tatumella; however, the analysis with the core genes placed Pantoea and Tatumella as sister taxa followed by two Mixta species whereas the analysis with ribosomal MLSA genes placed Pantoea and Tatumella as sister taxa followed by Erwinia and then by the Mixta species. A possible cause behind these contentious results may be horizontal gene transfer (HGT) events within the Erwiniaceae or from unrelated species. HGT events occur frequently within bacteria and often the genes provide the bacteria the ability to adapt to new environments and/or survive selective pressures.

The objective of my honours thesis was to determine the cause behind the contentious phylogenetic results for Mixta species. To accomplish this, ten complete or partial genomes – two each for Mixta, Pantoea, Erwinia, Tatumella, and outgroup – were retrieved from public databases, annotated, and homologous genes were extracted. Genes were used to make individual phylogenetic trees, and the trees were examined for inconsistent evolutionary histories.

# Materials and Methods
## Genomes and Gene Sets
Publicly available full genomes were retrieved from NCBI comprising of ten type strains from Mixta, Pantoae, Erwinia, Tatumella, Citrobacter, and Enterobacter (Table 1). The genomes were annotated with Prokka v1.14.1 61 using default parameters and homologous genes were extracted using GET_HOMOLOGUES software package 62 with the bidirectional best-hit search algorithm using default parameters. GET_HOMOLOGUES extracted 1284 genes that were common between the ten species.
Within R 63, a filtering function was created to pass forward fasta files with ten, non-truncated sequences. The minimum gene length to pass through the filter was 90% of the length of the longest sequence in that file. This filter ensured that files containing split, missing, or paralogous genes were not included in the analysis. As a result, 1035 files of the 1284 passed this filtering stage.

## Gene Alignment
Gene alignments were done in R with a few software packages. The fasta files were read in using the readDNAStringSet() function 64 (package: Biostrings). The alignments were made using the ClustalW algorithm with the msaClustalW() 65 (package: msa). The number of maximum iterations was set to 100 and the other parameters were kept at default settings 65. Alignments were then converted into a readable format using msaConvert 65 (package: msa) and then written as fasta files using write.fasta() 66 (package: seqinr).

## Distance Matrices
Prior to computing distance matrices, model testing was done for each of the 1035 gene files through MEGA-CC 67 by calling a text file containing the file pathways and a MAO file detailing the analysis preferences. These results were exported as comma-separated values (CSV) and read into R. The best model was chosen based on the lowest Bayesian Information Criterion (BIC) and if it was available for distance matrices in MEGAX 68. General Time Reversible (GTR) models and Hasegawa-Kishino-Yano (HKY) models were not available for distance matrices and so these models were not considered during model testing. Furthermore, the rates among sites option invariant sites (+I) was also not available for distance matrices and were subsequently ignored.

The gene files were separated into groups according to the suggested model and text files were created with their file pathways. These text files were called in turn along with a corresponding MAO file through MEGA-CC to compute the distance matrices for each gene. Variance was estimated with 500 bootstraps, transitions and transversions were allowed, and gaps and missing data were treated with pairwise deletion.

The genetic distances and standard errors relative to M. calida and M. gaviniae were extracted to create plots in R, categorize the closest relatives, run generalized additive models to determine potential non-linear relationships between parameters and genetic distances, and identify genes for further analysis using phylogenetic trees. The ML phylogenetic trees were created using MEGAX and with a model suggested through model testing.

# Results and Discussion

# Conclusion

```{=openxml}
<w:p><w:r><w:br w:type="page"/></w:r></w:p>
```
# Literature Cited

### 2020-10-28 Results_Analysis.R
Update Results_Analysis.R

@@ -8,145 +8,266 @@ library("tidyverse")
 
 setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
 
+# Table 1. ---------------------------------------------------------------------------------------------------------------------------------------
+# ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
+                                 "Erwinia tasmaniensis", "Tatumella ptyseos", "Tatumella saanichensis", 
+                                 "Enterobacter cloacae subsp cloacae", "Pseudomonas syringiae pv syringae"),
+                     Strain = c("DSM_22759", "DSM_22758", "NBRC_102470", "LMG_5345", "CFBP_1232", "ET1/99", "NCTC_11468",
+                                "NML_06-3099", "ATCC_13047", "ICMP_3023"),
+                     `GenBank Accession No.` = c("GCA_002953215.1", "GCA_002953195.1", "GCA_001598475.1", "GCA_002095575.1", "GCA_000367625.2",
+                                                 "GCA_000026185.1", "GCA_900478715.1", "GCA_000439375.1", "GCA_000025565.1", "GCA_001401075.1"),
+                     Level = c("complete genome", "complete genome", "contigs", "contigs", "contigs", "complete genome", "complete genome", 
+                               "contigs", "complete genome", "scaffold"))
+# 
+# library("knitr")
+# kable(strain, caption = "Table 1. Genomes used in this study.")
+# ```
 
-# Project_TEN ------------------------------------------------------------------------------------------------------------------------------------
-NT_four_rel <- read.csv(file = "8_Results_Ten_NT/Four_Relatives_Ten_NT.csv")
-NT_rel_dist <- read.csv(file = "8_Results_Ten_NT/Relatives_Distances_Ten_NT.csv")
-
-AA_four_rel <- read.csv(file = "8_Results_Ten_AA/Four_Relatives_Ten_AA.csv")
-AA_rel_dist <- read.csv(file = "8_Results_Ten_AA/Relatives_Distances_Ten_AA.csv")
-
-## Double Check ==================================================================================================================================
-# This is to double check that the first and second relatives are Mixta species.
-
-# Nucleotides
[…]

### 2020-12-03 5_Distance_Matrices.R
Update 5_Distance_Matrices.R

@@ -60,13 +60,12 @@ first_four_rel <- function(distances, uniq_genes, mixta_spp) {
                              mixta_spp == "M_gaviniae" ~ "M_calida")
     
     four_rel <- mutate(four_rel,
-                       Mixta_check = First %in% c(mixta_spp, other_mixta),
-                       Closest_Relative = case_when(Mixta) ### HERE FOR ROW 433 Have to redo all of this
-                       
-                       
-                       
-                       case_when(Second %in% c(other_mixta, mixta_spp) ~ Third,
-                                 TRUE ~ Second))
+                       Mixta_check = case_when(First == mixta_spp ~ TRUE,
+                                               TRUE ~ FALSE),
+                       Closest_Relative = case_when(First %in% c(mixta_spp, other_mixta) ~
+                                                      case_when(Second %in% c(mixta_spp, other_mixta) ~ Third,
+                                                                TRUE ~ Second),
+                                                    TRUE ~ First))
     
     four <- rbind(four, four_rel)
   }

### 2020-12-03 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -8,7 +8,7 @@ theme_set(theme_bw())
 setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
 
 #
-# Table 1. -----------------------------------------------------------------------------------------------------------------------------------------------
+# Table 1. Genomes --------------------------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
                                  "Erwinia tasmaniensis", "Tatumella ptyseos", "Tatumella saanichensis", 
@@ -24,7 +24,7 @@ strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea aggl
 # kable(strain, caption = "Table 1. Genomes used in this study.")
 # ```
 
-# Table 2. -----------------------------------------------------------------------------------------------------------------------------------------------
+# Table 2. Models ---------------------------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
 phylo_models <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
 PM <- data.frame(table(phylo_models$BM))
@@ -35,10 +35,10 @@ PM <- mutate(PM,
              Percent = round((Number / sum(Number)) * 100, digits = 2))
 # 
 # library("knitr")
-# kable(strain, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC.")
+# kable(PM, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC.")
 # ```
 
-# Table 3 ------------------------------------------------------------------------------------------------------------------------------------------------
+# Table 3. First Relative Nucleotides -------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
 calida_four <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE)
 
[…]

### 2020-12-03 A_Nucleotide_Identity.R
Update A_Nucleotide_Identity.R

@@ -1,8 +1,89 @@
+# This is the first R File for this part of the project: nucleotide identity percentage.
+
+# The following code reads in the fasta files and aligns each species' sequence to a Mixta species for each gene. Then it counts the number of nucleotides
+# that are the same and in the same position (nucleotide identity). The higher the number of identical nucleotides, the more similar that sequence is to
+# Mixta.
+
 library("tidyverse")
 library("seqinr")
-# library("msa")
+library("msa")
 
 setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
+
+# Functions ----------------------------------------------------------------------------------------------------------------------------------------------
+calida_split_align <- function(gene_file_df, NTAA) {
+  for(row in 1:nrow(gene_file_df)) {
+    gene_pair <- rbind(gene_file_df[5, ],
+                      gene_file_df[row, ])
+    
+    test <- DNAStringSet(gene_pair$Sequences)
+    test@ranges@NAMES <- gene_pair$Names
+    
+    
+    
+    
+    
+    
+    
+    test <- align_gene(gene_pair, NTAA)
+  }
+}
+
+
+
+
+
+
+
+
+
+
+align_gene <- function(gene_pair, NTAA) {                                                 # Aligns sequences and converts to a dataframe
+  align <- msa::msaClustalW(inputSeqs = gene_pair, maxiters = 100, 
+                            type = NTAA, order = "input")                                 # Aligns the nucleotide sequences
+  
+  alignConv <- msaConvert(align, type = "seqinr::alignment")                              # Converts the aligned genes into a readable form
+  
+  alignFasta <- data.frame(sequences = alignConv$seq,
+                           species = alignConv$nam)                                       # Covert to format needed for write.fasta()
+}
+
+my_write.fasta <- function(align_file, file_row) {
+  aligned <- align_file
+  
+  write.fasta(sequences = as.list(aligned$sequences), names = aligned$species,
+              file.out = fastaFiles_fltr$Algn_path[file_row],
+              open = "w", nbchar = 10000, as.string = TRUE)
+}
+
+
+# Nucleotides --------------------------------------------------------------------------------------------------------------------------------------------
+fastaFiles <- data.frame(File_name = list.files(path = "4_Filtered_NT/",
+                                                pattern = ".fasta"))                      # List the fasta files in this folder as a dataframe
+
+
[…]

### 2020-12-05 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
     # toc_depth: 5
-# bibliography: References.bib
-csl: nature-genetics.csl
+bibliography: References.bib
+csl: IJSEM.csl
 ---
 
-```{=openxml}
-<w:p><w:r><w:br w:type="page"/></w:r></w:p>
-```
+\newpage
 # Abstract
@@
 Depending on the genes and method used for phylogenetic analysis, the genus with which *Mixta* shares the most recent ancestor differs.
-This study aimed to determine the cause behind these contentious results, and identified genes that resulted from horizontal gene transfer.
+This study aimed to determine the cause behind these contentious results, and identified genes that may have resulted from horizontal gene transfer.
 Ten complete or partial genomes – two *Mixta* species, six species from other *Erwiniaceae* genera, and two outgroups – were retrieved from NCBI and annotated.
-Homologous genes were extracted yielding a dataset of 1035 genes.
-The genes were aligned With the ClustalW algorithm and MEGA-CC was used for model testing and computing the distance matrix for each gene.
-*Pantoea* was the closest relative to the *Mixta* species in most of the analyses; however, results varied for some genes and gene datasets as they have likely resulted from horizontal gene transfer.
-Diligence must be given to the selection of genes for phylogenetic analysis in order to prevent xenologous signal from distorting the true relationships.
+Homologous genes were extracted yielding a dataset of 799 genes.
+The genes were aligned With the ClustalW algorithm and MEGA-CC was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
+Distance matrices were then extracted from these trees.
+Nucleotide and amino acid identity analyses were also done in `R` programming language.
+*Pantoea* was the closest relative to the *Mixta* species in most of the analyses; however, results were not consistent across analyses.
+Some genes were also found to be more similar to other, non-*Pantoea* species.
+Diligence must be given to the selection of genes for phylogenetic analysis and to the method chosen to prevent xenologous signal from distorting the true relationships.
+Furthermore, future research should consider that different phylogenetic analyses may give different results.
 
+\newpage
 # Introduction
-The bacterial family Erwiniaceae includes diverse Gram-negative 1 2 organisms including the genera Mixta, Pantoea, Erwinia, Tatumella, Buchnera, Wigglesworthia, Phaseolibacter, Kalamiella (LPSN, http://www.bacterio.net/).
-Three of these, Mixta, Pantoea, and Erwinia, have been the focus of numerous studies given their relevance to plant and human pathogens.
-For example, the genus Erwinia contains plant pathogens (e.g.
+The bacterial family Erwiniaceae includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, *Kalamiella* (LPSN, http://www.bacterio.net/).
+Three of these, *Mixta*, *Pantoea*, and *Erwinia*, have been the focus of numerous studies given their relevance to plant and human pathogens.
+For example, the genus *Erwinia* contains plant pathogens (e.g.
 E.
-amylovora) 3 4 5, epiphytes (e.g.
+amylovora) [@HaubenSwings2005; @Lopez2011; @Matsuura2012], epiphytes (e.g.
 E.
-tasmaniensis) 6 7, species that may contribute to plant disease 8, and others that are insect-associated 9 10 11.
-Similarly, Pantoea, a genus reported by Gavini et al.
-12, includes species that inhabit a variety of environments 12 13 14, including human and plant pathogens 2 15 16 17 18 19 20 as well as species associated with plant growth 21 22, insects 23, and fungi 24 25.
-Mixta, originally Pantoea, is a novel genus first suggested by Palmer et al.
-26 and then later proposed by the same researchers in 2018 27.
+tasmaniensis) [@Geider2006; @Mergaert1999], species that may contribute to plant disease [@Buonaurio2015], and others that are insect-associated [@Campillo2015; @Harada1997; @Skrodenyte-Arbaciauskiene2012].
+Similarly, *Pantoea*, a genus reported by Gavini et al.[@Gavini1989], includes species that inhabit a variety of environments [@Gavini1989; @Popp2010; @Liu2013], including human and plant pathogens [@Prakash2015; @deBaere2004; @Fullerton2007; @Kratz2003; @Lim2006; @Schmid2003; @vanRostenberghe2006] as well as species associated with plant growth [@Smits2011; @Kim2012], insects [@Palmer2016], and fungi [@Ma2016; @Rong2016].
+*Mixta*, originally *Pantoea*, is a novel genus first suggested by Palmer et al.
+[@Palmer2017] and then later proposed by the same researchers in 2018 [@Palmer2018].
 Two species, M.
 calida and M.
-gaviniae (previously species of Pantoea), were originally found in infant formula and its production environment 13 and Fritz et al.
-28 demonstrated that some strains of M.
+gaviniae (previously species of *Pantoea*), were originally found in infant formula and its production environment [@Popp2010].
+Fritz et al.
+[@Fritz2014] demonstrated that some strains of M.
 calida were linked with post-surgical meningitis in humans.
-Currently, there are only five Mixta species validly published (LPSN, http://www.bacterio.net/).
+Currently, there are only five *Mixta* species validly published (LPSN, http://www.bacterio.net/).
 
-Understanding the phylogeny of bacterial species and genes allows researchers to investigate the evolutionary history of the species, their relatedness, and the individual genes that allow the species to survive under various conditions 26 29 30 31 32 33.
-The most common methods in phylogenetic analysis are those including 16S ribosomal RNA (rRNA) sequences 34, multi-locus sequence analysis (MLSA) with several housekeeping genes 34 35 36, and ribosomal MLSA with 53 structural ribosomal proteins 26 29 37.
-However, the phylogenetic trees generated from these methods may not be very robust 34 36 38 because of the lower phylogenetic signal from the highly conserved housekeeping genes 34 35 39 40 41 42.
-This is particularly seen within Erwiniaceae wherein most of our understanding of the phylogeny and interrelationships are based on the 16S rRNA genes 43 44 45 46 47 because of its ubiquity.
-However, it has insufficient resolution at the species level 36 and the genera Pantoea, Erwinia, Tatumella, and Mixta are often not monophyletic 48.
+Understanding the phylogeny of bacterial species and genes allows researchers to investigate the evolutionary history of the species, their relatedness, and the individual genes that allow the species to survive under various conditions [@Palmer2017; @Bennett2012; @Prasanna2013; @Angus2014; @deMaayer2014; @Fouts2016].
+The most common methods in phylogenetic analysis are those including 16S ribosomal RNA (rRNA) sequences [@Konstantinidis2007], multi-locus sequence analysis (MLSA) with several housekeeping genes [@Konstantinidis2007; @Gevers2005; @Glaeser2015], and ribosomal MLSA with 53 structural ribosomal proteins [@Palmer2017; @Bennett2012; @Jolley2012].
+However, the phylogenetic trees generated from these methods may not be very robust [@Konstantinidis2007; @Glaeser2015; @Brady2008] because of the lower phylogenetic signal from the highly conserved housekeeping genes [@Konstantinidis2007; @Gevers2005; @Fox1992; @Konstantinidis2005; @Staley2006; @RichterRosselloMora2009].
+This is particularly seen within Erwiniaceae wherein most of our understanding of the phylogeny and interrelationships are based on the 16S rRNA genes [@Francino2006; @Hauben1998; @Naum2008; @Sproer1999; @Adeolu2016] because of its ubiquity.
+However, it has insufficient resolution at the species level [@Glaeser2015] and the genera *Pantoea*, *Erwinia*, *Tatumella*, and *Mixta* are often not monophyletic [@Rezzonico2016].
 Another issue that may arise with MLSA is unknowingly including paralogues (i.e.
 genes originating from an intragenomic duplication event) or xenologues (i.e.
-genes originating from horizontal gene transfer) 26 *49.
+genes originating from horizontal gene transfer) [@Palmer2017; @Koonin2005].
 
-The DNA topoisomerase (gyrase) subunit (gyrB), DNA-directed RNA polymerase subunit (rpoB), ATP synthase subunit (atpD), and initiation translation factor (infB) have been shown to be more reliable than 16S rRNA gene sequences for determining intra- and intergeneric relationships 50 51 52 53 and a robust method for the Erwiniaceae family 54.
-The genera within Erwiniaceae are monophyletic when concatenated partial gyrB, rpoB, atpD, and infB gene sequences are used 38 55 56.
+The DNA topoisomerase (gyrase) subunit (gyrB), DNA-directed RNA polymerase subunit (rpoB), ATP synthase subunit (atpD), and initiation translation factor (infB) have been shown to be more reliable than 16S rRNA gene sequences for determining intra- and intergeneric relationships [@Hedegaard1999; @Dauga2002; @Mollet1997; @Paradis2005] and a robust method for the Erwiniaceae family [@Brady2010].
+The genera within Erwiniaceae are monophyletic when concatenated partial gyrB, rpoB, atpD, and infB gene sequences are used [@Brady2008; @Brady2009; @Moretti2011].
 
-Nevertheless, despite efforts to find a suitable method for identifying the species tree for Erwiniaceae genera, Mixta’s position with relation to Pantoea, Erwinia, and Tatumella vastly differs between studies 2 26 27 47 57 48 58 59.
-Not only do different gene combinations provide different results, but different algorithms and models for alignments and phylogenetic trees also contribute the contentious results 45 60.
+Nevertheless, despite efforts to find a suitable method for identifying the species tree for Erwiniaceae genera, *Mixta*’s position with relation to *Pantoea*, *Erwinia*, and *Tatumella* vastly differs between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
+Not only do different gene combinations provide different results, but different algorithms and models for alignments and phylogenetic trees also contribute to the contentious results [@Naum2008; @Octavia2014].
 For example, Prakash et al.
-2, Palmer et al.
-26, Rezzonico et al.
-48, and Gueule et al.
-59 all used gyrB, rpoB, atpD, and infB nucleotide gene sequences in their phylogenetic analyses.
+[@Prakash2015], Palmer et al.
+[@Palmer2017], Rezzonico et al.
+[@Rezzonico2016], and Gueule et al.
+[@Gueule2015] all used gyrB, rpoB, atpD, and infB nucleotide gene sequences in their phylogenetic analyses.
 Prakash et al.
@@
 Three of the studies – Rezzonico et al., Prakash et al., and Gueule et al.
-– placed the Mixta species within the Pantoea genera whereas the analysis in Palmer et al.
-places Mixta outside of the other three genera.
+– placed the *Mixta* species within the *Pantoea* genera whereas the analysis in Palmer et al.
+placed *Mixta* outside of the other three genera.
 Palmer et al.
-27, Rezzonico et al.
-48, and Brady et al.
-58 also ran analyses with the four MLSA genes as amino acid sequences.
+[@Palmer2018], Rezzonico et al.
+[@Rezzonico2016], and Brady et al.
+[@Brady2012] also ran analyses with the four MLSA genes as amino acid sequences.
 These three studies show contentious results as Rezzonico et al.
-shows that Mixta is related closest with Tatumella, Palmer et al.
-shows that Mixta is related closest with Pantoea, and Brady et al.
-shows Mixta within the Pantoea genus.
+found that *Mixta* is related closest with *Tatumella*, Palmer et al.
+found that *Mixta* is related closest with *Pantoea*, and Brady et al.
+found *Mixta* within the *Pantoea* genus.
 In a separate study conducted by Palmer et al.
-26, two ML trees were computed, one with 1039 core amino acid sequence and the other with 52 ribosomal MLSA nucleotide genes.
-Both ML trees contained species from Mixta, Pantoea, Erwinia, and Tatumella; however, the analysis with the core genes placed Pantoea and Tatumella as sister taxa followed by two Mixta species whereas the analysis with ribosomal MLSA genes placed Pantoea and Tatumella as sister taxa followed by Erwinia and then by the Mixta species.
-A possible cause behind these contentious results may be horizontal gene transfer (HGT) events within the Erwiniaceae or from unrelated species.
+@Palmer2017, two ML trees were computed, one with 1039 core amino acid sequence and the other with 52 ribosomal MLSA nucleotide genes.
+Both ML trees contained species from *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*; however, the analysis with the core genes placed *Pantoea* and *Tatumella* as sister taxa followed by two *Mixta* species whereas the analysis with ribosomal MLSA genes placed *Pantoea* and *Tatumella* as sister taxa followed by *Erwinia* and then by the *Mixta* species.
+
+There are two possible causes behind contentious results in phylogenetic analyses.
+One is the use of different methods, including the use of a different phylogenetic tree model from another study despite both studies conducting maximum likelihood trees and the same sequence type.
+Additionally, there is no common phylogenetic analysis model between nucleotide and amino acid sequences; therefore, evolutionary history is estimated differently.
+The other reason may be horizontal gene transfer (HGT) events within the *Erwiniaceae* or from a non-closely-related species.
 HGT events occur frequently within bacteria and often the genes provide the bacteria the ability to adapt to new environments and/or survive selective pressures.
 
-The objective of my honours thesis was to determine the cause behind the contentious phylogenetic results for Mixta species.
-To accomplish this, ten complete or partial genomes – two each for Mixta, Pantoea, Erwinia, Tatumella, and outgroup – were retrieved from public databases, annotated, and homologous genes were extracted.
+The objective of this paper was to determine the cause behind the contentious phylogenetic results for the *Mixta* species.
+To accomplish this, homologous genes were extracted from ten complete or partial genomes - two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*, and two outgroups - that were retrieved from public databases.
+Model testing was done for each gene in the dataset and phylogenetic trees were computed using the recommended model.
+Distance matrices were exported from the phylogenetic trees and these were used to determine which species were most closely related to *Mixta* for each gene.
+Nucleotide and amino acid identity analyses were also conducted for each gene in order to compare
+
+The objective of my honours thesis was to determine the cause behind the contentious phylogenetic results for *Mixta* species.
+To accomplish this, ten complete or partial genomes – two each for *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, and outgroup – were retrieved from public databases, annotated, and homologous genes were extracted.
 Genes were used to make individual phylogenetic trees, and the trees were examined for inconsistent evolutionary histories.
@@
 ## Genomes and Gene Sets
-Publicly available full genomes were retrieved from NCBI comprising of ten type strains from Mixta, Pantoae, Erwinia, Tatumella, Citrobacter, and Enterobacter (Table 1).
+Publicly available full genomes were retrieved from NCBI comprising of ten type strains from *Mixta*, Pantoae, *Erwinia*, *Tatumella*, Citrobacter, and Enterobacter (Table 1).
 The genomes were annotated with Prokka v1.14.1 61 using default parameters and homologous genes were extracted using GET_HOMOLOGUES software package 62 with the bidirectional best-hit search algorithm using default parameters.
@@
 # Literature Cited
+<div id="refs"></div>
+
+\newpage
+# Appendix

### 2020-12-06 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -24,21 +24,42 @@ strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea aggl
 # kable(strain, caption = "Table 1. Genomes used in this study.")
 # ```
 
-# Table 2. Models ---------------------------------------------------------------------------------------------------------------------------------------
+# Table 2. Nucleotide Models -----------------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
-phylo_models <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
-PM <- data.frame(table(phylo_models$BM))
+phylo_models_NT <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
+PM_NT <- data.frame(table(phylo_models_NT$BM))
 
-colnames(PM) <- c("Model", "Number")
+colnames(PM_NT) <- c("Model", "Number")
 
-PM <- mutate(PM,
+PM_NT <- mutate(PM_NT,
              Percent = round((Number / sum(Number)) * 100, digits = 2))
 # 
 # library("knitr")
-# kable(PM, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC.")
+# kable(PM_NT, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for 
+# nucleotide sequences.")
 # ```
 
-# Table 3. First Relative Nucleotides -------------------------------------------------------------------------------------------------------------------
+# Table 3. Amino Acid Models ----------------------------------------------------------------------------------------------------------------------------
+# ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+phylo_models_AA <- read.csv("9_Results_AA/Phylo_Models_AA.csv", stringsAsFactors = FALSE)
+PM_AA <- data.frame(table(phylo_models_AA$BM))
+
+colnames(PM_AA) <- c("Model", "Number")
+
+PM_AA <- mutate(PM_AA,
+                Percent = round((Number / sum(Number)) * 100, digits = 2))
+# 
+# library("knitr")
+# kable(PM_AA, caption = "Table 3. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for
+# amino acid sequences.")
+# ```
+
+
+
+
+
+
[…]

### 2020-12-06 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 To accomplish this, homologous genes were extracted from ten complete or partial genomes - two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*, and two outgroups - that were retrieved from public databases.
+In this study, both the nucleotide and amino acid sequences for the same genes were separately analyzed.
 Model testing was done for each gene in the dataset and phylogenetic trees were computed using the recommended model.
 Distance matrices were exported from the phylogenetic trees and these were used to determine which species were most closely related to *Mixta* for each gene.
-Nucleotide and amino acid identity analyses were also conducted for each gene in order to compare
-
-The objective of my honours thesis was to determine the cause behind the contentious phylogenetic results for *Mixta* species.
-To accomplish this, ten complete or partial genomes – two each for *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, and outgroup – were retrieved from public databases, annotated, and homologous genes were extracted.
-Genes were used to make individual phylogenetic trees, and the trees were examined for inconsistent evolutionary histories.
+Nucleotide and amino acid identity analyses were also conducted for each gene in order to compare the results between evolutionary history and sequence similarity.
+Phylogenetic trees were then done with either one gene or multiple genes to confirm results.
 
@@
 ## Genomes and Gene Sets
-Publicly available full genomes were retrieved from NCBI comprising of ten type strains from *Mixta*, Pantoae, *Erwinia*, *Tatumella*, Citrobacter, and Enterobacter (Table 1).
-The genomes were annotated with Prokka v1.14.1 61 using default parameters and homologous genes were extracted using GET_HOMOLOGUES software package 62 with the bidirectional best-hit search algorithm using default parameters.
-GET_HOMOLOGUES extracted 1284 genes that were common between the ten species.
-Within R 63, a filtering function was created to pass forward fasta files with ten, non-truncated sequences.
-The minimum gene length to pass through the filter was 90% of the length of the longest sequence in that file.
-This filter ensured that files containing split, missing, or paralogous genes were not included in the analysis.
-As a result, 1035 files of the 1284 passed this filtering stage.
+Publicly available full genomes were retrieved from NCBI comprising of ten type strains from *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Enterobacter*, and *Pseudomonas* (Table 1).
+*Enterobacter cloacae* is in the same order of *Enterobacterales* as *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella* whereas *Pseudomonas syringiae* is in the same class of *Gammaproteobacteria*.
+The genomes were annotated with Prokka v1.14.1 [@Seemann2014] using default parameters and homologous genes were extracted using GET_HOMOLOGUES software package [@Contreras-Moreira2013] with the bidirectional best-hit search algorithm using default parameters.
+GET_HOMOLOGUES extracted 954 homologous genes between the ten species.
+
+```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
+                                 "Erwinia tasmaniensis", "Tatumella ptyseos", "Tatumella saanichensis", 
+                                 "Enterobacter cloacae subsp cloacae", "Pseudomonas syringae pv syringae"),
+                     Strain = c("DSM_22759", "DSM_22758", "NBRC_102470", "LMG_5345", "CFBP_1232", "ET1/99", "NCTC_11468",
+                                "NML_06-3099", "ATCC_13047", "ICMP_3023"),
+                     `GenBank Accession No.` = c("GCA_002953215.1", "GCA_002953195.1", "GCA_001598475.1", "GCA_002095575.1", "GCA_000367625.2",
+                                                 "GCA_000026185.1", "GCA_900478715.1", "GCA_000439375.1", "GCA_000025565.1", "GCA_001401075.1"),
+                     Level = c("complete genome", "complete genome", "contigs", "contigs", "contigs", "complete genome", "complete genome", 
+                               "contigs", "complete genome", "scaffold"))
+
+library("knitr")
+kable(strain, caption = "Table 1. Genomes used in this study.")
+```
+
+In `R` [@RCore2019], two filter functions were written to pass through gene files of the nucleotide sequences.
+Since both nucleotide and amino acid sequences were analyzed, the filters were only applied to the nucleotide sequences.
+A list of the passed nucleotide gene files was used to isolate the corresponding amino acid gene files to be used in later analyses.
+
+The first filter passed through gene files that contained ten, single copy sequences.
+A minimum relative length was set where all sequences in a gene file had to be at least 90% the length of the longest sequence in the same file.
+For example, if the longest sequence in a file is 1000 bp, then the rest of the sequences had to be at least 900 bp.
+Of 954 genes, 38 genes were excluded for having at least extra sequence and an additional 175 genes were excluded because at least one sequence did not meet the length requirements.
+
+However, there may be an evolutionary reason why at least one of the sequences did not meet the length requirements.
+The second filter took the 175 gene files that were denied by the first filter one at a time and split the sequences into two groups; the first group contained the gene sequences that were over the 90% length cutoff and the second contained the rest that were shorter than the cutoff length.
+The first requirement for this filter was that all sequences in the file were longer than 80% of the length of the longest genes in the file.
+The second requirement was that both representative of a genus were in the same group, either above or below the 90% cutoff.
+The two outgroup species were not required to be in the same group since they are not from the same genus.
+An additional 58 genes were added to the working dataset, resulting in 799 genes.
+A list of the 799 gene names was used to grab the corresponding amino acid files.
 
 ## Gene Alignment
-Gene alignments were done in R with a few software packages.
-The fasta files were read in using the readDNAStringSet() function 64 (package: Biostrings).
-The alignments were made using the ClustalW algorithm with the msaClustalW() 65 (package: msa).
-The number of maximum iterations was set to 100 and the other parameters were kept at default settings 65.
-Alignments were then converted into a readable format using msaConvert 65 (package: msa) and then written as fasta files using write.fasta() 66 (package: seqinr).
+Gene alignments were in `R` using a ClustalW algorithm.
+The gene fasta files were read into `R` using the `readDNAStringSet()` function (package: Biostrings) [@Pages2020].
+Alignments were done using the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and setting maximum number of iterations to 100.
+The aligned functions were then converted into a readble format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
 
 ## Distance Matrices
-Prior to computing distance matrices, model testing was done for each of the 1035 gene files through MEGA-CC 67 by calling a text file containing the file pathways and a MAO file detailing the analysis preferences.
-These results were exported as comma-separated values (CSV) and read into R.
-The best model was chosen based on the lowest Bayesian Information Criterion (BIC) and if it was available for distance matrices in MEGAX 68.
-General Time Reversible (GTR) models and Hasegawa-Kishino-Yano (HKY) models were not available for distance matrices and so these models were not considered during model testing.
-Furthermore, the rates among sites option invariant sites (+I) was also not available for distance matrices and were subsequently ignored.
+After alignments, model testing was done for each gene through `MEGA-CC` [@Kumar2012] for both nucleotide and amino acid sequences.
+In a command line terminal, `MEGA-CC` is called along with a text file listing the file pathways to all 799 genes and a MAO file detailling the analysis preferences.
+Default paramaters were used for model testing.
+This resulted in a comma-separated values (CSV) file for each gene.
+These CSV files were read into `R` and the model for phylogenetic analysis was extracted based on lowest Bayesian Information Criterion (BIC).
+A list of the unique models and the number of genes that required each model are given in Table 2 for nucleotide sequences and in Table 3 for amino acid sequences.
 
-The gene files were separated into groups according to the suggested model and text files were created with their file pathways.
-These text files were called in turn along with a corresponding MAO file through MEGA-CC to compute the distance matrices for each gene.
-Variance was estimated with 500 bootstraps, transitions and transversions were allowed, and gaps and missing data were treated with pairwise deletion.
+```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+library("tidyverse")
 
-The genetic distances and standard errors relative to M.
-calida and M.
-gaviniae were extracted to create plots in R, categorize the closest relatives, run generalized additive models to determine potential non-linear relationships between parameters and genetic distances, and identify genes for further analysis using phylogenetic trees.
-The ML phylogenetic trees were created using MEGAX and with a model suggested through model testing.
+phylo_models_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
+PM_NT <- data.frame(table(phylo_models_NT$BM))
+
+colnames(PM_NT) <- c("Model", "Number")
+
+PM_NT <- mutate(PM_NT,
+             Percent = round((Number / sum(Number)) * 100, digits = 2))
+
+library("knitr")
+kable(PM_NT, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for nucleotide sequences.")
+```
+
+```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+
+phylo_models_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_AA/Phylo_Models_AA.csv", stringsAsFactors = FALSE)
+PM_AA <- data.frame(table(phylo_models_AA$BM))
+
+colnames(PM_AA) <- c("Model", "Number")
+
+PM_AA <- mutate(PM_AA,
+                Percent = round((Number / sum(Number)) * 100, digits = 2))
+
+library("knitr")
+kable(PM_AA, caption = "Table 3. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for amino acid sequences.")
+```
+
+For each unique model, a text file was written listing the gene file pathways of each gene that required that model.
+In a command line terminal, `MEGA-CC` [@Kumar2012] was called with a text file listing gene file pathways for a certain model and the corresponding MAO file.
+This was done for each unique model.
+These MAO files specified the use of a maximum likelihood statistical method, 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
+If a model required a Gamma distribution, the number of discrete Gamma categories was set to 5, which is default.
+
+Distance matrices were extracted from each of the phylogenetic trees and saved as Excel files.
+These files were read into `R` and the genetic distances between both of the *Mixta* species and all of the sequences were extracted for further analysis in `R`.
+Tables were created in `R Markdown` using the `kable()` function (package: knitr) [@Xie2020] and figures were created using the `ggplot()` function (package: ggplot2) [@Wickham2016].
 
@@
 
-```{=openxml}
-<w:p><w:r><w:br w:type="page"/></w:r></w:p>
-```
+\newpage
 # Literature Cited

### 2020-12-08 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -3,10 +3,15 @@
 # The following code analyzes the distance matrices.
 
 library("tidyverse")
-theme_set(theme_bw())
+library("msa")
+library("ggplot2")
+library("cowplot")
+library("RColorBrewer")
+library("extrafont")
+# loadfonts(device = "win")
+theme_set(theme_classic())
 
 setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
-
 #
 # Table 1. Genomes --------------------------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
@@ -24,9 +29,9 @@ strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea aggl
 # kable(strain, caption = "Table 1. Genomes used in this study.")
 # ```
 
-# Table 2. Nucleotide Models -----------------------------------------------------------------------------------------------------------------------------
+# Table 2. Nucleotide Models ----------------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
-phylo_models_NT <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
+phylo_models_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
 PM_NT <- data.frame(table(phylo_models_NT$BM))
 
 colnames(PM_NT) <- c("Model", "Number")
@@ -41,7 +46,7 @@ PM_NT <- mutate(PM_NT,
 
 # Table 3. Amino Acid Models ----------------------------------------------------------------------------------------------------------------------------
 # ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
-phylo_models_AA <- read.csv("9_Results_AA/Phylo_Models_AA.csv", stringsAsFactors = FALSE)
+phylo_models_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_AA/Phylo_Models_AA.csv", stringsAsFactors = FALSE)
 PM_AA <- data.frame(table(phylo_models_AA$BM))
 
 colnames(PM_AA) <- c("Model", "Number")
@@ -53,83 +58,224 @@ PM_AA <- mutate(PM_AA,
[…]

### 2020-12-08 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
-```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+```{r Table1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
@@
 Alignments were done using the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and setting maximum number of iterations to 100.
-The aligned functions were then converted into a readble format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
+The aligned functions were then converted into a readable format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
 
@@
 After alignments, model testing was done for each gene through `MEGA-CC` [@Kumar2012] for both nucleotide and amino acid sequences.
-In a command line terminal, `MEGA-CC` is called along with a text file listing the file pathways to all 799 genes and a MAO file detailling the analysis preferences.
-Default paramaters were used for model testing.
+In a command line terminal, `MEGA-CC` is called along with a text file listing the file pathways to all 799 genes and a MAO file detailing the analysis preferences.
+Default parameters were used for model testing.
 This resulted in a comma-separated values (CSV) file for each gene.
@@
 
-```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+```{r Table2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
 library("tidyverse")
@@
 
-colnames(PM_NT) <- c("Model", "Number")
+colnames(PM_NT) <- c("Model", "#")
 
 PM_NT <- mutate(PM_NT,
-             Percent = round((Number / sum(Number)) * 100, digits = 2))
+                `%` = round((`#` / sum(`#`)) * 100, digits = 2))
+
+PM_NT <- cbind(PM_NT[1:4, ], PM_NT[5:8, ], PM_NT[9:12, ])
 
@@
 
-```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
 
@@
 
-colnames(PM_AA) <- c("Model", "Number")
+colnames(PM_AA) <- c("Model", "#")
 
 PM_AA <- mutate(PM_AA,
-                Percent = round((Number / sum(Number)) * 100, digits = 2))
+                `%` = round((`#` / sum(`#`)) * 100, digits = 2))
+
+PM_AA <- rbind(PM_AA, c("--", "--", "--"))
+
+PM_AA <- cbind(PM_AA[1:8, ], PM_AA[9:16, ], PM_AA[17:24, ])
 
@@
 # Results and Discussion
+After the two filters, 799 of of 954 homologous genes remained.
+The genes included those that encoded rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the *M.
+calida* and *M.
+gaviniae* genomes.
 
+## Closest Relative Based on Genetic Distance
+Genetic distance gives a numeric estimate of evolutionary change between species.
+The species that has the shortest genetic distance to a species of interest is considered to be closest relative.
+For the following analysis, model testing was done first to estimate the best model for each gene.
+Next, individual phylogenetic gene trees were created and it was from these that distance matrices were exported.
+*P.
+septica* was the closest relative to 66.3% and 66.5% of the *M.
+calida* and *M.
+gaviniae* genes, respectively, for the nucleotide sequences (Figure 1).
+In comparison, only 42.7% and 42.8% of *M.
+calida* and *M.
+gaviniae* genes, respectively, are closely related to *P.
+septica* for the amino acid sequences (Figure 1), but this is still the majority.
+These results suggest that *P.
+septica* is the closest relative to *Mixta* followed by *P.
+agglomerans*.
+For the amino acid sequences, instead of the *Mixta* genes being more closely related to *P.
+septica*, they are more closely related to the other species in this study (Figure 1).
+For example, the percentage of genes related to *P.
+agglomerans* increased from 13.9% for both of the *Mixta* species to 22.3 and 22.4% to *M.
+calida* and *M.
+gaviniae*, respectively.
+This may be due to a loss of phylogenetic signal when using amino acid sequences as a result of codon degeneracy and having a third of available sites.
+
+```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
+The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using genetic distances.
+The species with the shortest genetic distance from both of the *Mixta* species is considered to be the closest relative.
+Blue bars are the nucleotide sequences of the genes with light blue bars representing the percentage of *M.
+calida* genes and dark blue bars representing *M.
+gaviniae*.
+Green bars are the corresponding amino acid sequences with light green representing *M.
+calida* genes and dark blue bars representing *M.
+gaviniae*.
+Approximately 66% of *M.
+calida* and *M.
+gaviniae* genes were most closely related to *P.
+septica* when using nucleotide sequences in comparison to only 43% when using amino acid sequences.", fig.width=6.5}
+library("tidyverse")
+library("ggplot2")
+library("RColorBrewer")
+library("extrafont")
+theme_set(theme_classic())
+
+count_relatives <- function(data) {
+  num_rel <- c(sum(data$Closest_Relative == "P_septica"), sum(data$Closest_Relative == "P_agglomerans"), 
+               sum(data$Closest_Relative == "E_tasmaniensis"), sum(data$Closest_Relative == "E_amylovora"),
+               sum(data$Closest_Relative == "T_ptyseos"), sum(data$Closest_Relative == "T_saanichensis"),
+               sum(data$Closest_Relative == "E_cloacae"), sum(data$Closest_Relative == "P_syringae"))
+}
+
+datasets <- data.frame(project = c("NT", "NT", "AA", "AA"),
+                       dataset = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"))
+
+number_relatives <- data.frame(matrix(ncol = 4, nrow = 0))
+
+for(row in 1:nrow(datasets)) {
+  four_rel <- read.csv(paste("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_", 
+                             datasets$project[row], "/four_relatives_",
+                             datasets$dataset[row], ".csv", sep = ""), 
+                       stringsAsFactors = FALSE)
+
+  num_rel <- data.frame(Species = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                    "E. cloacae", "P. syringae"),
+                        Number = count_relatives(four_rel))
+
+  num_rel <- mutate(num_rel,
+                    Species = factor(Species, levels = Species, ordered = TRUE),
+                    Percent = round((Number / sum(Number)) * 100, digits = 1),
+                    Type = datasets$dataset[row])
+
+  number_relatives <- rbind(number_relatives, num_rel)
+}; rm(four_rel, num_rel, row)
+
+number_relatives <- mutate(number_relatives,
+                           Type = factor(Type, levels = datasets$dataset, ordered = TRUE))
+
+# The plot:
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+```
+
+Unexpectedly, the number of *Mixta* genes that are more closely related increased from 2.9% to 6.0% when using the amino acid sequences instead of the nucleotide sequences (Figure 1).
+This, along with the other differences in relatedness for *Mixta* genes between nucleotide and amino acid sequences, suggests that the sequence type used in the analysis will give different results, even when using the same genes.
+
+If genetic distance infers the same closest relative for a *Mixta* gene for both nucleotide and amino acid analyses, then these results may be more reliable.
+Thus, for genes where the closest relative is an *Erwinia*, *Tatumella*, or *Enterobacter* species both nucletodie and amino acid analyses, they may have resulted from a horizontal gene transfer (HGT) event rather than vertical evolution from a *Pantoea* species.
+No *Mixta* genes were found to be closely related to *P.
+syringiae* for either the nucleotide or amino acid sequences (Figure 1).
+This confirms that *P.
+syringiae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to the order *Enterobacterales*.
+
+## Best Model vs. Best Available Model
+In MEGAX [@Kumar2018], the user is restricted to what models can be used when estimating genetic distance; however, the user is not restricted when estimating phylogentic trees.
+Model testing in MEGAX tests all of the models that are available for phylogenetic tree analysis.
+Therefore, the recommended model based on lowest BIC value may not be available for genetic distance analysis.
+GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models, as well as the invariant sites (+I) parameter, are not available for genetic distance analysis.
+In order to determine if the limited options would affect the percentage of each species being the closest relative to the *Mixta* genes, genetic distances were estimated for each gene for the nucleotide sequences in MEGA-CC.
+The model testing results were filtered to remove the models that are not available for genetic distances in MEGAX and then distance matrices were estimated for each gene using the best available model based on lowest BIC value.
+As a result, only four of twelve models were available with more than 87% of the genes now using either the Tamura-Nei with a Gamma parameter (TN93_G) or the Tamura 3-parameter with a Gamma parameter (T92_G) (Table 4).
+
+```{r Table4, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
+DM_models_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/DM_Models_NT.csv", stringsAsFactors = FALSE)
+phylo_models_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
+
+models <- data.frame(Model = c("GTR_G", "GTR_G_I", "HKY_G", "HKY_G_I", "K2", "K2_G", "K2_G_I", "K2_I", "T92_G", "T92_G_I", "TN93_G", "TN93_G_I")) %>%
+  mutate(required = c(sum(phylo_models_NT$BM == "GTR_G"), sum(phylo_models_NT$BM == "GTR_G_I"), sum(phylo_models_NT$BM == "HKY_G"),
+                      sum(phylo_models_NT$BM == "HKY_G_I"), sum(phylo_models_NT$BM == "K2"), sum(phylo_models_NT$BM == "K2_G"),
+                      sum(phylo_models_NT$BM == "K2_G_I"), sum(phylo_models_NT$BM == "K2_I"), sum(phylo_models_NT$BM == "T92_G"),
+                      sum(phylo_models_NT$BM == "T92_G_I"), sum(phylo_models_NT$BM == "TN93_G"), sum(phylo_models_NT$BM == "TN93_G_I")),
+         available = c(sum(DM_models_NT$Code == "GTR_G"), sum(DM_models_NT$Code == "GTR_G_I"), sum(DM_models_NT$Code == "HKY_G"),
+                       sum(DM_models_NT$Code == "HKY_G_I"), sum(DM_models_NT$Code == "K2"), sum(DM_models_NT$Code == "K2_G"),
+                       sum(DM_models_NT$Code == "K2_G_I"), sum(DM_models_NT$Code == "K2_I"), sum(DM_models_NT$Code == "T92_G"),
+                       sum(DM_models_NT$Code == "T92_G_I"), sum(DM_models_NT$Code == "TN93_G"), sum(DM_models_NT$Code == "TN93_G_I"))) %>%
+  na_if(0)
+
+colnames(models) <- c("Model", "# of genes requiring each model", "# of genes requiring each available model")
+
+library("knitr")
+kable(models, caption = "Table 4. The number of genes requiring each model when using phylogentic tree analysis vs genetic distance analysis in MEGAX for nucleotide sequences.")
+```
+
+One benefit to using MEGAX is that genetic distance can be exported from phylogenetic trees.
+The disadvantage, however, is that this has to be done manually by opening the tree in MEGAX and exporting the distance matrix as either an Excel or CSV file.
+
+```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended model from model testing vs the restricted model options available for genetic distance analysis in MEGAX for nucleotide sequences.
+The species with the shortest genetic distance from both of the *Mixta* species is considered to be the closest relative.
+Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation.
+Nevertheless, this model may not be available in some software programs or analysis methods within a software program.
+Blue bars represent the percentage of genes that have each species as the closest relative to *Mixta* species when using the recommended model.
+Green bars represent the percentage of genes that have each species as the closest relative to *Mixta* species when using the best available model for genetic distance estimation in MEGA-CC.", fig.width=6.5}
+
+GD_first_relative <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/O8_Results_Ten_NT/Four_Relatives_Ext_Ten_NT.csv", stringsAsFactors = FALSE)
+
+GD_calida <- subset(GD_first_relative, select = c(gene:cal_four)) %>%
+  pivot_wider(names_from = order, values_from = cal_four) %>%
+  mutate(First_check = substring(First_check, first = 1, last = 12),
+         Second = substring(Second, first = 1, last = 12),
+         Third = substring(Third, first = 1, last = 12),
+         Four = substring(Four, first = 1, last = 12),
+         Closest_Relative = case_when(First_check %in% c("Mixta_calida", "Mixta_gavini") ~
+                                                      case_when(Second %in% c("Mixta_calida", "Mixta_gavini") ~ Third,
+                                                                TRUE ~ Second),
+                                                    TRUE ~ First_check))
+
+GD_gaviniae <- subset(GD_first_relative, select = c(gene, order, gav_four)) %>%
+  pivot_wider(names_from = order, values_from = gav_four) %>%
+  mutate(First_check = substring(First_check, first = 1, last = 12),
+         Second = substring(Second, first = 1, last = 12),
+         Third = substring(Third, first = 1, last = 12),
+         Four = substring(Four, first = 1, last = 12),
+         Closest_Relative = case_when(First_check %in% c("Mixta_calida", "Mixta_gavini") ~
+                                                      case_when(Second %in% c("Mixta_calida", "Mixta_gavini") ~ Third,
+                                                                TRUE ~ Second),
+                                                    TRUE ~ First_check))
+
+PD_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE)
+PD_gaviniae <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/four_relatives_gaviniae_NT.csv", stringsAsFactors = FALSE)
+
+unique(GD_calida$gene == GD_gaviniae$gene)
+unique(GD_calida$gene == PD_calida$Gene)
+unique(GD_calida$gene == PD_calida$Gene)
+
+# The plot:
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))  
+```
+
+To assess whether there were any positional effects of the genes and spatial patterns of the closest relatives, circular plots were generated using the *Mixta* gene IDs for nucleotide sequences (Figure 2) and amino acid sequences (Figure 3).
+The gene IDs give the approximate location of each gene relative to the *Mixta* genome.
+The length and colour of a bar represents the species that was the closest relative to the *Mixta* gene at that position.
+Since the closest relatives were the same both *Mixta* species except for 1 and 2 for nucleotide and amino acid sequences, respectively, only the results for *M.
+calida* are shown.
+The differing gene
+
+Since the closest relatives for all but one gene (ttuB where M.
+calida is with T.
+ptyseos and M.
+gaviniae is with P.
+septica) for both *Mixta* species for the nucleotide sequences, only *M.
+calida* is shown
+Amino acids: only two (ttuB where M. calida is with T. saanichensis and M. gaviniae is with P_septica and secY where M. calida is with E. tasmaniensis and M. gaviniae is with P. agglomerans)
+
+To assess whether there were any positional effects of the genes, circular plots were generated using gene IDs for M.
+calida and M.
+gaviniae and the closest relatives (Figures 1 and 2).
+The genes showing greatest identity (lowest distance) to P.
+septica are distributed across the genome for both M.
+calida and M.
+gaviniae (Figures 1 and 2), further suggesting Mixta and Pantoea species have a recent common ancestor.
+If many genes were horizontally transferred in a single event, then large blocks of genes from a single species would be expected.
+However, the Erwinia lines (green) are fairly even throughout the genome and the outgroup lines (red and orange) are separated with genes in between.
+This suggests that Mixta underwent many HGT after speciation from Pantoea.
+
+Positional effects of all of the species and range of the distances were also evaluated (Figures 3 and 4).
+For most genes, the genetic distance for each of the eight species ranged from 0 to 1.5 with only a few greater than 2.
+The cluster of genes around ID 425 in both M.
+calida and M.
+gaviniae have a very small range and so are highly conserved among all of the species.
+These genes were annotated as being ribosomal genes.
+Furthermore, Pantoea species (blue dots) are typically the closest relative to Mixta, followed by Erwinia (green).
+The Tatumella species (purple and pink) and the outgroup species (red and orange) vary as to which are the furthest relative to Mixta.
+
+```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap = "Figure 2.", fig.width=6.5}
+FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE)
+
+FR_nucl_calida <- mutate(FR_nucl_calida,
+                         Relative_Number = case_when(Closest_Relative == "P_septica" ~ 3, Closest_Relative == "P_agglomerans" ~ 4,
+                                                     Closest_Relative == "E_tasmaniensis" ~ 5, Closest_Relative == "E_amylovora" ~ 6,
+                                                     Closest_Relative == "T_ptyseos" ~ 7, Closest_Relative == "T_saanichensis" ~ 8,
+                                                     Closest_Relative == "E_cloacae" ~ 9, Closest_Relative == "P_syringiae" ~ 0))
+
+extra_genes_NT <- data.frame(Gene = NA_character_, First = NA_character_, Second = NA_character_, Third = NA_character_, Fourth = NA_character_,
+                          Mixta_check = NA, Closest_Relative = "P_syringiae", Gene_Length = NA_real_, Beg = NA_real_, End = NA_real_, 
+                          ID = as.numeric((max(FR_nucl_calida$ID) + 1):4084), Relative_Number = 0)
+
+FR_nucl_calida <- rbind(FR_nucl_calida, extra_genes_NT)                                   # Combines M_calida with the extra genes
+
+FR_nucl_calida <- mutate(FR_nucl_calida,
+                         Closest_Relative = factor(Closest_Relative, 
+                                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+# Plot
+nucleotides <- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 5) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringiae")) +
+  scale_y_continuous(limits = c(0, 10), breaks = c(0, 2, 4, 6, 8, 10)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman")); nucleotides
+```
+
+```{r Figure3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap = "Figure 3.", fig.width=6.5}
+FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_AA/four_relatives_calida_AA.csv", stringsAsFactors = FALSE)
+
+FR_aa_calida <- mutate(FR_aa_calida,
+                       Relative_Number = case_when(Closest_Relative == "P_septica" ~ 3, Closest_Relative == "P_agglomerans" ~ 4,
+                                                   Closest_Relative == "E_tasmaniensis" ~ 5, Closest_Relative == "E_amylovora" ~ 6,
+                                                   Closest_Relative == "T_ptyseos" ~ 7, Closest_Relative == "T_saanichensis" ~ 8,
+                                                   Closest_Relative == "E_cloacae" ~ 9, Closest_Relative == "P_syringiae" ~ 0))
+
+extra_genes_AA <- data.frame(Gene = NA_character_, First = NA_character_, Second = NA_character_, Third = NA_character_, Fourth = NA_character_,
+                             Mixta_check = NA, Closest_Relative = "P_syringiae", Gene_Length = NA_real_, Beg = NA_real_, End = NA_real_, 
+                             ID = as.numeric((max(FR_aa_calida$ID) + 1):4084), Relative_Number = 0)
+
+FR_aa_calida <- rbind(FR_aa_calida, extra_genes_AA)                                   # Combines M_calida with the extra genes
+
+FR_aa_calida <- mutate(FR_aa_calida,
+                       Closest_Relative = factor(Closest_Relative, 
+                                                 levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                                            "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+
+# Plot
+amino_acids <- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 5) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringiae")) +
+  scale_y_continuous(limits = c(0, 10), breaks = c(0, 2, 4, 6, 8, 10)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman")); amino_acids
+```
 # Conclusion

### 2020-12-09 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

--- 2020-12-06 Manuscript/Manuscript_Paper.Rmd
@@
 ---
-
+```{r Packages, echo = FALSE, message = FALSE, warning=FALSE}
+library("knitr")
+library("tidyverse")
+library("ggplot2")
+library("RColorBrewer")
+library("extrafont")
+theme_set(theme_classic())
+```
 \newpage
@@
 
-```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
-strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
-                                 "Erwinia tasmaniensis", "Tatumella ptyseos", "Tatumella saanichensis", 
-                                 "Enterobacter cloacae subsp cloacae", "Pseudomonas syringae pv syringae"),
-                     Strain = c("DSM_22759", "DSM_22758", "NBRC_102470", "LMG_5345", "CFBP_1232", "ET1/99", "NCTC_11468",
-                                "NML_06-3099", "ATCC_13047", "ICMP_3023"),
-                     `GenBank Accession No.` = c("GCA_002953215.1", "GCA_002953195.1", "GCA_001598475.1", "GCA_002095575.1", "GCA_000367625.2",
-                                                 "GCA_000026185.1", "GCA_900478715.1", "GCA_000439375.1", "GCA_000025565.1", "GCA_001401075.1"),
-                     Level = c("complete genome", "complete genome", "contigs", "contigs", "contigs", "complete genome", "complete genome", 
-                               "contigs", "complete genome", "scaffold"))
+```{r Table1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table1_Genomes.csv", stringsAsFactors = FALSE)
 
-library("knitr")
 kable(strain, caption = "Table 1. Genomes used in this study.")
@@
 
-```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
-library("tidyverse")
+```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE)
 
-phylo_models_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
-PM_NT <- data.frame(table(phylo_models_NT$BM))
-
-colnames(PM_NT) <- c("Model", "Number")
-
-PM_NT <- mutate(PM_NT,
-             Percent = round((Number / sum(Number)) * 100, digits = 2))
-
-library("knitr")
 kable(PM_NT, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for nucleotide sequences.")
@@
 
-```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
+```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
+PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_PmodelsAA.csv", stringsAsFactors = FALSE)
 
-phylo_models_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_AA/Phylo_Models_AA.csv", stringsAsFactors = FALSE)
-PM_AA <- data.frame(table(phylo_models_AA$BM))
-
-colnames(PM_AA) <- c("Model", "Number")
-
-PM_AA <- mutate(PM_AA,
-                Percent = round((Number / sum(Number)) * 100, digits = 2))
-
-library("knitr")
 kable(PM_AA, caption = "Table 3. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for amino acid sequences.")
@@
 # Results and Discussion
+After the two filters, 799 of of 954 homologous genes remained.
+The genes included those that encoded rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the *M.
+calida* and *M.
+gaviniae* genomes.
 
+## Genetic Distance
+### Closest Relative
+Genetic distance gives a numeric estimate of evolutionary change between species.
+The species that has the shortest genetic distance to a species of interest is considered to be closest relative.
+For the following analysis, model testing was done first to estimate the best model for each gene.
+Next, individual phylogenetic gene trees were created and it was from these that distance matrices were exported.
+*P.
+septica* was the closest relative to 66.3% and 66.5% of the *M.
+calida* and *M.
+gaviniae* genes, respectively, for the nucleotide sequences (Figure 1).
+In comparison, only 42.7% and 42.8% of *M.
+calida* and *M.
+gaviniae* genes, respectively, are closely related to *P.
+septica* for the amino acid sequences (Figure 1), but this is still the majority.
+These results suggest that *P.
+septica* is the closest relative to *Mixta* followed by *P.
+agglomerans*.
+For the amino acid sequences, instead of the *Mixta* genes being more closely related to *P.
+septica*, they are more closely related to the other species in this study (Figure 1).
+For example, the percentage of genes related to *P.
+agglomerans* increased from 13.9% for both of the *Mixta* species to 22.3 and 22.4% to *M.
+calida* and *M.
+gaviniae*, respectively.
+This may be due to a loss of phylogenetic signal when using amino acid sequences as a result of codon degeneracy and having a third of available sites.
+
+```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
+The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using genetic distances.
+The species with the shortest genetic distance from both of the *Mixta* species is considered to be the closest relative.
+Blue bars are the nucleotide sequences of the genes with light blue bars representing the percentage of *M.
+calida* genes and dark blue bars representing *M.
+gaviniae*.
+Green bars are the corresponding amino acid sequences with light green representing *M.
+calida* genes and dark blue bars representing *M.
+gaviniae*.
+Approximately 66% of *M.
+calida* and *M.
+gaviniae* genes were most closely related to *P.
+septica* when using nucleotide sequences in comparison to only 43% when using amino acid sequences.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeMixtaNTAA.csv", 
+                             stringsAsFactors = FALSE) %>%
+  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                              "E. cloacae", "P. syringae"), ordered = TRUE),
+         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
+
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+Unexpectedly, the number of *Mixta* genes that are more closely related increased from 2.9% to 6.0% when using the amino acid sequences instead of the nucleotide sequences (Figure 1).
+This, along with the other differences in relatedness for *Mixta* genes between nucleotide and amino acid sequences, suggests that the sequence type used in the analysis will give different results, even when using the same genes.
+
+If genetic distance infers the same closest relative for a *Mixta* gene for both nucleotide and amino acid analyses, then these results may be more reliable.
+Thus, for genes where the closest relative is an *Erwinia*, *Tatumella*, or *Enterobacter* species both nucleotide and amino acid analyses, they may have resulted from a horizontal gene transfer (HGT) event rather than vertical evolution from a *Pantoea* species.
+No *Mixta* genes were found to be closely related to *P.
+syringiae* for either the nucleotide or amino acid sequences (Figure 1).
+This confirms that *P.
+syringiae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to the order *Enterobacterales*.
+
+### Best Model vs. Best Available Model
+In MEGAX [@Kumar2018], the user is restricted to what models can be used when estimating genetic distance; however, the user is not restricted when estimating phylogenetic trees.
+Model testing in MEGAX tests all of the models that are available for phylogenetic tree analysis.
+Therefore, the recommended model based on lowest BIC value may not be available for genetic distance analysis.
+GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models, as well as the invariant sites (+I) parameter, are not available for genetic distance analysis.
+In order to determine if the limited options would affect the percentage of each species being the closest relative to the *Mixta* genes, genetic distances were estimated for each gene for the nucleotide sequences in MEGA-CC.
+The model testing results were filtered to remove the models that are not available for genetic distances in MEGAX and then distance matrices were estimated for each gene using the best available model based on lowest BIC value.
+As a result, only four of twelve models were available with more than 87% of the genes now using either the Tamura 3-parameter with a Gamma parameter (T92_G) or the Tamura-Nei with a Gamma parameter (TN93_G) (Table 4).
+This is in comparison to the original 58% of genes that required these models when not restricted by availability.
+Additionally, the best model was not available for approximately 43% of the genes (242 genes).
+The mean BIC difference between the best model and the best available model was 2.65 with a range of 0 to 122.6 where 0 means that the best available model is the same as the best model.
+When ignoring cases where the best available model was also the best recommended model, the mean BIC difference was 13.8 with a range of 0.1 to 122.6.
+These difference may result in a less accurate estimation of genetic distance between species.
+
+```{r Table4, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
+models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table4_GenDistModelsNT.csv", stringsAsFactors = FALSE)
+
+kable(models, caption = "Table 4. The number of genes requiring each model when using phylogentic tree analysis vs genetic distance analysis in MEGAX for nucleotide sequences.")
+```
+
+The closest relatives to both *Mixta* species were extracted from the distance matrices.
+*P.
+septica* was estimated to be the closest relative to 12.6 and 14.4% more genes for *M.
+calida* and *M.
+gaviniae*, respectively, when the best available models were used (Figure 2).
+The next largest difference was for genes that were calculated to be most closely related to *P.
+agglomerans* (Figure 2).
+Approximately 5.5 and 7.1% less genes were most closely related to *P.
+agglomerans* when the best available model was used (Figure 2).
+Additionally, there was more disparity in the closest relatives between *> calida* and *M.
+gaviniae* when only the best available model was used (Figure 2).
+When the best model was used, the two *Mixta* typically had the same closest relative for each gene; however, this was the not the case when model options were limited to those available.
+Due to differences in evolutionary change estimation between models and the addition or exclusion of additional parameters, genetic distances may be inaccurately calculated when the best model is not used and such results may even suggest that sibling taxa are not as closely related as originally hypothesized.
+
+```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences.
+The species with the shortest genetic distance from both of the *Mixta* species is considered to be the closest relative.
+Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation.
+Nevertheless, this model may not be available in some software programs or analysis methods within a software program.
+Pink bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the recommended model.
+Orange bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the best available model for genetic distance analysis in MEGAX in which GTR and HKY models and invariant sites parameter are not offered.
+The lighter shades represent *M.
+calida* and the darker shades represent *M.
+gavinaie*.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv", 
+                             stringsAsFactors = FALSE) %>%
+  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                              "E. cloacae", "P. syringae"), ordered = TRUE),
+         Type = factor(Type, levels = c("PD_calida", "PD_gaviniae", "GD_calida", "GD_gaviniae"), ordered = TRUE))
+
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"),
+                    labels = c(expression(paste("Best Model and ", italic("M. calida"), sep = "")), 
+                               expression(paste("Best Model and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("Best Available Model and ", italic("M. calida"), sep = "")), 
+                               expression(paste("Best Available Model and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Model and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+When the closest relative results were compared between when using the best model vs when using the best available, about 20.0 and 22.3% of *M.
+calida* and *M.
+gaviniae* genes were estimated to have a different closest relative.
+Thus, for 82 *M.
+calida* genes and 64 *M.
+gaviniae* genes, the same closest relative was called despite using a different model.
+This may because the best model and the best available for those genes were similar and explained approximately the same amount of variance.
+
+Each model estimates genetic distance differently.
+The best model explains most of the variance between sequences and gives a more accurate estimation of genetic distance.
+Using a model that does not explain most of the variance between sequences will estimate genetic distance less accurately which may result in related taxa (such as those from the same species) having a longer genetic distance between them than in actuality.
+This may also result in the opposite for taxa that are not closely related (i.e.
+*Pantoea* spp.
+and *E.
+cloacae*).
+Even though only two of the nucleotide models and one parameter were unavailable for genetic distance analysis in MEGAX, the results for closest relative to *Mixta* genes was substantially impacted.
+This was particularly the case between the two *Mixta* as one would expect them to have the same closest relative for most of the homologous genes.
+Therefore, using the best model is recommended.
+However, this may not be possible when estimating genetic distance using more than one gene that may require different models.
+Thus, it must be considered that one or more the genes in the concatenation may have a different evolutionary pattern than the others and results should be investigated further to ensure accuracy.
+
+Even though genetic distance estimation in MEGAX did not offer all of the models and parameters as suggested by model testing, all of the models and parameters were available for phylogentic tree analysis and genetic distance can be exported from these trees.
+However, this has to be done manually by opening the tree file in MEGAX and exporting the distance matrix as either an Excel or CSV file and is therefore not feasible when hundreds or thousands of genetic distance matrices are needed.
+
+### Spatial Patterns of Closest Relatives
+To assess whether there were any positional effects of the genes and spatial patterns of the closest relatives, circular plots were generated using the *Mixta* gene IDs for nucleotide sequences (Figure 3) and amino acid sequences (Figure 4).
+The gene IDs give the approximate location of each gene relative to the *Mixta* genome.
+The length and colour of a bar represents the species that was the closest relative to the *Mixta* gene at that position.
+Since the closest relatives were the same both *Mixta* species except for 1 and 2 genes for nucleotide and amino acid sequences, respectively, only the results for *M.
+calida* are shown.
+These genes will be discussed later on.
+
+The genes showing greatest identity (lowest distance) to P.
+septica are distributed across the genome for both M.
+calida and M.
+gaviniae (Figures 1 and 2), further suggesting Mixta and Pantoea species have a recent common ancestor.
+If many genes were horizontally transferred in a single event, then large blocks of genes from a single species would be expected.
+However, the Erwinia lines (green) are fairly even throughout the genome and the outgroup lines (red and orange) are separated with genes in between.
+This suggests that Mixta underwent many HGT after speciation from Pantoea.
+
+Positional effects of all of the species and range of the distances were also evaluated (Figures 3 and 4).
+For most genes, the genetic distance for each of the eight species ranged from 0 to 1.5 with only a few greater than 2.
+The cluster of genes around ID 425 in both M.
+calida and M.
+gaviniae have a very small range and so are highly conserved among all of the species.
+These genes were annotated as being ribosomal genes.
+Furthermore, Pantoea species (blue dots) are typically the closest relative to Mixta, followed by Erwinia (green).
+The Tatumella species (purple and pink) and the outgroup species (red and orange) vary as to which are the furthest relative to Mixta.
+
+```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.", fig.width=6.5, fig.height=5}
+FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+
+ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringiae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.", fig.width=6.5, fig.height=5}
+FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+
+ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringiae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+Nucleotides: ttuB where M. calida is with T. ptyseos and M. gaviniae is with P. septica
+Amino acids: ttuB where M. calida is with T. saanichensis and M. gaviniae is with P. septica 
+             secY where M. calida is with E. tasmaniensis and M. gaviniae is with P. agglomerans
 # Conclusion

### 2020-12-09 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -13,8 +13,18 @@ theme_set(theme_classic())
 
 setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
 #
+### Functions and Frequently used Variables #############################################################################################################
+count_relatives <- function(data) {
+  num_rel <- c(sum(data$Closest_Relative %in% c("P_septica", "Pantoea_sept")), sum(data$Closest_Relative %in% c("P_agglomerans", "Pantoea_aggl")), 
+               sum(data$Closest_Relative %in% c("E_tasmaniensis", "Erwinia_tasm")), sum(data$Closest_Relative %in% c("E_amylovora", "Erwinia_amyl")),
+               sum(data$Closest_Relative %in% c("T_ptyseos", "Tatumella_pt")), sum(data$Closest_Relative %in% c("T_saanichensis", "Tatumella_sa")),
+               sum(data$Closest_Relative %in% c("E_cloacae", "Enterobacter")), sum(data$Closest_Relative %in% c("P_syringae", "Pseudomonas_")))
+}
+
+Spp <- c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+         "E. cloacae", "P. syringae")
+#
 # Table 1. Genomes --------------------------------------------------------------------------------------------------------------------------------------
-# ```{r echo = FALSE, message = FALSE, fig.cap=TRUE}
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
                                  "Erwinia tasmaniensis", "Tatumella ptyseos", "Tatumella saanichensis", 
                                  "Enterobacter cloacae subsp cloacae", "Pseudomonas syringae pv syringae"),
@@ -24,49 +34,57 @@ strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea aggl
                                                  "GCA_000026185.1", "GCA_900478715.1", "GCA_000439375.1", "GCA_000025565.1", "GCA_001401075.1"),
                      Level = c("complete genome", "complete genome", "contigs", "contigs", "contigs", "complete genome", "complete genome", 
                                "contigs", "complete genome", "scaffold"))
-# 
[…]

### 2020-12-09 A_Identity.R
Create A_Identity.R

@@ -1,0 +1,212 @@
+# This is the first R File for this part of the project: nucleotide identity percentage.
+
+# The following code reads in the fasta files and aligns each species' sequence to a Mixta species for each gene. Then it counts the number of 
+# nucleotides that are the same and in the same position (nucleotide identity). The higher the number of identical nucleotides, the more similar that 
+# sequence is to Mixta.
+
+library("tidyverse")
+library("seqinr")
+library("msa")
+library("Rfast")
+
+setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
+
+# Functions ---------------------------------------------------------------------------------------------------------------------------------------------
+align_gene <- function(gene_pair, dna_prot) {                                             # Aligns sequences and converts to a dataframe
+  align <- msa::msaClustalW(inputSeqs = gene_pair, maxiters = 100, 
+                            type = dna_prot, order = "input")                             # Aligns the nucleotide sequences
+  
+  alignConv <- msaConvert(align, type = "seqinr::alignment")                              # Converts the aligned genes into a readable form
+  
+  alignFasta <- data.frame(sequences = alignConv$seq,
+                           species = alignConv$nam)                                       # Covert to format needed for write.fasta()
+}
+
+site_identity <- function(gene_file_df, mixta_spp, dna_prot, fasta_row) {
+  identity <- data.frame(matrix(ncol = 4, nrow = 0))
+  for(row2 in 1:nrow(gene_file_df)) {
+    if(mixta_spp == "calida") {
+      gene_pair <- rbind(gene_file_df[5, ],
+                         gene_file_df[row2, ])                                              # Take two sequences at a time, one being M. calida
+    } else if(mixta_spp == "gaviniae") {
+      gene_pair <- rbind(gene_file_df[6, ],
+                         gene_file_df[row2, ])                                              # Take two sequences at a time, one being M. gaviniae
+    }
+    
+    if(dna_prot == "dna") {
+      gene_pair_DSS <- DNAStringSet(gene_pair$Sequences)                                    # Convert two sequences into DNAStringSet
+      
+    } else if(dna_prot == "protein") {
[…]

### 2020-12-09 A_Nucleotide_Identity.R
Delete A_Nucleotide_Identity.R

@@ -1,109 +1,0 @@
-# This is the first R File for this part of the project: nucleotide identity percentage.
-
-# The following code reads in the fasta files and aligns each species' sequence to a Mixta species for each gene. Then it counts the number of nucleotides
-# that are the same and in the same position (nucleotide identity). The higher the number of identical nucleotides, the more similar that sequence is to
-# Mixta.
-
-library("tidyverse")
-library("seqinr")
-library("msa")
-
-setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
-
-# Functions ----------------------------------------------------------------------------------------------------------------------------------------------
-calida_split_align <- function(gene_file_df, NTAA) {
-  for(row in 1:nrow(gene_file_df)) {
-    gene_pair <- rbind(gene_file_df[5, ],
-                      gene_file_df[row, ])
-    
-    test <- DNAStringSet(gene_pair$Sequences)
-    test@ranges@NAMES <- gene_pair$Names
-    
-    
-    
-    
-    
-    
-    
-    test <- align_gene(gene_pair, NTAA)
-  }
-}
-
-
-
-
-
-
-
-
-
-
-align_gene <- function(gene_pair, NTAA) {                                                 # Aligns sequences and converts to a dataframe
-  align <- msa::msaClustalW(inputSeqs = gene_pair, maxiters = 100, 
-                            type = NTAA, order = "input")                                 # Aligns the nucleotide sequences
-  
-  alignConv <- msaConvert(align, type = "seqinr::alignment")                              # Converts the aligned genes into a readable form
-  
-  alignFasta <- data.frame(sequences = alignConv$seq,
-                           species = alignConv$nam)                                       # Covert to format needed for write.fasta()
-}
-
-my_write.fasta <- function(align_file, file_row) {
-  aligned <- align_file
-  
-  write.fasta(sequences = as.list(aligned$sequences), names = aligned$species,
-              file.out = fastaFiles_fltr$Algn_path[file_row],
-              open = "w", nbchar = 10000, as.string = TRUE)
-}
-
-
-# Nucleotides --------------------------------------------------------------------------------------------------------------------------------------------
-fastaFiles <- data.frame(File_name = list.files(path = "4_Filtered_NT/",
[…]

### 2020-12-09 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
+## Nucleotide and Amino Acid Identity
+Individual species sequences of genes were aligned to both of the corresponding *Mixta* sequences in turn.
+Alignments were done in `R` using the same methods as for gene alignment of the whole gene file.
+The aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@].
+Using base `R` code, a function was created to calculate the number of identical sites between the two sequences over the total number of sites to give a percentage of similar identity.
+Values that equal or are close to 100% are more similar to the *Mixta* sequences.
+
 # Results and Discussion
@@
 
-Even though genetic distance estimation in MEGAX did not offer all of the models and parameters as suggested by model testing, all of the models and parameters were available for phylogentic tree analysis and genetic distance can be exported from these trees.
+Even though genetic distance estimation in MEGAX did not offer all of the models and parameters as suggested by model testing, all of the models and parameters were available for phylogenetic tree analysis and genetic distance can be exported from these trees.
 However, this has to be done manually by opening the tree file in MEGAX and exporting the distance matrix as either an Excel or CSV file and is therefore not feasible when hundreds or thousands of genetic distance matrices are needed.
@@
 
-The genes showing greatest identity (lowest distance) to P.
-septica are distributed across the genome for both M.
-calida and M.
-gaviniae (Figures 1 and 2), further suggesting Mixta and Pantoea species have a recent common ancestor.
-If many genes were horizontally transferred in a single event, then large blocks of genes from a single species would be expected.
-However, the Erwinia lines (green) are fairly even throughout the genome and the outgroup lines (red and orange) are separated with genes in between.
-This suggests that Mixta underwent many HGT after speciation from Pantoea.
-
-Positional effects of all of the species and range of the distances were also evaluated (Figures 3 and 4).
-For most genes, the genetic distance for each of the eight species ranged from 0 to 1.5 with only a few greater than 2.
-The cluster of genes around ID 425 in both M.
-calida and M.
-gaviniae have a very small range and so are highly conserved among all of the species.
-These genes were annotated as being ribosomal genes.
-Furthermore, Pantoea species (blue dots) are typically the closest relative to Mixta, followed by Erwinia (green).
-The Tatumella species (purple and pink) and the outgroup species (red and orange) vary as to which are the furthest relative to Mixta.
+For the nucleotide sequences, there were no discernible patterns in closest relatives to *M.
+calida* genes with the exception of the seven genes that are most closely related to a *Tatumella* species (pink) which are only in the first half of the *M.
+calida* genome (Figure 3).
+Genes that are most closely related to *Pantoea* species (blue) were distributed throughout the genome (Figure 3), which further suggests that *Mixta* and *Pantoea* have a recent ancestor.
+Unexpectedly, genes that are most closely related to other genera, particularly to *E.
+cloacae* (orange) are also distributed throughout the genome (Figure 3).
 
@@
 
+Less genes are closely related to *P.
+septica* when analyzing the amino acid sequences of the *M.
+calida* genes and more are closely related to *P.
+agglomerans*, the *Erwinia* species, the *Tatumella* species, and *E.
+cloacae* (Figure 4).
+Nevertheless, no obvious patterns arose in closest relatives to the *Mixta* genes (Figure 4).
+The *Mixta* genes that have changed their allegiance between the nucleotide and amino acid analyses may be a result of redundancy of the genetic code.
+
 ```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.", fig.width=6.5, fig.height=5}
@@
 
+If many genes were horizontally transferred in a single event, then large blocks of genes from a single species would be expected.
+However, this is not the case for any of species.
+There are several hypotheses as to why.
+One is that it is possible that the *Mixta* genome underwent many HGT events after speciation from *Pantoea*; however, further research into the origins of these genes and the frequency of HGT events with *Mixta* and other genera within the *Erwiniaceae* is needed.
+Another hypothesis is that the gene sequence is so well-conserved that the genetic distance between *Mixta* and more than one of the other species are near identical.
+Since only one species was extracted for the closest relative to each *Mixta* gene, it is possible that a species was called at random if there was tie between more than one species for having the shortest distance to *Mixta*.
+For example, if the genetic distance from *M.
+calida* to both *P.
+septica* and *E.
+amylovora* for a single gene was 0.012 and this was the shortest distance for that gene, then both species are tied for closest relative for that gene.
+However, since only one can be determined as closest relative, then closest relative may have been decided by the computer program based on another characteristic such as alphabetical order.
+Therefore, quantitative analysis or computing a phylogenetic tree to view the two-dimensional results is needed.
+
+## Nucleotide and Amino Acid Identity
+### Nucleotide Identity
+
+### Amino Acid Identity
+
+## A Comparison Between Evolutionary Change and Site Identity
+
+## Phylogenetic Trees
+
 Nucleotides: ttuB where M. calida is with T. ptyseos and M. gaviniae is with P. septica

### 2021-01-19 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -1,6 +1,6 @@
 # This is the sixth R File for this project.
 
-# The following code analyzes the distance matrices.
+# The following code analyzes the results from the distance matrices and creates the tables and figures for the paper.
 
 library("tidyverse")
 library("msa")
@@ -21,8 +21,6 @@ count_relatives <- function(data) {
                sum(data$Closest_Relative %in% c("E_cloacae", "Enterobacter")), sum(data$Closest_Relative %in% c("P_syringae", "Pseudomonas_")))
 }
 
-Spp <- c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
-         "E. cloacae", "P. syringae")
 #
 # Table 1. Genomes --------------------------------------------------------------------------------------------------------------------------------------
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
@@ -351,28 +349,32 @@ FR_nucl_calida <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsA
   mutate(Relative_Number = case_when(Closest_Relative == "P_septica" ~ 3, Closest_Relative == "P_agglomerans" ~ 4,
                                      Closest_Relative == "E_tasmaniensis" ~ 5, Closest_Relative == "E_amylovora" ~ 6,
                                      Closest_Relative == "T_ptyseos" ~ 7, Closest_Relative == "T_saanichensis" ~ 8,
-                                     Closest_Relative == "E_cloacae" ~ 9, Closest_Relative == "P_syringiae" ~ 0))
+                                     Closest_Relative == "E_cloacae" ~ 9, Closest_Relative == "P_syringae" ~ 0))
 
 extra_genes_NT <- data.frame(Gene = NA_character_, First = NA_character_, Second = NA_character_, Third = NA_character_, Fourth = NA_character_,
-                             Mixta_check = NA, Closest_Relative = "P_syringiae", Gene_Length = NA_real_, Beg = NA_real_, End = NA_real_, 
+                             Mixta_check = NA, Closest_Relative = "P_syringae", Gene_Length = NA_real_, Beg = NA_real_, End = NA_real_, 
[…]

### 2021-01-19 A_Identity.R
Update A_Identity.R

@@ -1,4 +1,4 @@
-# This is the first R File for this part of the project: nucleotide identity percentage.
+# This is the first R File for this part of the project: nucleotide and amino acid identity percentage.
 
 # The following code reads in the fasta files and aligns each species' sequence to a Mixta species for each gene. Then it counts the number of 
 # nucleotides that are the same and in the same position (nucleotide identity). The higher the number of identical nucleotides, the more similar that 

### 2021-01-19 B_Results_Analysis.R
Create B_Results_Analysis.R

@@ -1,0 +1,214 @@
+# This is the second R File for this part of the project: nucleotide and amino acid identity percentage.
+
+# The following code analyzes the results from the identity percentages and creates the tables and figures for the paper.
+
+library("tidyverse")
+library("msa")
+library("ggplot2")
+library("cowplot")
+library("RColorBrewer")
+library("extrafont")
+# loadfonts(device = "win")
+theme_set(theme_classic())
+
+setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
+
+### Functions and Frequently used Variables #############################################################################################################
+count_relatives <- function(data) {
+  num_rel <- c(sum(data$Closest_Relative %in% c("P_septica", "Pantoea_sept")), sum(data$Closest_Relative %in% c("P_agglomerans", "Pantoea_aggl")), 
+               sum(data$Closest_Relative %in% c("E_tasmaniensis", "Erwinia_tasm")), sum(data$Closest_Relative %in% c("E_amylovora", "Erwinia_amyl")),
+               sum(data$Closest_Relative %in% c("T_ptyseos", "Tatumella_pt")), sum(data$Closest_Relative %in% c("T_saanichensis", "Tatumella_sa")),
+               sum(data$Closest_Relative %in% c("E_cloacae", "Enterobacter")), sum(data$Closest_Relative %in% c("P_syringae", "Pseudomonas_")))
+}
+
+# Figure 5. First Relatives Percentage ------------------------------------------------------------------------------------------------------------------
+datasets <- data.frame(project = c("NT", "NT", "AA", "AA"),
+                       dataset = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"))
+
+number_relatives <- data.frame(matrix(ncol = 4, nrow = 0))
+
+for(row in 1:nrow(datasets)) {
+  four_rel <- read.csv(paste("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_", 
+                             datasets$project[row], "/four_identities_",
+                             datasets$dataset[row], ".csv", sep = ""), 
+                       stringsAsFactors = FALSE)
+  
+  num_rel <- data.frame(Species = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                    "E. cloacae", "P. syringae"),
[…]

### 2021-01-19 i_Tree_Genes.R
Create i_Tree_Genes.R

@@ -1,0 +1,6 @@
+# This is all for phylogenetic trees
+
+# atp, gyr, inf, and rpo Genes ----------------
+fastaFileNT <- data.frame(File_name = list.files(path = "5_2_Renamed_NT/", 
+                                                 pattern = ".fasta")) %>%
+  mutate(Path_name = paste("5_2_Renamed_NT", File_name, sep = "/"))

### 2021-01-19 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 ---
-title: "Untitled"
+title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
 author: "Kimberly Hinz"
@@
 Depending on the genes and method used for phylogenetic analysis, the genus with which *Mixta* shares the most recent ancestor differs.
-This study aimed to determine the cause behind these contentious results, and identified genes that may have resulted from horizontal gene transfer.
+This study aimed to determine the cause behind these contentious results and identified genes that may have resulted from horizontal gene transfer.
 Ten complete or partial genomes – two *Mixta* species, six species from other *Erwiniaceae* genera, and two outgroups – were retrieved from NCBI and annotated.
 Homologous genes were extracted yielding a dataset of 799 genes.
-The genes were aligned With the ClustalW algorithm and MEGA-CC was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
+The genes were aligned with the ClustalW algorithm, and MEGA-CC was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
 Distance matrices were then extracted from these trees.
-Nucleotide and amino acid identity analyses were also done in `R` programming language.
+Nucleotide and amino acid identity analyses were also done using the programming language `R`.
 *Pantoea* was the closest relative to the *Mixta* species in most of the analyses; however, results were not consistent across analyses.
 Some genes were also found to be more similar to other, non-*Pantoea* species.
-Diligence must be given to the selection of genes for phylogenetic analysis and to the method chosen to prevent xenologous signal from distorting the true relationships.
-Furthermore, future research should consider that different phylogenetic analyses may give different results.
+Diligence must be given to selecting genes for phylogenetic analysis and the method chosen to prevent any xenologous signal from distorting the actual relationships.
+Furthermore, future research should consider that different phylogenetic analyses may provide different results.
 
@@
 # Introduction
-The bacterial family Erwiniaceae includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, *Kalamiella* (LPSN, http://www.bacterio.net/).
-Three of these, *Mixta*, *Pantoea*, and *Erwinia*, have been the focus of numerous studies given their relevance to plant and human pathogens.
+The bacterial family Erwiniaceae includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms, including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, *Kalamiella* (LPSN, http://www.bacterio.net/).
+Numerous studies have focused on three of these genera – *Mixta*, *Pantoea*, and *Erwinia* – given their relevance to plant and human pathogens.
 For example, the genus *Erwinia* contains plant pathogens (e.g.
-E.
-amylovora) [@HaubenSwings2005; @Lopez2011; @Matsuura2012], epiphytes (e.g.
-E.
-tasmaniensis) [@Geider2006; @Mergaert1999], species that may contribute to plant disease [@Buonaurio2015], and others that are insect-associated [@Campillo2015; @Harada1997; @Skrodenyte-Arbaciauskiene2012].
+*E.
+amylovora*) [@HaubenSwings2005; @Lopez2011; @Matsuura2012], epiphytes (e.g.
+*E.
+tasmaniensis*) [@Geider2006; @Mergaert1999], species that may contribute to plant disease [@Buonaurio2015], and others that are insect-associated [@Campillo2015; @Harada1997; @Skrodenyte-Arbaciauskiene2012].
 Similarly, *Pantoea*, a genus reported by Gavini et al.[@Gavini1989], includes species that inhabit a variety of environments [@Gavini1989; @Popp2010; @Liu2013], including human and plant pathogens [@Prakash2015; @deBaere2004; @Fullerton2007; @Kratz2003; @Lim2006; @Schmid2003; @vanRostenberghe2006] as well as species associated with plant growth [@Smits2011; @Kim2012], insects [@Palmer2016], and fungi [@Ma2016; @Rong2016].
 *Mixta*, originally *Pantoea*, is a novel genus first suggested by Palmer et al.
-[@Palmer2017] and then later proposed by the same researchers in 2018 [@Palmer2018].
-Two species, M.
-calida and M.
-gaviniae (previously species of *Pantoea*), were originally found in infant formula and its production environment [@Popp2010].
+[@Palmer2017] and later proposed by the same researchers in 2018 [@Palmer2018].
+*M.
+calida* and *M.
+gaviniae* (previously species of *Pantoea*) were initially found in infant formula and its production environment [@Popp2010].
 Fritz et al.
-[@Fritz2014] demonstrated that some strains of M.
-calida were linked with post-surgical meningitis in humans.
+[@Fritz2014] demonstrated that some strains of *M.
+calida* were linked with post-surgical meningitis in humans.
 Currently, there are only five *Mixta* species validly published (LPSN, http://www.bacterio.net/).
 
-Understanding the phylogeny of bacterial species and genes allows researchers to investigate the evolutionary history of the species, their relatedness, and the individual genes that allow the species to survive under various conditions [@Palmer2017; @Bennett2012; @Prasanna2013; @Angus2014; @deMaayer2014; @Fouts2016].
-The most common methods in phylogenetic analysis are those including 16S ribosomal RNA (rRNA) sequences [@Konstantinidis2007], multi-locus sequence analysis (MLSA) with several housekeeping genes [@Konstantinidis2007; @Gevers2005; @Glaeser2015], and ribosomal MLSA with 53 structural ribosomal proteins [@Palmer2017; @Bennett2012; @Jolley2012].
-However, the phylogenetic trees generated from these methods may not be very robust [@Konstantinidis2007; @Glaeser2015; @Brady2008] because of the lower phylogenetic signal from the highly conserved housekeeping genes [@Konstantinidis2007; @Gevers2005; @Fox1992; @Konstantinidis2005; @Staley2006; @RichterRosselloMora2009].
-This is particularly seen within Erwiniaceae wherein most of our understanding of the phylogeny and interrelationships are based on the 16S rRNA genes [@Francino2006; @Hauben1998; @Naum2008; @Sproer1999; @Adeolu2016] because of its ubiquity.
-However, it has insufficient resolution at the species level [@Glaeser2015] and the genera *Pantoea*, *Erwinia*, *Tatumella*, and *Mixta* are often not monophyletic [@Rezzonico2016].
-Another issue that may arise with MLSA is unknowingly including paralogues (i.e.
-genes originating from an intragenomic duplication event) or xenologues (i.e.
-genes originating from horizontal gene transfer) [@Palmer2017; @Koonin2005].
+Understanding the phylogeny of bacterial species and genes allows researchers to investigate the species' evolutionary histories, relatedness, and the individual genes that enable the species to survive under various conditions [@Palmer2017; @Bennett2012; @Prasanna2013; @Angus2014; @deMaayer2014; @Fouts2016].
+The most common methods in phylogenetic analysis are those that use ubiquitous genes including 16S ribosomal RNA (rRNA) sequences [@Konstantinidis2007], multilocus sequence analysis (MLSA) with several housekeeping genes [@Konstantinidis2007; @Gevers2005; @Glaeser2015], and ribosomal MLSA with 53 structural ribosomal proteins [@Palmer2017; @Bennett2012; @Jolley2012].
+However, the phylogenetic trees generated from these methods may not be robust [@Konstantinidis2007; @Glaeser2015; @Brady2008] because of the lower phylogenetic signal from the highly conserved genes [@Konstantinidis2007; @Gevers2005; @Fox1992; @Konstantinidis2005; @Staley2006; @RichterRosselloMora2009].
+This issue is particularly the case with *Erwiniaceae* because much of our understanding is based on the 16S rRNA and other housekeeping genes [@Francino2006; @Hauben1998; @Naum2008; @Sproer1999; @Adeolu2016].
+However, such analyses have insufficient resolution at the species level [@Glaeser2015], and the genera *Pantoea*, *Erwinia*, *Tatumella*, and *Mixta* are often not monophyletic [@Rezzonico2016].
+Another issue is the unintentional inclusion of paralogues (originating from an intragenomic duplication event) or xenologues (arising from horizontal gene transfer) in MLSA gene analyses [@Palmer2017; @Koonin2005].
 
-The DNA topoisomerase (gyrase) subunit (gyrB), DNA-directed RNA polymerase subunit (rpoB), ATP synthase subunit (atpD), and initiation translation factor (infB) have been shown to be more reliable than 16S rRNA gene sequences for determining intra- and intergeneric relationships [@Hedegaard1999; @Dauga2002; @Mollet1997; @Paradis2005] and a robust method for the Erwiniaceae family [@Brady2010].
-The genera within Erwiniaceae are monophyletic when concatenated partial gyrB, rpoB, atpD, and infB gene sequences are used [@Brady2008; @Brady2009; @Moretti2011].
+The DNA gyrase subunit (gyrB), DNA-directed RNA polymerase subunit (rpoB), ATP synthase subunit (atpD), and initiation translation factor (infB) are more reliable than 16S rRNA gene sequences for determining intra- and intergeneric relationships [@Hedegaard1999; @Dauga2002; @Mollet1997; @Paradis2005] and a robust method for the Erwiniaceae family [@Brady2010].
+The genera within Erwiniaceae are monophyletic when the analysis uses concatenated partial gyrB, rpoB, atpD, and infB gene sequences [@Brady2008; @Brady2009; @Moretti2011].
 
-Nevertheless, despite efforts to find a suitable method for identifying the species tree for Erwiniaceae genera, *Mixta*’s position with relation to *Pantoea*, *Erwinia*, and *Tatumella* vastly differs between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
+Nevertheless, despite efforts to find a suitable method for identifying the species tree for Erwiniaceae genera, *Mixta*’s position relative to *Pantoea*, *Erwinia*, and *Tatumella* vastly differs between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
 Not only do different gene combinations provide different results, but different algorithms and models for alignments and phylogenetic trees also contribute to the contentious results [@Naum2008; @Octavia2014].
@@
 Three of the studies – Rezzonico et al., Prakash et al., and Gueule et al.
-– placed the *Mixta* species within the *Pantoea* genera whereas the analysis in Palmer et al.
-placed *Mixta* outside of the other three genera.
+– placed the *Mixta* species within the *Pantoea* genera.
+In contrast.
+the analysis in Palmer et al.
+identified *Mixta* outside of the other three genera.
 Palmer et al.
@@
 [@Brady2012] also ran analyses with the four MLSA genes as amino acid sequences.
+
+These three studies showed contentious results as *Mixta* was said to be related closest with *Tatumella* by Rezzonico et al., with *Pantoea* by Palmer et al., and within the *Pantoea* clade by
+
+These three studies showed contentious results as
+
 These three studies show contentious results as Rezzonico et al.
@@
 found *Mixta* within the *Pantoea* genus.
+
 In a separate study conducted by Palmer et al.
@@
 The other reason may be horizontal gene transfer (HGT) events within the *Erwiniaceae* or from a non-closely-related species.
-HGT events occur frequently within bacteria and often the genes provide the bacteria the ability to adapt to new environments and/or survive selective pressures.
+HGT events frequently occur within bacteria, and often the genes provide the bacteria with the ability to adapt to new environments and/or survive selective pressures.
 
 The objective of this paper was to determine the cause behind the contentious phylogenetic results for the *Mixta* species.
-To accomplish this, homologous genes were extracted from ten complete or partial genomes - two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*, and two outgroups - that were retrieved from public databases.
+Homologous genes were extracted from ten complete or partial genomes - two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*, and two outgroups - that were retrieved from public databases.
 In this study, both the nucleotide and amino acid sequences for the same genes were separately analyzed.
-Model testing was done for each gene in the dataset and phylogenetic trees were computed using the recommended model.
-Distance matrices were exported from the phylogenetic trees and these were used to determine which species were most closely related to *Mixta* for each gene.
+Model testing was done for each gene in the dataset, and phylogenetic trees were computed using the recommended model.
+Distance matrices were exported from the phylogenetic trees, and these were used to determine which species were most closely related to *Mixta* for each gene.
 Nucleotide and amino acid identity analyses were also conducted for each gene in order to compare the results between evolutionary history and sequence similarity.
@@
 Publicly available full genomes were retrieved from NCBI comprising of ten type strains from *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Enterobacter*, and *Pseudomonas* (Table 1).
-*Enterobacter cloacae* is in the same order of *Enterobacterales* as *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella* whereas *Pseudomonas syringiae* is in the same class of *Gammaproteobacteria*.
-The genomes were annotated with Prokka v1.14.1 [@Seemann2014] using default parameters and homologous genes were extracted using GET_HOMOLOGUES software package [@Contreras-Moreira2013] with the bidirectional best-hit search algorithm using default parameters.
+*Enterobacter cloacae* is in the same order of *Enterobacterales* as *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella* whereas *Pseudomonas syringae* is in the same class of *Gammaproteobacteria*.
+The genomes were annotated with Prokka v1.14.1 [@Seemann2014] using default parameters, and homologous genes were extracted using GET_HOMOLOGUES software package [@Contreras-Moreira2013] with the bidirectional best-hit search algorithm using default parameters.
 GET_HOMOLOGUES extracted 954 homologous genes between the ten species.
@@
 ## Genetic Distance
-### Closest Relative
+### Closest Relative from Genetic Distances
 Genetic distance gives a numeric estimate of evolutionary change between species.
@@
 No *Mixta* genes were found to be closely related to *P.
-syringiae* for either the nucleotide or amino acid sequences (Figure 1).
+syringae* for either the nucleotide or amino acid sequences (Figure 1).
 This confirms that *P.
-syringiae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to the order *Enterobacterales*.
+syringae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to the order *Enterobacterales*.
 
@@
 
-```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.", fig.width=6.5, fig.height=5}
+```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
+Closest relatives to *M.
+calida* genes for nucleotide sequences around the *M.
+calida* genome.
+Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
+The *M.
+calida* genome has 4084 genes.
+No distinct patterns emerged; however, genes that are most closely related to *P.
+septca* are the most common and occur throughout the entire genome.
+Therefore, *P.
+septica* and *Mixta* are likely to have a common recent ancestor.", fig.width=6.5, fig.height=5}
 FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
@@
                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
-                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
 
@@
                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringiae")) +
+                               "P. syringae")) +
   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
@@
 
-```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.", fig.width=6.5, fig.height=5}
+```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.
+Closest relatives to *M.
+calida* genes for amino acid sequences around the *M.
+calida* genome.
+Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
+The *M.
+calida* genome has 4084 genes.
+No distinct patterns emerged; however, genes that are most closely related to *P.
+septca* are the most common and occur throughout the entire genome.
+Therefore, *P.
+septica* and *Mixta* are likely to have a common recent ancestor.", fig.width=6.5, fig.height=5}
 FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
@@
 ## Nucleotide and Amino Acid Identity
-### Nucleotide Identity
+### Most Similar Taxa from Site Identity
+Site identity is here defined as the amount of similarity between two sequences.
+If two sequences have a site identity of 100%, then they are identical.
+If site identity is 99%, then there is up to a few bases or amino acids that are different between the two sequences depending on the length of the sequences.
+Within each gene file, the ten sequences were aligned to both *M.
+calida* and *M.
+gaviniae* in turn and the site identity was calculated.
+Site identity was divided by the length of aligned genes to give a percent identity that could be compared across the genes and between the nucleotide and amino acid sequences.
+From these values, the closest non-*Mixta* relative was extracted for each gene.
 
-### Amino Acid Identity
+The majority of the *Mixta* sequences were most identical with *P.
+septica* for nucleotide sequences (Figure 5) with 80.6 and 79.3% of *M.
+calida* and *M.
+gaviniae* sequences, respectively.
+This is an increase of approximately 15% from the genetic distance results for the closest relative (Figure 1).
+The rest of the genes are most similar to *P.
+agglomerans*, followed by the *Erwinia* species, *E.
+cloacae*, and then the *Tatumella* species.
+Since mutations that do not disrupt the bacterium's ability to thrive and reproduce do not occur frequently and yet may occur at constant rate, then these results suggest that *P.
+septica* has the most recent common ancestor with the *Mixta* species.
 
-## A Comparison Between Evolutionary Change and Site Identity
+```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5.
+The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using percent site identities.
+The species with the highest percent identity to each of the *Mixta* species is considered to be most similar taxon for a single gene.
+Blue bars are the results for the nucleotide sequences where the light blue bars represent *M.
+calida* sequences and the dark blue bars represent *M.
+gaviniae* sequences.
+Green bars are the results for the amino acid sequences where the light green bars represent *M.
+calida* sequences and the dark green bars represent *M.
+gaviniae* sequences.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5_FirstRelativeMixtaNTAAIdentity.csv", 
+                             stringsAsFactors = FALSE) %>%
+  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                              "E. cloacae", "P. syringae"), ordered = TRUE),
+         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
+
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+In contrast, only 47.8 and 47.6% of *M.
+calida* and *M.
+gaviniae* sequences were most similar to *P.
+septica* for the amino acid sequences (Figure 5).
+Instead, the genes are most similar to either *P.
+agglomerans* or one of the two *Erwinia* species (Figure 5).
+Additionally, 3.1 and 2.5% more genes were most similar to the *E.
+cloacae* sequences for *M.
+calida* and *M.
+gaviniae*, respectively (Figure 5).
+In the identity analyses, amino acids that are functionally equivalent were not considered.
+This may explain the large differences between the nucleotide and amino acid identities for which species was most similar to the *Mixta* sequences.
+Additionally, at every site for a nucleotide sequence, each base has a 1:3 ratio of being identical to the base at the corresponding location of the other sequence whereas there is only a 1:19 ratio of an amino acid being identical to the amino acid at the same position of the other sequence.
+Thus, it is far more likely for a nucleotide at a random position to be identical the nucleotide at the same position of the other sequence.
+This may lead to problems when only amino acid sequences are considered, especially if functional equivalence is not considered.
+
+Although the numbers of genes that are most similar to *P.
+septica* are similar between the two *Mixta* species for both nucleotide and amino acid sequences (Figure 5), there were many genes in which the same species was the not the most similar taxa to both *Mixta* species.
+There were 86 genes for the nucleotide sequences and 147 genes for the amino acid sequences.
+However, of the 86 genes for nucleotide sequences, it was the case for 52% that one of the *Mixta* sequences was most similar to one representative of a genus and the other *Mixta* sequences was most similar to the other species of the same genus.
+This was also the case for 62% of the 147 amino acid genes.
+
+Furthermore, percent identity analyses give only how similar are two sequences.
+They do not use evolutionary models or penalize transitions and transversions in nucleotide sequences.
+However, because mutations that do not disrupt any vital processes are rare, percent identity analyses may lend support to relatedness between two taxa.
+
+### Spatial Patterns of Most Similar Taxa
+Similar to what was done with the closest relatives in the genetic distance analysis, spatial patterns of the most similar taxa to the *M.
+calida* sequences based on site identity were investigated for the nucleotide sequences (Figure 6) and the amino acid sequences (Figure 7).
+Once again, no discernible patterns arose when spatial position on the *M.
+calida* genome are considered when either the nucleotide or amino acid sequences are analyzed (Figure 6 and 7).
+However, majority of the nucleotide genes that were most similar to the two *Tatumella* species and to *E.
+amylovora* occur in the first half of the genome (Figure 6).
+
+```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6.
+The most similar taxa to *M.
+calida* genes for nucleotide sequences around the *M.
+calida* genome.
+The most similar taxa is defined to be that which sequence is most identical to the *M.
+calida* sequence.
+Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5}
+FI_calida_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6_CircularIdentities_NT.csv", 
+                         stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+
+ggplot(FI_calida_NT, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Most Similar Taxa") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+```{r Figure7, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 7.
+The most similar taxa to *M.
+calida* genes for amino acid sequences around the *M.
+calida* genome.
+The most similar taxa is defined to be that which sequence is most identical to the *M.
+calida* sequence.
+Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5}
+FI_calida_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_CircularIdentities_AA.csv", 
+                         stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+
+ggplot(FI_calida_AA, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Most Similar Taxa") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+While no distinct patterns emerged, genes that are most similar to *P.
+septica* predominates the whole of the *M.
+calida* genome (Figures 6 and 7).
+This suggests that *Mixta* and *P.
+septica* may be related as their sequences are similar and few mutations have occurred after they diverged.
+
+## A Comparison Between Evolutionary Change and Site Identity (Keep?)
+### *Mixta calida*
+Since all four of the analyses gave contentious results, it was necessary to compare the results for the first relative from genetic distance and the most similar taxa from site identity for both the nucleotide and amino acid sequences.
+For approximately 39% of the *M.
+calida* genes, all four analyses gave the same species.
+Of these genes, 255 agreed with *P.
+septica*, 22 with *P.
+agglomerans*, 17 with *E.
+tasmaniensis*, 8 with *E.
+amylovora*, 1 with *T.
+saanichensis*, and 6 with *E.
+cloacae*.
+Additionally, the four analyses gave the same genus approximately 63% of the genes where in 454 agreed with *Pantoea* followed by *Erwinia* with 44, and *Tatumella* with 3.
+Thus, because all four of the analyses agreed on a single species or genus, these results are more likely to be correct.
+
+The two analyses conducted with nucleotide sequences agreed on a species 78% of the time whereas the two analyses conducted with amino acid sequences agreed only 70% of the time.
+It is not unexpected that the two nucleotide analyses would agree on taxa more often than the amino acid analyses since there are more sites in nucleotide sequences and no degeneracy in the code.
+
+```{r Figure8, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 8. Sample for first 100 genes.", fig.width=6.5, fig.height=3}
+all_results_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureN_AllResultsCalida.csv", 
+                               stringsAsFactors = TRUE) %>%
+  pivot_longer(cols = c(Rel_NT:Ide_AA), names_to = "Analysis", values_to = "Closest_Similar") %>%
+  mutate(Closest_Similar = factor(Closest_Similar, 
+                                  levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", 
+                                             "E_cloacae", "P_syringiae"), 
+                                  ordered = TRUE),
+         Analysis = factor(Analysis, levels = c("Ide_AA", "Rel_AA", "Ide_NT", "Rel_NT"), ordered = TRUE),
+         Gene_Name = reorder(Gene_Name, ID))
+
+ggplot(all_results_calida[1:400, ], aes(x = Gene_Name, y = Analysis, fill = Closest_Similar)) +
+  geom_tile() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_discrete(labels = c("AA Identity", "AA Closest", "NT Identity", "NT Closest")) +
+  labs(fill = "Closest Relative or \n Most Similar Taxa") +
+  theme(legend.text = element_text(face = "italic"),
+        # legend.position = "bottom",
+        text = element_text(size = 12,  family = "Times New Roman"),
+        axis.text.x = element_text(angle = 45, hjust = 1))               # Rotates x axis labels
+
+stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE)
+
+# table(stats_on_all$Same_Species)
+# table(stats_on_all$Same_Genus)
+# table(stats_on_all$NT_Same)
+# table(stats_on_all$AA_Same)
+# table(stats_on_all$Rel_Same)
+# table(stats_on_all$Ide_Same)
+```
+Visual comparison of the results from the four analyses does not reveal any patterns (Figure 8).
+The four analyses are plotted on y axis where NT Closest is the closest relative according to genetic distance for nucleotide sequences  (Figure 8), NT Identity is the most similar taxa according to percent identity of nucleotide sequences, AA Closest is the closest relative according to genetic distance for amino acid sequences, and AA Identity is the most similar taxa according to percent identity of amino acid sequences (Figure 8).
+The first 100 genes according to *M.
+calida* gene ID are plotted on the x axis and the colour of the block represents the species that is the closest relative or most similar taxa to that gene for that analysis (Figure 8).
 
 ## Phylogenetic Trees
+### Single Gene Trees
+From the genetic distance analyses, only one gene separated *M.
+calida* and *M.
+gaviniae* for the nucleotide sequence and two genes for the amino acid sequence.
+The one gene for the nucleotide sequences was ttuB.
+According to genetic distance, *M.
+calida* was more closely related to *T.
+ptyseos* and *M.
+gaviniae* was more closely related to *P.
+septica* rather than to each other.
+A phylogenetic tree of this gene confirms that *M.
+calida* is more closely related to the *Tatumella* species where *M.
+gaviniae* to *Pantoea* and *Erwinia* (Figure 9A).
+The gene ttuB also had the two *Mixta* species separated for the amino acid sequences and the resulting tree shows the same results (Figure 9B).
 
-Nucleotides: ttuB where M. calida is with T. ptyseos and M. gaviniae is with P. septica
-Amino acids: ttuB where M. calida is with T. saanichensis and M. gaviniae is with P. septica 
-             secY where M. calida is with E. tasmaniensis and M. gaviniae is with P. agglomerans
+```{r Figure9, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 9. Maximum likelihood gene trees for ttuB for nucleotide (A) and amino acid sequences (B).", fig.width=3.25}
+knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_NT.png")
+knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_AA.png")
+```
+
+The second gene that had the two *Mixta* species separated for the amino acid sequences was secY.
+From the nucleotide sequences, all of the taxa except for *P.
+syringae* are very closely related and bifurcations cannot be determined from the radial tree (Figure 10A).
+The same is the case for the amino acid sequences (Figure 10B).
+Condensed tree at 70% cutoff was created to show the relationships for both the nucleotide sequences (Figure 10C) and the amino acid sequences (Figure 10D).
+In the nucleotide tree, the two *Mixta* species are paired together, yet they are placed in a group with the two outgroups and the *Erwinia* species (Figure 10C).
+A condensed tree for the amino acid sequences showed that birfucations between most of the species were not supported above the 70% cutoff.
+These gene trees suggest that the differences between the *Enterobacterales* are not great enough to accurately infer the *Mixta* relative.
+
+```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10.
+Maximum likelihood gene trees for secY for nucleotide (A) and amino acid sequences (B).
+Since bifucations could not be seen in radial trees, condensed trees were computed with a 70% cutoff for both the nucleotide sequnces (C) and amino acid sequences (D).", fig.width=3.25}
+knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_NT.png")
+knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_AA.png")
+knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_NT70.png")
+knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_AA70.png")
+```
+
 # Conclusion
+Since the analyses provided different results for many of the genes, a single analysis may not be enough to conclude that a species is related to another.
 

### 2021-01-20 Manuscript/Manuscript_Paper.Rmd
New edits

@@
 # Abstract
-*Mixta* is a recently described genus within the bacterial family *Erwiniaceae* that is closely related to *Pantoea*, *Erwinia*, and *Tatumella*.
+*Mixta* is a recently described genus within the bacterial family *Erwiniaceae* and is closely related to *Pantoea*, *Erwinia*, and *Tatumella*.
 Depending on the genes and method used for phylogenetic analysis, the genus with which *Mixta* shares the most recent ancestor differs.
@@
 Nucleotide and amino acid identity analyses were also done using the programming language `R`.
-*Pantoea* was the closest relative to the *Mixta* species in most of the analyses; however, results were not consistent across analyses.
+*Pantoea* was the closest relative to the *Mixta* species in most analyses; however, results were not consistent.
 Some genes were also found to be more similar to other, non-*Pantoea* species.
@@
 
-Understanding the phylogeny of bacterial species and genes allows researchers to investigate the species' evolutionary histories, relatedness, and the individual genes that enable the species to survive under various conditions [@Palmer2017; @Bennett2012; @Prasanna2013; @Angus2014; @deMaayer2014; @Fouts2016].
+Understanding the phylogeny of bacterial species and genes allows researchers to investigate the species' evolutionary histories, relatedness, and the individual genes that enable them to survive under various conditions [@Palmer2017; @Bennett2012; @Prasanna2013; @Angus2014; @deMaayer2014; @Fouts2016].
 The most common methods in phylogenetic analysis are those that use ubiquitous genes including 16S ribosomal RNA (rRNA) sequences [@Konstantinidis2007], multilocus sequence analysis (MLSA) with several housekeeping genes [@Konstantinidis2007; @Gevers2005; @Glaeser2015], and ribosomal MLSA with 53 structural ribosomal proteins [@Palmer2017; @Bennett2012; @Jolley2012].
@@
 This issue is particularly the case with *Erwiniaceae* because much of our understanding is based on the 16S rRNA and other housekeeping genes [@Francino2006; @Hauben1998; @Naum2008; @Sproer1999; @Adeolu2016].
-However, such analyses have insufficient resolution at the species level [@Glaeser2015], and the genera *Pantoea*, *Erwinia*, *Tatumella*, and *Mixta* are often not monophyletic [@Rezzonico2016].
+However, such analyses have an insufficient resolution at the species level [@Glaeser2015], and the genera *Pantoea*, *Erwinia*, *Tatumella*, and *Mixta* are often not monophyletic [@Rezzonico2016].
 Another issue is the unintentional inclusion of paralogues (originating from an intragenomic duplication event) or xenologues (arising from horizontal gene transfer) in MLSA gene analyses [@Palmer2017; @Koonin2005].
@@
 [@Brady2012] also ran analyses with the four MLSA genes as amino acid sequences.
+*Mixta* was found to be related closest to *Tatumella* by Rezzonico et al.
+and *Pantoea* by Palmer et al.
+Brady et al.
+suggested that *Mixta* was a part of the *Pantoea* genus and not separate.
+A separate study by Palmer et al.
+@Palmer2017 computed two ML trees; one used 1039 core amino acid sequences, and the other used 52 ribosomal MLSA nucleotide sequences.
+The 1039 core gene tree suggested *Pantoea* and *Tatumella* were sister taxa followed by two *Mixta* species.
+However, the ribosomal MLSA gene tree proposed *Pantoea* and *Tatumella* were sister taxa followed by *Erwinia* and then by *Mixta*.
 
-These three studies showed contentious results as *Mixta* was said to be related closest with *Tatumella* by Rezzonico et al., with *Pantoea* by Palmer et al., and within the *Pantoea* clade by
+Two possible causes may be behind the contentious results in phylogenetic analyses.
+The first is the methods; this includes the software and the applied model; two different phylogenetic tree models will provide different results even if both trees are ML and use the same genes and sequence type.
+The second potential cause is the materials: the gene sequences.
+The inclusion of different genes between analyses may give different results if one or more in the group is xenologous or has a different evolutionary rate.
+Xenologous genes need to be identified and removed from the dataset before phylogenetic analysis of gene sets since horizontal gene transfer (HGT) events frequently occur within bacteria and may occur between members of the same family.
+Sequence type may also affect results due to codon redundancy and codon usage bias.
+Additionally, there is no common phylogenetic tree model between nucleotide and amino acid sequences; therefore, evolutionary history is estimated differently between sequence types.
 
-These three studies showed contentious results as
-
-These three studies show contentious results as Rezzonico et al.
-found that *Mixta* is related closest with *Tatumella*, Palmer et al.
-found that *Mixta* is related closest with *Pantoea*, and Brady et al.
-found *Mixta* within the *Pantoea* genus.
-
-In a separate study conducted by Palmer et al.
-@Palmer2017, two ML trees were computed, one with 1039 core amino acid sequence and the other with 52 ribosomal MLSA nucleotide genes.
-Both ML trees contained species from *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*; however, the analysis with the core genes placed *Pantoea* and *Tatumella* as sister taxa followed by two *Mixta* species whereas the analysis with ribosomal MLSA genes placed *Pantoea* and *Tatumella* as sister taxa followed by *Erwinia* and then by the *Mixta* species.
-
-There are two possible causes behind contentious results in phylogenetic analyses.
-One is the use of different methods, including the use of a different phylogenetic tree model from another study despite both studies conducting maximum likelihood trees and the same sequence type.
-Additionally, there is no common phylogenetic analysis model between nucleotide and amino acid sequences; therefore, evolutionary history is estimated differently.
-The other reason may be horizontal gene transfer (HGT) events within the *Erwiniaceae* or from a non-closely-related species.
-HGT events frequently occur within bacteria, and often the genes provide the bacteria with the ability to adapt to new environments and/or survive selective pressures.
-
-The objective of this paper was to determine the cause behind the contentious phylogenetic results for the *Mixta* species.
-Homologous genes were extracted from ten complete or partial genomes - two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella*, and two outgroups - that were retrieved from public databases.
-In this study, both the nucleotide and amino acid sequences for the same genes were separately analyzed.
-Model testing was done for each gene in the dataset, and phylogenetic trees were computed using the recommended model.
+This paper's objective was to determine the cause behind the *Mixta* species' contentious phylogenetic results.
+Ten complete or partial genomes – two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella* and two outgroups – were retrieved from a public database.
+Homologous genes were extracted from these genomes in both the nucleotide and amino acid sequences.
 Distance matrices were exported from the phylogenetic trees, and these were used to determine which species were most closely related to *Mixta* for each gene.
-Nucleotide and amino acid identity analyses were also conducted for each gene in order to compare the results between evolutionary history and sequence similarity.
+Nucleotide and amino acid identity analyses were also conducted for each gene to compare the results between evolutionary history and sequence similarity.
 Phylogenetic trees were then done with either one gene or multiple genes to confirm results.
@@
 ## Genomes and Gene Sets
-Publicly available full genomes were retrieved from NCBI comprising of ten type strains from *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Enterobacter*, and *Pseudomonas* (Table 1).
-*Enterobacter cloacae* is in the same order of *Enterobacterales* as *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella* whereas *Pseudomonas syringae* is in the same class of *Gammaproteobacteria*.
+Ten publicly available full or partial genomes of the species' type strains were retrieved from the National Center for Biotechnology Information (NCBI).
+Each of the four *Erwiniaceae* genera had two representatives and *Enterobacter cloacae* and *Pseudomonas syringae* are the outgroups (Table 1).
+*E.
+cloacae* belongs to the same order as *Erwiniacae* (*Enterobacterales*) and *P.
+syringae* belongs to the same class (*Gammaproteobacteria*).
 The genomes were annotated with Prokka v1.14.1 [@Seemann2014] using default parameters, and homologous genes were extracted using GET_HOMOLOGUES software package [@Contreras-Moreira2013] with the bidirectional best-hit search algorithm using default parameters.
@@
 
-In `R` [@RCore2019], two filter functions were written to pass through gene files of the nucleotide sequences.
+In `R` [@RCore2019], two filter functions were written to pass through the nucleotide sequence gene files.
 Since both nucleotide and amino acid sequences were analyzed, the filters were only applied to the nucleotide sequences.
-A list of the passed nucleotide gene files was used to isolate the corresponding amino acid gene files to be used in later analyses.
+A list of the passed nucleotide gene files was used to isolate the corresponding amino acid gene files.
+The first filter kept genes files that contained ten, single-copy sequences that were approximately the same length.
+The shortest gene length allowed was 90% the length of the longest sequence in the gene file.
+For example, if the longest sequence in a gene file was 1000 bp, then the other sequences had to be at least 900 bp.
+Consequently, 741 gene files passed through this filter.
+Excluded from the dataset were 38 gene files because at least one species had a second copy and another 175 were excluded for not meeting the length requirements.
 
-The first filter passed through gene files that contained ten, single copy sequences.
-A minimum relative length was set where all sequences in a gene file had to be at least 90% the length of the longest sequence in the same file.
-For example, if the longest sequence in a file is 1000 bp, then the rest of the sequences had to be at least 900 bp.
-Of 954 genes, 38 genes were excluded for having at least extra sequence and an additional 175 genes were excluded because at least one sequence did not meet the length requirements.
-
-However, there may be an evolutionary reason why at least one of the sequences did not meet the length requirements.
-The second filter took the 175 gene files that were denied by the first filter one at a time and split the sequences into two groups; the first group contained the gene sequences that were over the 90% length cutoff and the second contained the rest that were shorter than the cutoff length.
-The first requirement for this filter was that all sequences in the file were longer than 80% of the length of the longest genes in the file.
-The second requirement was that both representative of a genus were in the same group, either above or below the 90% cutoff.
-The two outgroup species were not required to be in the same group since they are not from the same genus.
+Since there may be an evolutionary reason why at least one of the sequences did not meet the length requires, a second filter was applied to the 175 gene files that did not meet the initial length requirements.
+This filter took each file at a time and separated the sequences into two groups.
+One group contained the sequences that were longer than the initial 90% length cutoff and the other contained the sequences between 80 and 90% the length of the longest gene in the file.
+If at least one of the sequences was below this new 80% cutoff, then the file was excluded.
+To pass through, both representatives of a genus had to be in the same group.
 An additional 58 genes were added to the working dataset, resulting in 799 genes.
@@
 The gene fasta files were read into `R` using the `readDNAStringSet()` function (package: Biostrings) [@Pages2020].
-Alignments were done using the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and setting maximum number of iterations to 100.
-The aligned functions were then converted into a readable format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
+Alignments were done using the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and setting the maximum number of iterations to 100.
+The aligned sequences were then converted into a readable format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
 
@@
 After alignments, model testing was done for each gene through `MEGA-CC` [@Kumar2012] for both nucleotide and amino acid sequences.
-In a command line terminal, `MEGA-CC` is called along with a text file listing the file pathways to all 799 genes and a MAO file detailing the analysis preferences.
+`MEGA-CC` is called in a command-line terminal along with a text file listing the file pathways to all 799 genes and a MAO file detailing the analysis preferences.
 Default parameters were used for model testing.
-This resulted in a comma-separated values (CSV) file for each gene.
-These CSV files were read into `R` and the model for phylogenetic analysis was extracted based on lowest Bayesian Information Criterion (BIC).
+The output was a comma-separated values (CSV) file for each gene.
+These CSV files were read into `R`, and the model for phylogenetic analysis was extracted based on the lowest Bayesian Information Criterion (BIC).
+
 A list of the unique models and the number of genes that required each model are given in Table 2 for nucleotide sequences and in Table 3 for amino acid sequences.

### 2021-01-26 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -21,6 +21,8 @@ count_relatives <- function(data) {
                sum(data$Closest_Relative %in% c("E_cloacae", "Enterobacter")), sum(data$Closest_Relative %in% c("P_syringae", "Pseudomonas_")))
 }
 
+Spp <- c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", 
+         "E_cloacae", "P_syringae")
 #
 # Table 1. Genomes --------------------------------------------------------------------------------------------------------------------------------------
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
@@ -45,12 +47,12 @@ kable(strain, caption = "Table 1. Genomes used in this study.")
 phylo_models_NT <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
 PM_NT <- data.frame(table(phylo_models_NT$BM))
 
-colnames(PM_NT) <- c("Model", "#")
+colnames(PM_NT) <- c("Model", "Number")
 
 PM_NT <- mutate(PM_NT,
-                `%` = round((`#` / sum(`#`)) * 100, digits = 2))
+                Percentage = round((Number / sum(Number)) * 100, digits = 2))
 
-PM_NT <- cbind(PM_NT[1:4, ], PM_NT[5:8, ], PM_NT[9:12, ])
+# PM_NT <- cbind(PM_NT[1:4, ], PM_NT[5:8, ], PM_NT[9:12, ])
 
 write.csv(PM_NT, "10_Tables_and_Figures/Table2_PmodelsNT.csv", row.names = FALSE)
 
@@ -65,14 +67,14 @@ kable(PM_NT, caption = "Table 2. The number of genes that required each phylogen
 phylo_models_AA <- read.csv("9_Results_AA/Phylo_Models_AA.csv", stringsAsFactors = FALSE)
 PM_AA <- data.frame(table(phylo_models_AA$BM))
 
-colnames(PM_AA) <- c("Model", "#")
+colnames(PM_AA) <- c("Model", "Number")
 
 PM_AA <- mutate(PM_AA,
-                `%` = round((`#` / sum(`#`)) * 100, digits = 2))
+                Percentage = round((Number / sum(Number)) * 100, digits = 2))
 
-PM_AA <- rbind(PM_AA, c("--", "--", "--"))
-
[…]

### 2021-01-26 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 Homologous genes were extracted yielding a dataset of 799 genes.
-The genes were aligned with the ClustalW algorithm, and MEGA-CC was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
+The genes were aligned with the ClustalW algorithm, and `MEGA-CC` was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
 Distance matrices were then extracted from these trees.
@@
 
-This paper's objective was to determine the cause behind the *Mixta* species' contentious phylogenetic results.
+This paper aims to determine the cause behind the *Mixta* species' contentious phylogenetic results.
 Ten complete or partial genomes – two for each of *Mixta*, *Pantoea*, *Erwinia*, and *Tatumella* and two outgroups – were retrieved from a public database.
@@
 Ten publicly available full or partial genomes of the species' type strains were retrieved from the National Center for Biotechnology Information (NCBI).
-Each of the four *Erwiniaceae* genera had two representatives and *Enterobacter cloacae* and *Pseudomonas syringae* are the outgroups (Table 1).
+Each of the four *Erwiniaceae* genera had two representatives, and *Enterobacter cloacae* and *Pseudomonas syringae* are the outgroups (Table 1).
 *E.
-cloacae* belongs to the same order as *Erwiniacae* (*Enterobacterales*) and *P.
+cloacae* belongs to the same order as *Erwiniacae* (*Enterobacterales*), and *P.
 syringae* belongs to the same class (*Gammaproteobacteria*).
+
 The genomes were annotated with Prokka v1.14.1 [@Seemann2014] using default parameters, and homologous genes were extracted using GET_HOMOLOGUES software package [@Contreras-Moreira2013] with the bidirectional best-hit search algorithm using default parameters.
@@
 These CSV files were read into `R`, and the model for phylogenetic analysis was extracted based on the lowest Bayesian Information Criterion (BIC).
-
-A list of the unique models and the number of genes that required each model are given in Table 2 for nucleotide sequences and in Table 3 for amino acid sequences.
+Twelve and twenty-three unique models were used for nucleotide (Table 2) and amino acid sequences (Table 3) were needed to create phylogenetic gene trees.
 
@@
 
-kable(PM_NT, caption = "Table 2. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for nucleotide sequences.")
+kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.")
 ```
@@
 
-kable(PM_AA, caption = "Table 3. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for amino acid sequences.")
+kable(PM_AA, caption = "Table 3. According to model testing and the lowest BIC for amino acid sequences, the number of genes required each phylogenetic tree model.")
 ```
 
-For each unique model, a text file was written listing the gene file pathways of each gene that required that model.
-In a command line terminal, `MEGA-CC` [@Kumar2012] was called with a text file listing gene file pathways for a certain model and the corresponding MAO file.
-This was done for each unique model.
-These MAO files specified the use of a maximum likelihood statistical method, 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
-If a model required a Gamma distribution, the number of discrete Gamma categories was set to 5, which is default.
+The gene file paths were grouped according to unique models.
+For each model, a text file was written containing the pathways.
+`MEGA-CC` [@Kumar2012] was called in a command-line terminal along with a file pathways text file and the corresponding MAO file detailing the unique model.
+This step was repeated for each model.
+The MAO files specified the use of a maximum likelihood statistical method, 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
+If a model required a Gamma distribution, the number of discrete Gamma categories was set to 5, which is the default.
 
 Distance matrices were extracted from each of the phylogenetic trees and saved as Excel files.
-These files were read into `R` and the genetic distances between both of the *Mixta* species and all of the sequences were extracted for further analysis in `R`.
+These files were then read into `R`.
+The genetic distances between *Mixta* and the other species were extracted for further analysis in `R`.
 Tables were created in `R Markdown` using the `kable()` function (package: knitr) [@Xie2020] and figures were created using the `ggplot()` function (package: ggplot2) [@Wickham2016].
@@
 Alignments were done in `R` using the same methods as for gene alignment of the whole gene file.
-The aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@].
-Using base `R` code, a function was created to calculate the number of identical sites between the two sequences over the total number of sites to give a percentage of similar identity.
-Values that equal or are close to 100% are more similar to the *Mixta* sequences.
+The aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@Wickham2019].
+A function was created using base `R` code to calculate the number of identical sites between the two sequences over the total number of sites to give a similar identity percentage.
+Values that equal or are close to 100% are more identical to the *Mixta* sequences.
 
 # Results and Discussion
-After the two filters, 799 of of 954 homologous genes remained.
-The genes included those that encoded rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the *M.
+After the two filters, 799 of 954 homologous genes remained.
+The genes included those encoding rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the *M.
 calida* and *M.
@@
 Genetic distance gives a numeric estimate of evolutionary change between species.
-The species that has the shortest genetic distance to a species of interest is considered to be closest relative.
+The species that has the shortest genetic distance to a species of interest is considered the closest relative.
 For the following analysis, model testing was done first to estimate the best model for each gene.
-Next, individual phylogenetic gene trees were created and it was from these that distance matrices were exported.
+Next, individual phylogenetic gene trees were created, and it was from these that distance matrices were extracted.
 *P.
@@
 These results suggest that *P.
-septica* is the closest relative to *Mixta* followed by *P.
+septica* is the closest relative to *Mixta*, followed by *P.
 agglomerans*.
-For the amino acid sequences, instead of the *Mixta* genes being more closely related to *P.
-septica*, they are more closely related to the other species in this study (Figure 1).
-For example, the percentage of genes related to *P.
-agglomerans* increased from 13.9% for both of the *Mixta* species to 22.3 and 22.4% to *M.
+In contrast, the *Mixta* genes are more closely related to the other species in the study than *P.
+septica* in the amino acid analysis (Figure 1).
+For instance, the percentage of genes related to *P.
+agglomerans* increased by 8.4 and 8.5% for *M.
 calida* and *M.
-gaviniae*, respectively.
-This may be due to a loss of phylogenetic signal when using amino acid sequences as a result of codon degeneracy and having a third of available sites.
+gaviniae*, respectively (Figure 1).
+This may be due to a loss of phylogenetic signal when using amino acid sequences caused by codon degeneracy and having a third of available sites.
 
@@
 The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using genetic distances.
-The species with the shortest genetic distance from both of the *Mixta* species is considered to be the closest relative.
+The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative.
 Blue bars are the nucleotide sequences of the genes with light blue bars representing the percentage of *M.
@@
 gaviniae* genes were most closely related to *P.
-septica* when using nucleotide sequences in comparison to only 43% when using amino acid sequences.", fig.width=6.5}
+septica* when using nucleotide sequences compared to only 43% when using amino acid sequences.", fig.width=6.5}
 number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeMixtaNTAA.csv", 
@@
 
-Unexpectedly, the number of *Mixta* genes that are more closely related increased from 2.9% to 6.0% when using the amino acid sequences instead of the nucleotide sequences (Figure 1).
-This, along with the other differences in relatedness for *Mixta* genes between nucleotide and amino acid sequences, suggests that the sequence type used in the analysis will give different results, even when using the same genes.
+Unexpectedly, the number of *Mixta* genes that were more closely related to *E.
+cloacae* increased by 3.1% when amino acid sequences were used (Figure 1).
+The differences in relatedness for *Mixta* genes between nucleotide and amino acid sequences suggests that sequence type impacts the results, even when using the same genes.
 
-If genetic distance infers the same closest relative for a *Mixta* gene for both nucleotide and amino acid analyses, then these results may be more reliable.
+If genetic distance infers the same closest relative for a *Mixta* gene for both nucleotide and amino acid analyses, these results may be more reliable.
 Thus, for genes where the closest relative is an *Erwinia*, *Tatumella*, or *Enterobacter* species both nucleotide and amino acid analyses, they may have resulted from a horizontal gene transfer (HGT) event rather than vertical evolution from a *Pantoea* species.
-No *Mixta* genes were found to be closely related to *P.
+No *Mixta* genes were closely related to *P.
 syringae* for either the nucleotide or amino acid sequences (Figure 1).
-This confirms that *P.
-syringae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to the order *Enterobacterales*.
+This result confirms that *P.
+syringae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to  *Enterobacterales*.
 
-### Best Model vs. Best Available Model
-In MEGAX [@Kumar2018], the user is restricted to what models can be used when estimating genetic distance; however, the user is not restricted when estimating phylogenetic trees.
-Model testing in MEGAX tests all of the models that are available for phylogenetic tree analysis.
-Therefore, the recommended model based on lowest BIC value may not be available for genetic distance analysis.
+### Best Model vs Best Available Model
+In `MEGAX` [@Kumar2018], there are fewer model options for genetic distance than for phylogenetic trees.
+However, model testing using this software tests all of the available models for phylogenetic tree analysis.
+Therefore, the recommended model based on the lowest BIC value may not be available for genetic distance analysis.
 GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models, as well as the invariant sites (+I) parameter, are not available for genetic distance analysis.
-In order to determine if the limited options would affect the percentage of each species being the closest relative to the *Mixta* genes, genetic distances were estimated for each gene for the nucleotide sequences in MEGA-CC.
-The model testing results were filtered to remove the models that are not available for genetic distances in MEGAX and then distance matrices were estimated for each gene using the best available model based on lowest BIC value.
+Genetic distances for the nucleotide sequences were estimated in `MEGA-CC` using the limited model options to determine if this would affect closest relative percentages.
+The model testing results were filtered to remove the models not available for genetic distances in `MEGAX`.
+Then distance matrices were estimated for each gene using the best available model based on the lowest BIC value.
 As a result, only four of twelve models were available with more than 87% of the genes now using either the Tamura 3-parameter with a Gamma parameter (T92_G) or the Tamura-Nei with a Gamma parameter (TN93_G) (Table 4).
-This is in comparison to the original 58% of genes that required these models when not restricted by availability.
-Additionally, the best model was not available for approximately 43% of the genes (242 genes).
-The mean BIC difference between the best model and the best available model was 2.65 with a range of 0 to 122.6 where 0 means that the best available model is the same as the best model.
-When ignoring cases where the best available model was also the best recommended model, the mean BIC difference was 13.8 with a range of 0.1 to 122.6.
-These difference may result in a less accurate estimation of genetic distance between species.
+When not restricted by availability, only 58% of the genes required these models (Table 4).
+Furthermore, the best model was not available for approximately 30.3% of the genes (242 genes).
+The mean BIC difference between the best and best available models  was 2.65 (range: 0, 122.6).
+In this instance, 0 means that the best available model is the same as the best model.
+When these zeroes are ignored, the mean BIC difference was 13.8 (range: 0.1, 122.6).
+These differences may result in a less accurate estimation of genetic distance between species.
 
@@
 
-kable(models, caption = "Table 4. The number of genes requiring each model when using phylogentic tree analysis vs genetic distance analysis in MEGAX for nucleotide sequences.")
+colnames(models) <- c("Model", "Number of genes", "Number of genes, if available")
+
+kable(models, caption = "Table 4. The number of genes requiring each model when using phylogenetic tree analysis vs genetic distance analysis in MEGAX for nucleotide sequences.")
 ```
@@
 gaviniae*, respectively, when the best available models were used (Figure 2).
-The next largest difference was for genes that were calculated to be most closely related to *P.
+The next most considerable difference was for genes calculated to be most closely related to *P.
 agglomerans* (Figure 2).
-Approximately 5.5 and 7.1% less genes were most closely related to *P.
+Approximately 5.5 and 7.1% fewer genes were most closely related to *P.
 agglomerans* when the best available model was used (Figure 2).
-Additionally, there was more disparity in the closest relatives between *> calida* and *M.
+Additionally, there was more disparity in the closest relatives between *M.
+calida* and *M.
 gaviniae* when only the best available model was used (Figure 2).
-When the best model was used, the two *Mixta* typically had the same closest relative for each gene; however, this was the not the case when model options were limited to those available.
-Due to differences in evolutionary change estimation between models and the addition or exclusion of additional parameters, genetic distances may be inaccurately calculated when the best model is not used and such results may even suggest that sibling taxa are not as closely related as originally hypothesized.
+When the best model was used, the two *Mixta* typically had the same closest relative for each gene; however, this was not the case when model options were limited to those available.
+Due to differences in evolutionary change estimation between models and the addition or exclusion of additional parameters, genetic distances may be inaccurately calculated when the best model is not used.
+Such results may even suggest that sibling taxa are not as closely related as initially hypothesized.
 
@@
 The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences.
-The species with the shortest genetic distance from both of the *Mixta* species is considered to be the closest relative.
+The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative.
 Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation.
-Nevertheless, this model may not be available in some software programs or analysis methods within a software program.
-Pink bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the recommended model.
-Orange bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the best available model for genetic distance analysis in MEGAX in which GTR and HKY models and invariant sites parameter are not offered.
+Nevertheless, this model may not be available in software programs or analysis methods within a software program.
+Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model.
+Orange bars represent the percentage of *Mixta* genes most closely related to any non-*Mixta* species when using the best available genetic distance model analysis in MEGAX.
+GTR and HKY models and invariant sites parameter were not available for this analysis.
 The lighter shades represent *M.
-calida* and the darker shades represent *M.
-gavinaie*.", fig.width=6.5}
+calida*, and the darker shades represent *M.
+gaviniae*.", fig.width=6.5}
 number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv", 
@@
 
-When the closest relative results were compared between when using the best model vs when using the best available, about 20.0 and 22.3% of *M.
+About 20.0 and 22.3% of *M.
 calida* and *M.
-gaviniae* genes were estimated to have a different closest relative.
-Thus, for 82 *M.
-calida* genes and 64 *M.
-gaviniae* genes, the same closest relative was called despite using a different model.
-This may because the best model and the best available for those genes were similar and explained approximately the same amount of variance.
+gaviniae* genes, respectively, had a different closest relative between the best and the best available models.
+Thus, of the 30.3% of genes with different models between the two analyses, 10.3% of the *M.
+calida* genes and 8.0% of the *M.
+gaviniae* genes had the same closest relative.
+These genes' results may be the same because the two models were similar and explained approximately the same amount of variance.
+Genetic distances between the species may also be sufficiently large that it does not change which species is most closely related to *Mixta*.
 
@@
 The best model explains most of the variance between sequences and gives a more accurate estimation of genetic distance.
-Using a model that does not explain most of the variance between sequences will estimate genetic distance less accurately which may result in related taxa (such as those from the same species) having a longer genetic distance between them than in actuality.
+Using a model that does not explain most of the variance will estimate genetic distance less accurately, resulting in related taxa (such as those from the same species) having a longer genetic distance between them than in actuality.
 This may also result in the opposite for taxa that are not closely related (i.e.
@@
 cloacae*).
-Even though only two of the nucleotide models and one parameter were unavailable for genetic distance analysis in MEGAX, the results for closest relative to *Mixta* genes was substantially impacted.
+Even though only two nucleotide models and one parameter were unavailable for genetic distance analysis in `MEGAX`, the results for the closest relative to *Mixta* genes were substantially impacted.
 This was particularly the case between the two *Mixta* as one would expect them to have the same closest relative for most of the homologous genes.
@@
 
-Even though genetic distance estimation in MEGAX did not offer all of the models and parameters as suggested by model testing, all of the models and parameters were available for phylogenetic tree analysis and genetic distance can be exported from these trees.
-However, this has to be done manually by opening the tree file in MEGAX and exporting the distance matrix as either an Excel or CSV file and is therefore not feasible when hundreds or thousands of genetic distance matrices are needed.
+Although genetic distance estimation in `MEGAX` did not offer all of the models and parameters as suggested by model testing, all of the models and parameters were available for phylogenetic tree analysis and genetic distance can be exported from these trees.
+However, this has to be done manually by opening the tree file in `MEGAX` and exporting the distance matrix as either an Excel or CSV file.
+Therefore, it is not feasible when hundreds or thousands of genetic distance matrices are needed.
 
 ### Spatial Patterns of Closest Relatives
-To assess whether there were any positional effects of the genes and spatial patterns of the closest relatives, circular plots were generated using the *Mixta* gene IDs for nucleotide sequences (Figure 3) and amino acid sequences (Figure 4).
+Circular plots were generated using the *Mixta* gene IDs for nucleotide (Figure 3) and amino acid sequences (Figure 4).
+These were created to assess whether there were any positional effects of the genes and spatial patterns in the closest relative.
 The gene IDs give the approximate location of each gene relative to the *Mixta* genome.
-The length and colour of a bar represents the species that was the closest relative to the *Mixta* gene at that position.
+The length and colour of a bar represent the species that was the closest relative to the *Mixta* gene at that position.
 Since the closest relatives were the same both *Mixta* species except for 1 and 2 genes for nucleotide and amino acid sequences, respectively, only the results for *M.
@@
 These genes will be discussed later on.
-
-For the nucleotide sequences, there were no discernible patterns in closest relatives to *M.
-calida* genes with the exception of the seven genes that are most closely related to a *Tatumella* species (pink) which are only in the first half of the *M.
-calida* genome (Figure 3).
-Genes that are most closely related to *Pantoea* species (blue) were distributed throughout the genome (Figure 3), which further suggests that *Mixta* and *Pantoea* have a recent ancestor.
-Unexpectedly, genes that are most closely related to other genera, particularly to *E.
-cloacae* (orange) are also distributed throughout the genome (Figure 3).
 
@@
 calida* genome has 4084 genes.
-No distinct patterns emerged; however, genes that are most closely related to *P.
-septca* are the most common and occur throughout the entire genome.
+No distinct patterns emerged; however, genes most closely related to *P.
+septica* are the most common and occur throughout the genome.
 Therefore, *P.
-septica* and *Mixta* are likely to have a common recent ancestor.", fig.width=6.5, fig.height=5}
+septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5}
 FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
@@
                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringae")) +
+                               "P. syringae"),
+                    na.value = "#ff7f00") +
   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
@@
 
-Less genes are closely related to *P.
-septica* when analyzing the amino acid sequences of the *M.
-calida* genes and more are closely related to *P.
+For the nucleotide sequences, there were no discernible patterns in closest relatives to *M.
+calida* genes except for the seven genes most closely related to a *Tatumella* species (pinks) which are only in the first half of the *M.
+calida* genome (Figure 3).
+Genes that are most closely related to *Pantoea* species (blue) were distributed throughout the genome (Figure 3), which further suggests that *Mixta* and *Pantoea* have a recent ancestor.
+Unexpectedly, genes closely related to other genera, particularly to *E.
+cloacae* (orange) are also distributed throughout the genome (Figure 3).
+
+Fewer genes were closely related to *P.
+septica* when analyzing the amino acid sequences (Figure 4).
+Instead, more genes were closely related to *P.
 agglomerans*, the *Erwinia* species, the *Tatumella* species, and *E.
 cloacae* (Figure 4).
-Nevertheless, no obvious patterns arose in closest relatives to the *Mixta* genes (Figure 4).
-The *Mixta* genes that have changed their allegiance between the nucleotide and amino acid analyses may be a result of redundancy of the genetic code.
+Nevertheless, no apparent patterns arose in the closest relatives to the *Mixta* genes (Figure 4).
+The *Mixta* genes that have changed their allegiance between the nucleotide and amino acid analyses may result from the redundancy of the genetic code.
 
@@
 calida* genome has 4084 genes.
-No distinct patterns emerged; however, genes that are most closely related to *P.
-septca* are the most common and occur throughout the entire genome.
+No distinct patterns emerged; however, genes most closely related to *P.
+septica* are the most common and occur throughout the genome.
 Therefore, *P.
-septica* and *Mixta* are likely to have a common recent ancestor.", fig.width=6.5, fig.height=5}
+septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5}
 FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
@@
 There are several hypotheses as to why.
-One is that it is possible that the *Mixta* genome underwent many HGT events after speciation from *Pantoea*; however, further research into the origins of these genes and the frequency of HGT events with *Mixta* and other genera within the *Erwiniaceae* is needed.
-Another hypothesis is that the gene sequence is so well-conserved that the genetic distance between *Mixta* and more than one of the other species are near identical.
-Since only one species was extracted for the closest relative to each *Mixta* gene, it is possible that a species was called at random if there was tie between more than one species for having the shortest distance to *Mixta*.
-For example, if the genetic distance from *M.
+The first is that may have *Mixta* genome underwent many HGT events after speciation from *Pantoea*; however, further research into the origins of these genes and the frequency of HGT events with *Mixta* and other genera within the *Erwiniaceae* is needed.
+A second hypothesis is that the gene sequence is so well-conserved that the genetic distance between *Mixta* and more than one of the other species are near identical.
+Since only one species was extracted for the closest relative to each *Mixta* gene, a species may have been at random if there was a tie between more than one species for having the shortest distance to *Mixta*.
+For example, suppose the genetic distance from *M.
 calida* to both *P.
 septica* and *E.
-amylovora* for a single gene was 0.012 and this was the shortest distance for that gene, then both species are tied for closest relative for that gene.
-However, since only one can be determined as closest relative, then closest relative may have been decided by the computer program based on another characteristic such as alphabetical order.
+amylovora* for a single gene was 0.012, and this was the shortest distance for that gene.
+In that case, both species are tied for the closest relative for that gene.
+However, since only one can be determined as the closest relative, the closest relative may have been decided by the computer program based on another characteristic such as alphabetical order.
 Therefore, quantitative analysis or computing a phylogenetic tree to view the two-dimensional results is needed.
@@
 If two sequences have a site identity of 100%, then they are identical.
-If site identity is 99%, then there is up to a few bases or amino acids that are different between the two sequences depending on the length of the sequences.
+If site identity is 99%, then there is up to a few bases or amino acids that are different between the two sequences depending on the sequences' length.
 Within each gene file, the ten sequences were aligned to both *M.
 calida* and *M.
-gaviniae* in turn and the site identity was calculated.
-Site identity was divided by the length of aligned genes to give a percent identity that could be compared across the genes and between the nucleotide and amino acid sequences.
+gaviniae* in turn, and the site identity was calculated.
+Site identity was divided by the length of aligned genes to give a per cent identity that could be compared across the genes and between the nucleotide and amino acid sequences.
 From these values, the closest non-*Mixta* relative was extracted for each gene.
 
-The majority of the *Mixta* sequences were most identical with *P.
+The majority of the *Mixta* sequences were similar to *P.
 septica* for nucleotide sequences (Figure 5) with 80.6 and 79.3% of *M.
@@
 gaviniae* sequences, respectively.
-This is an increase of approximately 15% from the genetic distance results for the closest relative (Figure 1).
+These results are approximately 15% more genes than in the genetic distance analysis.
 The rest of the genes are most similar to *P.
@@
 cloacae*, and then the *Tatumella* species.
-Since mutations that do not disrupt the bacterium's ability to thrive and reproduce do not occur frequently and yet may occur at constant rate, then these results suggest that *P.
-septica* has the most recent common ancestor with the *Mixta* species.
+Viable mutations that do not disrupt the ability to survive and reproduce are rare and occur at a constant rate.
+Thus, these results suggest that *P.
+septica* and the *Mixta* species share a recent common ancestor.
 

### 2021-01-27 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 ## Distance Matrices
-After alignments, model testing was done for each gene through `MEGA-CC` [@Kumar2012] for both nucleotide and amino acid sequences.
-`MEGA-CC` is called in a command-line terminal along with a text file listing the file pathways to all 799 genes and a MAO file detailing the analysis preferences.
+After alignments, model testing was done for each gene through `MEGA-CC` [@Kumar2012] for nucleotide and amino acid sequences.
+`MEGA-CC` is called in a command-line terminal along with a text file listing the file pathways to the 799 genes and a MAO file detailing the analysis preferences.
 Default parameters were used for model testing.
@@
 gaviniae*, respectively (Figure 1).
-This may be due to a loss of phylogenetic signal when using amino acid sequences caused by codon degeneracy and having a third of available sites.
+This difference may be due to a loss of phylogenetic signal when using amino acid sequences caused by codon degeneracy and having a third of available sites.
 
@@
 Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model.
-Orange bars represent the percentage of *Mixta* genes most closely related to any non-*Mixta* species when using the best available genetic distance model analysis in MEGAX.
+Orange bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the best available genetic distance model analysis in MEGAX.
 GTR and HKY models and invariant sites parameter were not available for this analysis.
@@
 
-Although genetic distance estimation in `MEGAX` did not offer all of the models and parameters as suggested by model testing, all of the models and parameters were available for phylogenetic tree analysis and genetic distance can be exported from these trees.
+Although genetic distance estimation in `MEGAX` did not offer all of the models and parameters suggested by model testing, all of the models and parameters were available for phylogenetic tree analysis and genetic distance can be exported from these trees.
 However, this has to be done manually by opening the tree file in `MEGAX` and exporting the distance matrix as either an Excel or CSV file.
@@
 Nevertheless, no apparent patterns arose in the closest relatives to the *Mixta* genes (Figure 4).
-The *Mixta* genes that have changed their allegiance between the nucleotide and amino acid analyses may result from the redundancy of the genetic code.
+The *Mixta* genes that have changed their allegiance between the nucleotide and amino acid analyses may result from the genetic code's redundancy.
 
@@
 The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using percent site identities.
-The species with the highest percent identity to each of the *Mixta* species is considered to be most similar taxon for a single gene.
-Blue bars are the results for the nucleotide sequences where the light blue bars represent *M.
-calida* sequences and the dark blue bars represent *M.
+The species with the highest percent identity to each of the *Mixta* species is considered the most similar taxon for a single gene.
+Blue bars represent the nucleotide sequences' results; the light blue bars represent *M.
+calida* sequences, and the dark blue bars represent *M.
 gaviniae* sequences.
-Green bars are the results for the amino acid sequences where the light green bars represent *M.
-calida* sequences and the dark green bars represent *M.
+Green bars represent the amino acid sequences' results; the light green bars represent *M.
+calida* sequences, and the dark green bars represent *M.
 gaviniae* sequences.", fig.width=6.5}
@@
 gaviniae*, respectively (Figure 5).
-In the identity analyses, amino acids that are functionally equivalent were not considered.
+In the identity analyses, functionally equivalent amino acids were not considered.
 This may explain the large differences between the nucleotide and amino acid identities for which species was most similar to the *Mixta* sequences.
-Additionally, at every site for a nucleotide sequence, each base has a 1:3 ratio of being identical to the base at the corresponding location of the other sequence whereas there is only a 1:19 ratio of an amino acid being identical to the amino acid at the same position of the other sequence.
-Thus, it is far more likely for a nucleotide at a random position to be identical the nucleotide at the same position of the other sequence.
+Furthermore, every site in a nucleotide sequence has a 1:3 chance of being identical to the site at the corresponding location of a homologous gene.
+In contrast, there is only a 1:19 chance for amino acids.
+Thus, two nucleotides at the same position on two homologues are more likely to be identical.
 This may lead to problems when only amino acid sequences are considered, especially if functional equivalence is not considered.
 
-Although the numbers of genes that are most similar to *P.
-septica* are similar between the two *Mixta* species for both nucleotide and amino acid sequences (Figure 5), there were many genes in which the same species was the not the most similar taxa to both *Mixta* species.
-There were 86 genes for the nucleotide sequences and 147 genes for the amino acid sequences.
-However, of the 86 genes for nucleotide sequences, it was the case for 52% that one of the *Mixta* sequences was most similar to one representative of a genus and the other *Mixta* sequences was most similar to the other species of the same genus.
-This was also the case for 62% of the 147 amino acid genes.
+The two *Mixta* species have a comparable number of genes that are most similar to *P.
+septica* for both nucleotide and amino acid sequences (Figure 5).
+Nevertheless, many genes in which the two *Mixta* species did not share the same closest relative.
+For the nucleotide sequences, 10.8% of the *Mixta* genes did not share the same species, and 5.6% did not share the same genus for the closest relative.
+For the amino acid sequences, 18.3% did not share the same species, and 11.4% did not share the same genus.
+
+Percent identity analyses provide only the similarity of two sequences.
+They do not estimate evolutionary history nor penalize changes in the sequences.
+However, since members of a genus should be genetically similar, the percent identity analyses lend support to two taxa's relatedness.
 
@@
 ### Spatial Patterns of Most Similar Taxa
+**HERE!!**
+
 Similar to what was done with the closest relatives in the genetic distance analysis, spatial patterns of the most similar taxa to the *M.

### 2021-02-02 Manuscript/Manuscript_Paper.Rmd
Update

@@
 title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
-author: "Kimberly Hinz"
-date: "2020/12/08"
+author: "Kimberly N. Hinz, John Stavrinides*"
 output:
   word_document:
-    reference_docx: Styles1.docx
+    reference_docx: Style_RMarkdown.docx
     fig_caption: true
@@
 ```
+
+Department of Biology
+University of Regina
+3737 Wascana Parkway
+Regina, Saskatchewan, Canada
+S4S0A2
+
+**Correspondence:*
+John Stavrinides
+Department of Biology
+University of Regina
+3737 Wascana Parkway
+Regina, Saskatchewan
+S4S0A2
+Ph: (306) 337-8478
+Email: john.stavrinides@uregina.ca
+
+*Keywords*:
+
 \newpage
@@
 
-Furthermore, percent identity analyses give only how similar are two sequences.
-They do not use evolutionary models or penalize transitions and transversions in nucleotide sequences.
-However, because mutations that do not disrupt any vital processes are rare, percent identity analyses may lend support to relatedness between two taxa.
-
 ### Spatial Patterns of Most Similar Taxa
-**HERE!!**
-
-Similar to what was done with the closest relatives in the genetic distance analysis, spatial patterns of the most similar taxa to the *M.
-calida* sequences based on site identity were investigated for the nucleotide sequences (Figure 6) and the amino acid sequences (Figure 7).
-Once again, no discernible patterns arose when spatial position on the *M.
-calida* genome are considered when either the nucleotide or amino acid sequences are analyzed (Figure 6 and 7).
-However, majority of the nucleotide genes that were most similar to the two *Tatumella* species and to *E.
-amylovora* occur in the first half of the genome (Figure 6).
+Similar to the genetic distance analysis, spatial patterns in percent identity in the *M.
+calida* sequences were investigated in the nucleotide (Figure 6) and amino acid sequences (Figure 7).
+While no discernible patterns were seen in either figure, most genes most similar to the *Tatumella* species occur in the first half of the genome (Figures 6 and 7).
+Additionally, the majority of the genes most similar to *E.
+amylovora* also occur in the first half of the genome for the nucleotide sequences (Figure 6).
+The rest of the taxa occur uniformly throughout the *M.
+calida* genome for both nucleotide and amino acid sequences (Figures 6 and 7).
 
@@
 The most similar taxa to *M.
-calida* genes for nucleotide sequences around the *M.
-calida* genome.
-The most similar taxa is defined to be that which sequence is most identical to the *M.
+calida* genes for nucleotide sequences around its genome.
+The most similar taxon is that which sequence is most identical to the *M.
 calida* sequence.
@@
 The most similar taxa to *M.
-calida* genes for amino acid sequences around the *M.
-calida* genome.
-The most similar taxa is defined to be that which sequence is most identical to the *M.
+calida* genes for amino acid sequences around its genome.
+The most similar taxon is that which sequence is most identical to the *M.
 calida* sequence.
@@
 
-While no distinct patterns emerged, genes that are most similar to *P.
-septica* predominates the whole of the *M.
+While no distinct patterns emerged, genes most similar to *P.
+septica* predominate the *M.
 calida* genome (Figures 6 and 7).
@@
 ## A Comparison Between Evolutionary Change and Site Identity (Keep?)
-### *Mixta calida*
-Since all four of the analyses gave contentious results, it was necessary to compare the results for the first relative from genetic distance and the most similar taxa from site identity for both the nucleotide and amino acid sequences.
-For approximately 39% of the *M.
+### *M. calida*
+Since all four analyses gave contentious results, it was necessary to compare the first relative results from the genetic distance and the most similar taxa from site identity for both the nucleotide and amino acid sequences (Table 5).
+
+**HERE!!**
+
+For approximately 38.7% of the *M.
 calida* genes, all four analyses gave the same species.
@@
 Thus, because all four of the analyses agreed on a single species or genus, these results are more likely to be correct.
+
+```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
+stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE)
+
+same_spp <- data.frame(table(stats_on_all$Same_Species))
+same_spp <- same_spp[which(same_spp$Var1 != "Nope"), ]
+
+same_spp <- rbind(same_spp, data.frame(Var1 = "T_ptyseos", Freq = 0))
+
+colnames(same_spp) <- c("Species", "Number of Genes")
+
+same_spp <- mutate(same_spp,
+                   Species = case_when(Species == "E_amylovora" ~ "E. amylovora",
+                                       Species == "E_tasmaniensis" ~ "E. tasmaniensis",
+                                       Species == "P_agglomerans" ~ "P. agglomerans",
+                                       Species == "P_septica" ~ "P. septica",
+                                       Species == "T_ptyseos" ~ "T. ptyseos",
+                                       Species == "T_saanichensis" ~ "T. saanichensis",
+                                       Species == "E_cloacae" ~ "E. cloacae")) %>%
+  mutate(Species = factor(Species, levels = c("P. agglomerans", "P. septica", "E. amylovora", "E. tasmaniensis",
+                                              "T. ptyseos", "T. saanichensis", "E. cloacae"), ordered = TRUE))
+
+same_spp <- arrange(same_spp, Species)
+
+kable(same_spp, caption = "Table 5. The number of genes that had each species as the closest or most similar relative for all four analyses.")
+
+# table(stats_on_all$Same_Genus)
+# table(stats_on_all$NT_Same)
+# table(stats_on_all$AA_Same)
+# table(stats_on_all$Rel_Same)
+# table(stats_on_all$Ide_Same)
+```
 
@@
 <div id="refs"></div>
-
-\newpage
-# Appendix

### 2021-02-02 6_Results_Analysis.R

@@ -592,6 +592,7 @@ for(row in 1:nrow(uniq_genes)) {
   
   stats_on_all <- rbind(stats_on_all, stats)
 }
+rm(one, stats, uniq_genes, row, Spp)
 
 write.csv(stats_on_all, "10_Tables_and_Figures/FigureO_AllStatsCalida.csv", row.names = FALSE)
 

### 2021-04-17 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -30,7 +30,7 @@ strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea aggl
                                  "Enterobacter cloacae subsp cloacae", "Pseudomonas syringae pv syringae"),
                      Strain = c("DSM_22759", "DSM_22758", "NBRC_102470", "LMG_5345", "CFBP_1232", "ET1/99", "NCTC_11468",
                                 "NML_06-3099", "ATCC_13047", "ICMP_3023"),
-                     `GenBank Accession No.` = c("GCA_002953215.1", "GCA_002953195.1", "GCA_001598475.1", "GCA_002095575.1", "GCA_000367625.2",
+                     GenBank = c("GCA_002953215.1", "GCA_002953195.1", "GCA_001598475.1", "GCA_002095575.1", "GCA_000367625.2",
                                                  "GCA_000026185.1", "GCA_900478715.1", "GCA_000439375.1", "GCA_000025565.1", "GCA_001401075.1"),
                      Level = c("complete genome", "complete genome", "contigs", "contigs", "contigs", "complete genome", "complete genome", 
                                "contigs", "complete genome", "scaffold"))
@@ -84,7 +84,59 @@ PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Fi
 kable(PM_AA, caption = "Table 3. The number of genes that required each phylogenetic tree model according to model testing and the lowest BIC for amino 
       acid sequences.")
 # ```
-# Figure 1. First Relatives Percentage ------------------------------------------------------------------------------------------------------------------
+# Figure 1. Nucleotide Models in MEGAX Genetic Distance --------------------------------------------------------------------------------------------------
+DM_models_NT <- read.csv("9_Results_NT/DM_Models_NT.csv", stringsAsFactors = FALSE)
+phylo_models_NT <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
+
+models <- data.frame(Model = c("GTR_G", "GTR_G_I", "HKY_G", "HKY_G_I", "K2", "K2_G", "K2_G_I", "K2_I", "T92_G", "T92_G_I", "TN93_G", "TN93_G_I")) %>%
[…]

### 2021-04-17 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
-author: "Kimberly N. Hinz, John Stavrinides*"
 output:
   word_document:
-    reference_docx: Style_RMarkdown.docx
+    reference_docx: Style_Manuscript.docx
     fig_caption: true
@@
 
+Kimberly N. Hinz, John Stavrinides^1^
+
 Department of Biology  
@@
 
-**Correspondence:*
+^1^**Correspondence**:  
 John Stavrinides  
@@
 
-*Keywords*:
+**Keywords:**
 
-\newpage
+<!-- \newpage -->
 # Abstract
-*Mixta* is a recently described genus within the bacterial family *Erwiniaceae* and is closely related to *Pantoea*, *Erwinia*, and *Tatumella*.
+<!-- *Mixta* is a recently described genus within the bacterial family *Erwiniaceae* and is closely related to *Pantoea*, *Erwinia*, and *Tatumella*.
 Depending on the genes and method used for phylogenetic analysis, the genus with which *Mixta* shares the most recent ancestor differs.
@@
 Homologous genes were extracted yielding a dataset of 799 genes.
-The genes were aligned with the ClustalW algorithm, and `MEGA-CC` was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
+The genes were aligned with the ClustalW algorithm, and MEGA-CC was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
 Distance matrices were then extracted from these trees.
@@
 Furthermore, future research should consider that different phylogenetic analyses may provide different results.
+-->
 
-\newpage
-# Introduction
-The bacterial family Erwiniaceae includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms, including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, *Kalamiella* (LPSN, http://www.bacterio.net/).
-Numerous studies have focused on three of these genera – *Mixta*, *Pantoea*, and *Erwinia* – given their relevance to plant and human pathogens.
-For example, the genus *Erwinia* contains plant pathogens (e.g.
-*E.
-amylovora*) [@HaubenSwings2005; @Lopez2011; @Matsuura2012], epiphytes (e.g.
-*E.
-tasmaniensis*) [@Geider2006; @Mergaert1999], species that may contribute to plant disease [@Buonaurio2015], and others that are insect-associated [@Campillo2015; @Harada1997; @Skrodenyte-Arbaciauskiene2012].
-Similarly, *Pantoea*, a genus reported by Gavini et al.[@Gavini1989], includes species that inhabit a variety of environments [@Gavini1989; @Popp2010; @Liu2013], including human and plant pathogens [@Prakash2015; @deBaere2004; @Fullerton2007; @Kratz2003; @Lim2006; @Schmid2003; @vanRostenberghe2006] as well as species associated with plant growth [@Smits2011; @Kim2012], insects [@Palmer2016], and fungi [@Ma2016; @Rong2016].
-*Mixta*, originally *Pantoea*, is a novel genus first suggested by Palmer et al.
-[@Palmer2017] and later proposed by the same researchers in 2018 [@Palmer2018].
+# Intro and MM
+<!-- \newpage -->
+<!-- # Introduction -->
+<!-- The bacterial family *Erwiniaceae* includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms, including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, and *Kalamiella* [REF CRAIG].
+*Mixta* is a relatively newer genus first identified by Palmer et al.
+in 2017 [@Palmer2017] and later described by the same researchers in 2018 [@Palmer2018].
+Currently, there are only five *Mixta* species validly published [@LPSN].
 *M.
@@
 calida* were linked with post-surgical meningitis in humans.
-Currently, there are only five *Mixta* species validly published (LPSN, http://www.bacterio.net/).
+The other *Mixta* species include *M.
+theicola* which was isolated from black tea extract [@Kato2015], *M.
+intestinalis* which was isolated from a healthy human fecal sample [@Prakash2015], and *M.
+tenebrionis* which was isolated from the gut microbiota of a plastic-eating mealworm [@Xia2020].
+The first two were previously named as *Pantoea* species.
+-->
 
-Understanding the phylogeny of bacterial species and genes allows researchers to investigate the species' evolutionary histories, relatedness, and the individual genes that enable them to survive under various conditions [@Palmer2017; @Bennett2012; @Prasanna2013; @Angus2014; @deMaayer2014; @Fouts2016].
-The most common methods in phylogenetic analysis are those that use ubiquitous genes including 16S ribosomal RNA (rRNA) sequences [@Konstantinidis2007], multilocus sequence analysis (MLSA) with several housekeeping genes [@Konstantinidis2007; @Gevers2005; @Glaeser2015], and ribosomal MLSA with 53 structural ribosomal proteins [@Palmer2017; @Bennett2012; @Jolley2012].
-However, the phylogenetic trees generated from these methods may not be robust [@Konstantinidis2007; @Glaeser2015; @Brady2008] because of the lower phylogenetic signal from the highly conserved genes [@Konstantinidis2007; @Gevers2005; @Fox1992; @Konstantinidis2005; @Staley2006; @RichterRosselloMora2009].
-This issue is particularly the case with *Erwiniaceae* because much of our understanding is based on the 16S rRNA and other housekeeping genes [@Francino2006; @Hauben1998; @Naum2008; @Sproer1999; @Adeolu2016].
-However, such analyses have an insufficient resolution at the species level [@Glaeser2015], and the genera *Pantoea*, *Erwinia*, *Tatumella*, and *Mixta* are often not monophyletic [@Rezzonico2016].
-Another issue is the unintentional inclusion of paralogues (originating from an intragenomic duplication event) or xenologues (arising from horizontal gene transfer) in MLSA gene analyses [@Palmer2017; @Koonin2005].
-
-The DNA gyrase subunit (gyrB), DNA-directed RNA polymerase subunit (rpoB), ATP synthase subunit (atpD), and initiation translation factor (infB) are more reliable than 16S rRNA gene sequences for determining intra- and intergeneric relationships [@Hedegaard1999; @Dauga2002; @Mollet1997; @Paradis2005] and a robust method for the Erwiniaceae family [@Brady2010].
-The genera within Erwiniaceae are monophyletic when the analysis uses concatenated partial gyrB, rpoB, atpD, and infB gene sequences [@Brady2008; @Brady2009; @Moretti2011].
-
-Nevertheless, despite efforts to find a suitable method for identifying the species tree for Erwiniaceae genera, *Mixta*’s position relative to *Pantoea*, *Erwinia*, and *Tatumella* vastly differs between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
-Not only do different gene combinations provide different results, but different algorithms and models for alignments and phylogenetic trees also contribute to the contentious results [@Naum2008; @Octavia2014].
-For example, Prakash et al.
+<!-- *Mixta*'s position relative to other members of the *Erwiniaceae*, in particular to *Pantoea*, *Erwinia*, *Tatumella*, differs vastly between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
+Prakash et al.
 [@Prakash2015], Palmer et al.
@@
 [@Rezzonico2016], and Gueule et al.
-[@Gueule2015] all used gyrB, rpoB, atpD, and infB nucleotide gene sequences in their phylogenetic analyses.
+[@Gueule2015] all used *gyrB*, *rpoB*, *atpD*, and *infB* nucleotide gene sequences in their phylogenetic analyses.
 Prakash et al.
-conducted a neighbour-joining statistical method while the others used a maximum-likelihood (ML) method.
-Three of the studies – Rezzonico et al., Prakash et al., and Gueule et al.
-– placed the *Mixta* species within the *Pantoea* genera.
-In contrast.
-the analysis in Palmer et al.
-identified *Mixta* outside of the other three genera.
-Palmer et al.
+[@Prakash2015] conducted a neighbour-joining statistical method while the others used a maximum-likelihood (ML) method [@Palmer2017; @Rezzonico2016; @Gueule2015].
+Three of the studies – Rezzonico et al.
+[@Rezzonico2016], Prakash et al.
+[@Prakash2015], and Gueule et al.
+[@Gueule2015] – placed the *Mixta* species within the *Pantoea* genera.
+In contrast, the analysis in Palmer et al.
+[@Palmer2017] identified *Mixta* outside of the other three genera.
+A separate study by Palmer et al.
 [@Palmer2018], Rezzonico et al.
@@
 *Mixta* was found to be related closest to *Tatumella* by Rezzonico et al.
-and *Pantoea* by Palmer et al.
+[@Rezzonico2016] and *Pantoea* by Palmer et al.
+[@Palmer2018].
 Brady et al.
-suggested that *Mixta* was a part of the *Pantoea* genus and not separate.
-A separate study by Palmer et al.
-@Palmer2017 computed two ML trees; one used 1039 core amino acid sequences, and the other used 52 ribosomal MLSA nucleotide sequences.
+[@Brady2012] suggested that *Mixta* was a part of the *Pantoea* genus and not separate.
+Additionally, Palmer et al.
+[@Palmer2017] computed two ML trees; one used 1039 core amino acid sequences, and the other used 52 ribosomal MLSA nucleotide sequences.
 The 1039 core gene tree suggested *Pantoea* and *Tatumella* were sister taxa followed by two *Mixta* species.
 However, the ribosomal MLSA gene tree proposed *Pantoea* and *Tatumella* were sister taxa followed by *Erwinia* and then by *Mixta*.
+-->
 
-Two possible causes may be behind the contentious results in phylogenetic analyses.
-The first is the methods; this includes the software and the applied model; two different phylogenetic tree models will provide different results even if both trees are ML and use the same genes and sequence type.
-The second potential cause is the materials: the gene sequences.
-The inclusion of different genes between analyses may give different results if one or more in the group is xenologous or has a different evolutionary rate.
-Xenologous genes need to be identified and removed from the dataset before phylogenetic analysis of gene sets since horizontal gene transfer (HGT) events frequently occur within bacteria and may occur between members of the same family.
-Sequence type may also affect results due to codon redundancy and codon usage bias.
-Additionally, there is no common phylogenetic tree model between nucleotide and amino acid sequences; therefore, evolutionary history is estimated differently between sequence types.
-
+<!-- What remains unclear is the relative contribution of disaparate evolutionary histories of *Mixta* genes, tree-building methods, and sequence evolution on the phylogenetic position of the genus.
 This paper aims to determine the cause behind the *Mixta* species' contentious phylogenetic results.
@@
 Nucleotide and amino acid identity analyses were also conducted for each gene to compare the results between evolutionary history and sequence similarity.
-Phylogenetic trees were then done with either one gene or multiple genes to confirm results.
+Phylogenetic trees were constructed for combinations of different gene sequences.
+-->
 
-# Materials and Methods
-## Genomes and Gene Sets
-Ten publicly available full or partial genomes of the species' type strains were retrieved from the National Center for Biotechnology Information (NCBI).
-Each of the four *Erwiniaceae* genera had two representatives, and *Enterobacter cloacae* and *Pseudomonas syringae* are the outgroups (Table 1).
-*E.
-cloacae* belongs to the same order as *Erwiniacae* (*Enterobacterales*), and *P.
-syringae* belongs to the same class (*Gammaproteobacteria*).
-
+<!-- # Materials and Methods -->
+<!-- ## Genomes and Gene Sets -->
+<!-- Ten publicly available full or partial genomes of the species' type strains were retrieved from the National Center for Biotechnology Information (NCBI).
+Two representative genomes were used for each of the four *Erwiniaceae* genera while *Enterobacter cloacae* (*Enterobacterales*) and *Pseudomonas syringae* (*Pseudomonaceae*) were used as outgroups (Table 1).
 The genomes were annotated with Prokka v1.14.1 [@Seemann2014] using default parameters, and homologous genes were extracted using GET_HOMOLOGUES software package [@Contreras-Moreira2013] with the bidirectional best-hit search algorithm using default parameters.
 GET_HOMOLOGUES extracted 954 homologous genes between the ten species.
+-->
 
-```{r Table1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
-strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table1_Genomes.csv", stringsAsFactors = FALSE)
+<!-- ```{r Table1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE} -->
+<!-- strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table1_Genomes.csv", stringsAsFactors = FALSE) -->
 
-kable(strain, caption = "Table 1. Genomes used in this study.")
-```
+<!-- colnames(strain)[3] <- "GenBank Accession No." -->
 
-In `R` [@RCore2019], two filter functions were written to pass through the nucleotide sequence gene files.
-Since both nucleotide and amino acid sequences were analyzed, the filters were only applied to the nucleotide sequences.
-A list of the passed nucleotide gene files was used to isolate the corresponding amino acid gene files.
+<!-- kable(strain, caption = "Table 1. Genomes used in this study.") -->
+<!-- ``` -->
+
+<!-- In `R` [@RCore2019], two filter functions were written for the nucleotide sequence gene files.
 The first filter kept genes files that contained ten, single-copy sequences that were approximately the same length.
 The shortest gene length allowed was 90% the length of the longest sequence in the gene file.
-For example, if the longest sequence in a gene file was 1000 bp, then the other sequences had to be at least 900 bp.
-Consequently, 741 gene files passed through this filter.
-Excluded from the dataset were 38 gene files because at least one species had a second copy and another 175 were excluded for not meeting the length requirements.
+Of the 954 homologous gene sets, 38 multi-copy genes were excluded, and another 175 were excluded for not meeting the length requirements.
+-->
 
-Since there may be an evolutionary reason why at least one of the sequences did not meet the length requires, a second filter was applied to the 175 gene files that did not meet the initial length requirements.
-This filter took each file at a time and separated the sequences into two groups.
-One group contained the sequences that were longer than the initial 90% length cutoff and the other contained the sequences between 80 and 90% the length of the longest gene in the file.
-If at least one of the sequences was below this new 80% cutoff, then the file was excluded.
-To pass through, both representatives of a genus had to be in the same group.
+<!-- A second filter was applied to these 175 genes to identify any genes wherein both representatives of each genera were approximately the same length.
+The filter split the genes into two groups: sequences that met the original 90% length cutoff and sequences that were between 80 and 90% the length of the longest sequence.
+If both representatives of a genus were in the same group, then the gene was added to the working dataset.
+If at least one of the sequences did not meet the 80% cutoff, then the gene was excluded.
 An additional 58 genes were added to the working dataset, resulting in 799 genes.
-A list of the 799 gene names was used to grab the corresponding amino acid files.
+The amino acid sequences for these genes were subsequently downloaded.
+-->
 
-## Gene Alignment
-Gene alignments were in `R` using a ClustalW algorithm.
-The gene fasta files were read into `R` using the `readDNAStringSet()` function (package: Biostrings) [@Pages2020].
-Alignments were done using the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and setting the maximum number of iterations to 100.
-The aligned sequences were then converted into a readable format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
+<!-- ## Gene Alignment -->
+<!-- Gene alignments were done in `R` using a ClustalW algorithm.
+The fasta files were read into the session using the `readDNAStringSet()` function (package: Biostrings) [@Pages2020].
+The function called to do the alignments was the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and 100 iterations.
+The aligned sequences were then converted into a comma-separated values (CSV)-compatible format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007] -->
 
-## Distance Matrices
-After alignments, model testing was done for each gene through `MEGA-CC` [@Kumar2012] for nucleotide and amino acid sequences.
-`MEGA-CC` is called in a command-line terminal along with a text file listing the file pathways to the 799 genes and a MAO file detailing the analysis preferences.
+<!-- ## Model Selection -->
+<!-- The best model of sequence evolution was identified for each nucleotide and amino acid alignment using Model Selection as implemented in MEGA-CC [@Kumar2012].
 Default parameters were used for model testing.
-The output was a comma-separated values (CSV) file for each gene.
+The output was a CSV file for each gene.
 These CSV files were read into `R`, and the model for phylogenetic analysis was extracted based on the lowest Bayesian Information Criterion (BIC).
+-->
+
+<!-- ## Phylogenetic Analysis -->
+<!-- Phylogenetic analysis was performed in MEGA-CC using maximum likelihood with 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
+If a model required a Gamma distribution, the number of discrete Gamma categories was set to 5, which is the default.
+-->
+
+<!-- Distance matrices were extracted from each of the phylogenetic trees and saved as Excel files.
+These files were then read into the `R` session.
+The genetic distances between *Mixta* and the other species were extracted for further analysis.
+Tables were created in `R Markdown` using the `kable()` function (package: knitr) [@Xie2020] and figures were created using the `ggplot()` function (package: ggplot2) [@Wickham2016].
+-->
+
+<!-- ## Nucleotide and Amino Acid Identity -->
+<!-- Each of the non-*Mixta* sequences were aligned to one of the corresponding *Mixta* sequences in turn.
+Alignments were performed in R as described above, and the aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@Wickham2019].
+A function was created using base `R` code to calculate the number of identical sites between each sequence and the *Mixta* homologs.
+Values that equal or are close to 100% are more identical to the *Mixta* sequences.
+-->
+
+# Results
 Twelve and twenty-three unique models were used for nucleotide (Table 2) and amino acid sequences (Table 3) were needed to create phylogenetic gene trees.
@@
 
-The gene file paths were grouped according to unique models.
-For each model, a text file was written containing the pathways.
-`MEGA-CC` [@Kumar2012] was called in a command-line terminal along with a file pathways text file and the corresponding MAO file detailing the unique model.
-This step was repeated for each model.
-The MAO files specified the use of a maximum likelihood statistical method, 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
-If a model required a Gamma distribution, the number of discrete Gamma categories was set to 5, which is the default.
-
-Distance matrices were extracted from each of the phylogenetic trees and saved as Excel files.
-These files were then read into `R`.
-The genetic distances between *Mixta* and the other species were extracted for further analysis in `R`.
-Tables were created in `R Markdown` using the `kable()` function (package: knitr) [@Xie2020] and figures were created using the `ggplot()` function (package: ggplot2) [@Wickham2016].
-
-## Nucleotide and Amino Acid Identity
-Individual species sequences of genes were aligned to both of the corresponding *Mixta* sequences in turn.
-Alignments were done in `R` using the same methods as for gene alignment of the whole gene file.
-The aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@Wickham2019].
-A function was created using base `R` code to calculate the number of identical sites between the two sequences over the total number of sites to give a similar identity percentage.
-Values that equal or are close to 100% are more identical to the *Mixta* sequences.
-
-# Results and Discussion
 After the two filters, 799 of 954 homologous genes remained.
@@
 
-## Genetic Distance
-### Closest Relative from Genetic Distances
-Genetic distance gives a numeric estimate of evolutionary change between species.
-The species that has the shortest genetic distance to a species of interest is considered the closest relative.
-For the following analysis, model testing was done first to estimate the best model for each gene.
-Next, individual phylogenetic gene trees were created, and it was from these that distance matrices were extracted.
-*P.
-septica* was the closest relative to 66.3% and 66.5% of the *M.
-calida* and *M.
-gaviniae* genes, respectively, for the nucleotide sequences (Figure 1).
-In comparison, only 42.7% and 42.8% of *M.
-calida* and *M.
-gaviniae* genes, respectively, are closely related to *P.
-septica* for the amino acid sequences (Figure 1), but this is still the majority.
-These results suggest that *P.
-septica* is the closest relative to *Mixta*, followed by *P.
-agglomerans*.
-In contrast, the *Mixta* genes are more closely related to the other species in the study than *P.
-septica* in the amino acid analysis (Figure 1).
-For instance, the percentage of genes related to *P.
-agglomerans* increased by 8.4 and 8.5% for *M.
-calida* and *M.
-gaviniae*, respectively (Figure 1).
-This difference may be due to a loss of phylogenetic signal when using amino acid sequences caused by codon degeneracy and having a third of available sites.
+## Best Model vs Best Available Model
+In MEGAX [@Kumar2018], the user can either calculate genetic distance or estimate a phylogenetic tree.
+However, although model testing accesses all possible models available in the software, all models are not available for genetic distance analysis.
+The GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models and the +I (invariant sites) parameter are available only for phylogenetic tree analysis.
 
 ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
-The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using genetic distances.
-The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative.
-Blue bars are the nucleotide sequences of the genes with light blue bars representing the percentage of *M.
-calida* genes and dark blue bars representing *M.
-gaviniae*.
-Green bars are the corresponding amino acid sequences with light green representing *M.
-calida* genes and dark blue bars representing *M.
-gaviniae*.
-Approximately 66% of *M.
-calida* and *M.
-gaviniae* genes were most closely related to *P.
-septica* when using nucleotide sequences compared to only 43% when using amino acid sequences.", fig.width=6.5}
-number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeMixtaNTAA.csv",
+The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange).
+Models are for nucleotide sequences.", fig.width=6.5}
+models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE)
+
+ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Dark2",
+                    labels = c("Phylogenetic", "Genetic Distance")) +
+  labs(y = "Percent of Genes \n Requiring Each Model",
+       fill = "Analysis") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels
+        legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (n = 799) most closely related to any of the eight non-*Mixta* species.
+Closest relative was inferred using shortest genetic distance.
+Blue bars represent the nucleotide sequences and green bars represent the amino acid sequences.
+The lighter of each colour represents *M.
+calida* and the darker *M.
+gaviniae*.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv",
                              stringsAsFactors = FALSE) %>%
@@
 ```
-
-Unexpectedly, the number of *Mixta* genes that were more closely related to *E.
-cloacae* increased by 3.1% when amino acid sequences were used (Figure 1).
-The differences in relatedness for *Mixta* genes between nucleotide and amino acid sequences suggests that sequence type impacts the results, even when using the same genes.
-
-If genetic distance infers the same closest relative for a *Mixta* gene for both nucleotide and amino acid analyses, these results may be more reliable.
-Thus, for genes where the closest relative is an *Erwinia*, *Tatumella*, or *Enterobacter* species both nucleotide and amino acid analyses, they may have resulted from a horizontal gene transfer (HGT) event rather than vertical evolution from a *Pantoea* species.
-No *Mixta* genes were closely related to *P.
-syringae* for either the nucleotide or amino acid sequences (Figure 1).
-This result confirms that *P.
-syringae* is a suitable outgroup for this analysis and suggests that any HGT events may be restricted to  *Enterobacterales*.
-
-### Best Model vs Best Available Model
-In `MEGAX` [@Kumar2018], there are fewer model options for genetic distance than for phylogenetic trees.
-However, model testing using this software tests all of the available models for phylogenetic tree analysis.
-Therefore, the recommended model based on the lowest BIC value may not be available for genetic distance analysis.
-GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models, as well as the invariant sites (+I) parameter, are not available for genetic distance analysis.
-Genetic distances for the nucleotide sequences were estimated in `MEGA-CC` using the limited model options to determine if this would affect closest relative percentages.
-The model testing results were filtered to remove the models not available for genetic distances in `MEGAX`.
-Then distance matrices were estimated for each gene using the best available model based on the lowest BIC value.
-As a result, only four of twelve models were available with more than 87% of the genes now using either the Tamura 3-parameter with a Gamma parameter (T92_G) or the Tamura-Nei with a Gamma parameter (TN93_G) (Table 4).
-When not restricted by availability, only 58% of the genes required these models (Table 4).
-Furthermore, the best model was not available for approximately 30.3% of the genes (242 genes).
-The mean BIC difference between the best and best available models  was 2.65 (range: 0, 122.6).
-In this instance, 0 means that the best available model is the same as the best model.
-When these zeroes are ignored, the mean BIC difference was 13.8 (range: 0.1, 122.6).
-These differences may result in a less accurate estimation of genetic distance between species.
-
-```{r Table4, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
-models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table4_GenDistModelsNT.csv", stringsAsFactors = FALSE)
-
-colnames(models) <- c("Model", "Number of genes", "Number of genes, if available")
-
-kable(models, caption = "Table 4. The number of genes requiring each model when using phylogenetic tree analysis vs genetic distance analysis in MEGAX for nucleotide sequences.")
-```
-
-The closest relatives to both *Mixta* species were extracted from the distance matrices.
-*P.
-septica* was estimated to be the closest relative to 12.6 and 14.4% more genes for *M.
-calida* and *M.
-gaviniae*, respectively, when the best available models were used (Figure 2).
-The next most considerable difference was for genes calculated to be most closely related to *P.
-agglomerans* (Figure 2).
-Approximately 5.5 and 7.1% fewer genes were most closely related to *P.
-agglomerans* when the best available model was used (Figure 2).
-Additionally, there was more disparity in the closest relatives between *M.
-calida* and *M.
-gaviniae* when only the best available model was used (Figure 2).
-When the best model was used, the two *Mixta* typically had the same closest relative for each gene; however, this was not the case when model options were limited to those available.
-Due to differences in evolutionary change estimation between models and the addition or exclusion of additional parameters, genetic distances may be inaccurately calculated when the best model is not used.
-Such results may even suggest that sibling taxa are not as closely related as initially hypothesized.
 
@@
 
-About 20.0 and 22.3% of *M.
-calida* and *M.
-gaviniae* genes, respectively, had a different closest relative between the best and the best available models.
-Thus, of the 30.3% of genes with different models between the two analyses, 10.3% of the *M.
-calida* genes and 8.0% of the *M.
-gaviniae* genes had the same closest relative.
-These genes' results may be the same because the two models were similar and explained approximately the same amount of variance.
-Genetic distances between the species may also be sufficiently large that it does not change which species is most closely related to *Mixta*.
-
-Each model estimates genetic distance differently.
-The best model explains most of the variance between sequences and gives a more accurate estimation of genetic distance.
-Using a model that does not explain most of the variance will estimate genetic distance less accurately, resulting in related taxa (such as those from the same species) having a longer genetic distance between them than in actuality.
-This may also result in the opposite for taxa that are not closely related (i.e.
-*Pantoea* spp.
-and *E.
-cloacae*).
-Even though only two nucleotide models and one parameter were unavailable for genetic distance analysis in `MEGAX`, the results for the closest relative to *Mixta* genes were substantially impacted.
-This was particularly the case between the two *Mixta* as one would expect them to have the same closest relative for most of the homologous genes.
-Therefore, using the best model is recommended.
-However, this may not be possible when estimating genetic distance using more than one gene that may require different models.
-Thus, it must be considered that one or more the genes in the concatenation may have a different evolutionary pattern than the others and results should be investigated further to ensure accuracy.
-
-Although genetic distance estimation in `MEGAX` did not offer all of the models and parameters suggested by model testing, all of the models and parameters were available for phylogenetic tree analysis and genetic distance can be exported from these trees.
-However, this has to be done manually by opening the tree file in `MEGAX` and exporting the distance matrix as either an Excel or CSV file.
-Therefore, it is not feasible when hundreds or thousands of genetic distance matrices are needed.
-
-### Spatial Patterns of Closest Relatives
-Circular plots were generated using the *Mixta* gene IDs for nucleotide (Figure 3) and amino acid sequences (Figure 4).
-These were created to assess whether there were any positional effects of the genes and spatial patterns in the closest relative.
-The gene IDs give the approximate location of each gene relative to the *Mixta* genome.
-The length and colour of a bar represent the species that was the closest relative to the *Mixta* gene at that position.
-Since the closest relatives were the same both *Mixta* species except for 1 and 2 genes for nucleotide and amino acid sequences, respectively, only the results for *M.
-calida* are shown.
-These genes will be discussed later on.
-
 ```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
@@
 ```
-
-For the nucleotide sequences, there were no discernible patterns in closest relatives to *M.
-calida* genes except for the seven genes most closely related to a *Tatumella* species (pinks) which are only in the first half of the *M.
-calida* genome (Figure 3).
-Genes that are most closely related to *Pantoea* species (blue) were distributed throughout the genome (Figure 3), which further suggests that *Mixta* and *Pantoea* have a recent ancestor.
-Unexpectedly, genes closely related to other genera, particularly to *E.
-cloacae* (orange) are also distributed throughout the genome (Figure 3).
-
-Fewer genes were closely related to *P.
-septica* when analyzing the amino acid sequences (Figure 4).
-Instead, more genes were closely related to *P.
-agglomerans*, the *Erwinia* species, the *Tatumella* species, and *E.
-cloacae* (Figure 4).
-Nevertheless, no apparent patterns arose in the closest relatives to the *Mixta* genes (Figure 4).
-The *Mixta* genes that have changed their allegiance between the nucleotide and amino acid analyses may result from the genetic code's redundancy.
 
@@
 
-If many genes were horizontally transferred in a single event, then large blocks of genes from a single species would be expected.
-However, this is not the case for any of species.
-There are several hypotheses as to why.
-The first is that may have *Mixta* genome underwent many HGT events after speciation from *Pantoea*; however, further research into the origins of these genes and the frequency of HGT events with *Mixta* and other genera within the *Erwiniaceae* is needed.
-A second hypothesis is that the gene sequence is so well-conserved that the genetic distance between *Mixta* and more than one of the other species are near identical.
-Since only one species was extracted for the closest relative to each *Mixta* gene, a species may have been at random if there was a tie between more than one species for having the shortest distance to *Mixta*.
-For example, suppose the genetic distance from *M.
-calida* to both *P.
-septica* and *E.
-amylovora* for a single gene was 0.012, and this was the shortest distance for that gene.
-In that case, both species are tied for the closest relative for that gene.
-However, since only one can be determined as the closest relative, the closest relative may have been decided by the computer program based on another characteristic such as alphabetical order.
-Therefore, quantitative analysis or computing a phylogenetic tree to view the two-dimensional results is needed.
-
-## Nucleotide and Amino Acid Identity
-### Most Similar Taxa from Site Identity
-Site identity is here defined as the amount of similarity between two sequences.
-If two sequences have a site identity of 100%, then they are identical.
-If site identity is 99%, then there is up to a few bases or amino acids that are different between the two sequences depending on the sequences' length.
-Within each gene file, the ten sequences were aligned to both *M.
-calida* and *M.
-gaviniae* in turn, and the site identity was calculated.
-Site identity was divided by the length of aligned genes to give a per cent identity that could be compared across the genes and between the nucleotide and amino acid sequences.
-From these values, the closest non-*Mixta* relative was extracted for each gene.
-
-The majority of the *Mixta* sequences were similar to *P.
-septica* for nucleotide sequences (Figure 5) with 80.6 and 79.3% of *M.
-calida* and *M.
-gaviniae* sequences, respectively.
-These results are approximately 15% more genes than in the genetic distance analysis.
-The rest of the genes are most similar to *P.
-agglomerans*, followed by the *Erwinia* species, *E.
-cloacae*, and then the *Tatumella* species.
-Viable mutations that do not disrupt the ability to survive and reproduce are rare and occur at a constant rate.
-Thus, these results suggest that *P.
-septica* and the *Mixta* species share a recent common ancestor.
-
 ```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5.
@@
 ```
-
-In contrast, only 47.8 and 47.6% of *M.
-calida* and *M.
-gaviniae* sequences were most similar to *P.
-septica* for the amino acid sequences (Figure 5).
-Instead, the genes are most similar to either *P.
-agglomerans* or one of the two *Erwinia* species (Figure 5).
-Additionally, 3.1 and 2.5% more genes were most similar to the *E.
-cloacae* sequences for *M.
-calida* and *M.
-gaviniae*, respectively (Figure 5).
-In the identity analyses, functionally equivalent amino acids were not considered.
-This may explain the large differences between the nucleotide and amino acid identities for which species was most similar to the *Mixta* sequences.
-Furthermore, every site in a nucleotide sequence has a 1:3 chance of being identical to the site at the corresponding location of a homologous gene.
-In contrast, there is only a 1:19 chance for amino acids.
-Thus, two nucleotides at the same position on two homologues are more likely to be identical.
-This may lead to problems when only amino acid sequences are considered, especially if functional equivalence is not considered.
-
-The two *Mixta* species have a comparable number of genes that are most similar to *P.
-septica* for both nucleotide and amino acid sequences (Figure 5).
-Nevertheless, many genes in which the two *Mixta* species did not share the same closest relative.
-For the nucleotide sequences, 10.8% of the *Mixta* genes did not share the same species, and 5.6% did not share the same genus for the closest relative.
-For the amino acid sequences, 18.3% did not share the same species, and 11.4% did not share the same genus.
-
-Percent identity analyses provide only the similarity of two sequences.
-They do not estimate evolutionary history nor penalize changes in the sequences.
-However, since members of a genus should be genetically similar, the percent identity analyses lend support to two taxa's relatedness.
-
-### Spatial Patterns of Most Similar Taxa
-Similar to the genetic distance analysis, spatial patterns in percent identity in the *M.
-calida* sequences were investigated in the nucleotide (Figure 6) and amino acid sequences (Figure 7).
-While no discernible patterns were seen in either figure, most genes most similar to the *Tatumella* species occur in the first half of the genome (Figures 6 and 7).
-Additionally, the majority of the genes most similar to *E.
-amylovora* also occur in the first half of the genome for the nucleotide sequences (Figure 6).
-The rest of the taxa occur uniformly throughout the *M.
-calida* genome for both nucleotide and amino acid sequences (Figures 6 and 7).
 
@@
 
-While no distinct patterns emerged, genes most similar to *P.
-septica* predominate the *M.
-calida* genome (Figures 6 and 7).
-This suggests that *Mixta* and *P.
-septica* may be related as their sequences are similar and few mutations have occurred after they diverged.
-
-## A Comparison Between Evolutionary Change and Site Identity (Keep?)
-### *M. calida*
-Since all four analyses gave contentious results, it was necessary to compare the first relative results from the genetic distance and the most similar taxa from site identity for both the nucleotide and amino acid sequences (Table 5).
-
-**HERE!!**
-
-For approximately 38.7% of the *M.
-calida* genes, all four analyses gave the same species.
-Of these genes, 255 agreed with *P.
-septica*, 22 with *P.
-agglomerans*, 17 with *E.
-tasmaniensis*, 8 with *E.
-amylovora*, 1 with *T.
-saanichensis*, and 6 with *E.
-cloacae*.
-Additionally, the four analyses gave the same genus approximately 63% of the genes where in 454 agreed with *Pantoea* followed by *Erwinia* with 44, and *Tatumella* with 3.
-Thus, because all four of the analyses agreed on a single species or genus, these results are more likely to be correct.
-
 ```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
@@
 ```
-
-The two analyses conducted with nucleotide sequences agreed on a species 78% of the time whereas the two analyses conducted with amino acid sequences agreed only 70% of the time.
-It is not unexpected that the two nucleotide analyses would agree on taxa more often than the amino acid analyses since there are more sites in nucleotide sequences and no degeneracy in the code.
 
@@
 ```
-Visual comparison of the results from the four analyses does not reveal any patterns (Figure 8).
-The four analyses are plotted on y axis where NT Closest is the closest relative according to genetic distance for nucleotide sequences  (Figure 8), NT Identity is the most similar taxa according to percent identity of nucleotide sequences, AA Closest is the closest relative according to genetic distance for amino acid sequences, and AA Identity is the most similar taxa according to percent identity of amino acid sequences (Figure 8).
-The first 100 genes according to *M.
-calida* gene ID are plotted on the x axis and the colour of the block represents the species that is the closest relative or most similar taxa to that gene for that analysis (Figure 8).
-
-## Phylogenetic Trees
-### Single Gene Trees
-From the genetic distance analyses, only one gene separated *M.
-calida* and *M.
-gaviniae* for the nucleotide sequence and two genes for the amino acid sequence.
-The one gene for the nucleotide sequences was ttuB.
-According to genetic distance, *M.
-calida* was more closely related to *T.
-ptyseos* and *M.
-gaviniae* was more closely related to *P.
-septica* rather than to each other.
-A phylogenetic tree of this gene confirms that *M.
-calida* is more closely related to the *Tatumella* species where *M.
-gaviniae* to *Pantoea* and *Erwinia* (Figure 9A).
-The gene ttuB also had the two *Mixta* species separated for the amino acid sequences and the resulting tree shows the same results (Figure 9B).
 
@@
 ```
-
-The second gene that had the two *Mixta* species separated for the amino acid sequences was secY.
-From the nucleotide sequences, all of the taxa except for *P.
-syringae* are very closely related and bifurcations cannot be determined from the radial tree (Figure 10A).
-The same is the case for the amino acid sequences (Figure 10B).
-Condensed tree at 70% cutoff was created to show the relationships for both the nucleotide sequences (Figure 10C) and the amino acid sequences (Figure 10D).
-In the nucleotide tree, the two *Mixta* species are paired together, yet they are placed in a group with the two outgroups and the *Erwinia* species (Figure 10C).
-A condensed tree for the amino acid sequences showed that birfucations between most of the species were not supported above the 70% cutoff.
-These gene trees suggest that the differences between the *Enterobacterales* are not great enough to accurately infer the *Mixta* relative.
 
@@
 
-# Conclusion
-Since the analyses provided different results for many of the genes, a single analysis may not be enough to conclude that a species is related to another.
+# Discussion
+
+# Acknowledgements
+
+# Contributions to the Field Statement
+
+# Figure Legends
+
+# Tables
 
 \newpage
-# Literature Cited
+# References
 <div id="refs"></div>

### 2021-04-18 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -32,8 +32,7 @@ strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea aggl
                                 "NML_06-3099", "ATCC_13047", "ICMP_3023"),
                      GenBank = c("GCA_002953215.1", "GCA_002953195.1", "GCA_001598475.1", "GCA_002095575.1", "GCA_000367625.2",
                                                  "GCA_000026185.1", "GCA_900478715.1", "GCA_000439375.1", "GCA_000025565.1", "GCA_001401075.1"),
-                     Level = c("complete genome", "complete genome", "contigs", "contigs", "contigs", "complete genome", "complete genome", 
-                               "contigs", "complete genome", "scaffold"))
+                     Level = c("Complete", "Complete", "Contigs", "Contigs", "Contigs", "Complete", "Complete", "Contigs", "Complete", "Scaffold"))
 
 write.csv(strain, "10_Tables_and_Figures/Table1_Genomes.csv", row.names = FALSE)
 
@@ -42,6 +41,39 @@ strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_F
 
 kable(strain, caption = "Table 1. Genomes used in this study.")
 # ```
+
+# Table 2. Best vs Available Models ----------------------------------------------------------------------------------------------------------------------
+DM_models_NT <- read.csv("9_Results_NT/DM_Models_NT.csv", stringsAsFactors = FALSE)
+phylo_models_NT <- read.csv("9_Results_NT/Phylo_Models_NT.csv", stringsAsFactors = FALSE)
+
+models <- data.frame(Model = c("GTR_G", "GTR_G_I", "HKY_G", "HKY_G_I", "K2", "K2_G", "K2_G_I", "K2_I", "T92_G", "T92_G_I", "TN93_G", "TN93_G_I")) %>%
+  mutate(Best = c(sum(phylo_models_NT$BM == "GTR_G"), sum(phylo_models_NT$BM == "GTR_G_I"), sum(phylo_models_NT$BM == "HKY_G"),
+                  sum(phylo_models_NT$BM == "HKY_G_I"), sum(phylo_models_NT$BM == "K2"), sum(phylo_models_NT$BM == "K2_G"),
[…]

### 2021-04-18 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 # Results
+```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/O10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE)
+
+kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.")
+
+```
+
+# Old Results
 Twelve and twenty-three unique models were used for nucleotide (Table 2) and amino acid sequences (Table 3) were needed to create phylogenetic gene trees.

### 2021-04-19 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -2,6 +2,7 @@
 
 # The following code analyzes the results from the distance matrices and creates the tables and figures for the paper.
 
+library("knitr")
 library("tidyverse")
 library("msa")
 library("ggplot2")
@@ -13,7 +14,7 @@ theme_set(theme_classic())
 
 setwd("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396")
 #
-### Functions and Frequently used Variables #############################################################################################################
+### Common Functions/Variables ############################################################################################################################
 count_relatives <- function(data) {
   num_rel <- c(sum(data$Closest_Relative %in% c("P_septica", "Pantoea_sept")), sum(data$Closest_Relative %in% c("P_agglomerans", "Pantoea_aggl")), 
                sum(data$Closest_Relative %in% c("E_tasmaniensis", "Erwinia_tasm")), sum(data$Closest_Relative %in% c("E_amylovora", "Erwinia_amyl")),
@@ -24,7 +25,7 @@ count_relatives <- function(data) {
 Spp <- c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", 
          "E_cloacae", "P_syringae")
 #
-# Table 1. Genomes --------------------------------------------------------------------------------------------------------------------------------------
+# Table 1. Genomes ----------------------------------------------------------------------------------------------------------------------------------------
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
                                  "Erwinia tasmaniensis", "Tatumella ptyseos", "Tatumella saanichensis", 
                                  "Enterobacter cloacae subsp cloacae", "Pseudomonas syringae pv syringae"),
@@ -42,7 +43,7 @@ strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_F
 kable(strain, caption = "Table 1. Genomes used in this study.")
 # ```
 
-# Table 2. Best vs Available Models ----------------------------------------------------------------------------------------------------------------------
+# Table 2. Best vs Available Models -----------------------------------------------------------------------------------------------------------------------
 DM_models_NT <- read.csv("9_Results_NT/DM_Models_NT.csv", stringsAsFactors = FALSE)
[…]

### 2021-04-19 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

--- 2021-04-17 Manuscript/Manuscript_Paper.Rmd
@@
 # Results
-Twelve and twenty-three unique models were used for nucleotide (Table 2) and amino acid sequences (Table 3) were needed to create phylogenetic gene trees.
+```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_BestAvailableModels.csv", stringsAsFactors = FALSE)
 
-```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
-PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE)
+colnames(models)[2:3] <- c("Best Model", "Best Available Model")
 
-kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.")
+kable(models, 
+      caption = "Table 2. The number of genes that required each model according to the best model with the lowest BIC and the best available model for genetic distance analysis.")
 ```
 
-```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
-PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_PmodelsAA.csv", stringsAsFactors = FALSE)
-
-kable(PM_AA, caption = "Table 3. According to model testing and the lowest BIC for amino acid sequences, the number of genes required each phylogenetic tree model.")
-```
-
-After the two filters, 799 of 954 homologous genes remained.
-The genes included those encoding rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the *M.
-calida* and *M.
-gaviniae* genomes.
-
-## Best Model vs Best Available Model
-In MEGAX [@Kumar2018], the user can either calculate genetic distance or estimate a phylogenetic tree.
-However, although model testing accesses all possible models available in the software, all models are not available for genetic distance analysis.
-The GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models and the +I (invariant sites) parameter are available only for phylogenetic tree analysis.
-
-```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
-The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange).
-Models are for nucleotide sequences.", fig.width=6.5}
-models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE)
-
-ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  scale_fill_brewer(palette = "Dark2",
-                    labels = c("Phylogenetic", "Genetic Distance")) +
-  labs(y = "Percent of Genes \n Requiring Each Model",
-       fill = "Analysis") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels
-        legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner
-        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
-        legend.text.align = 0,                                                            # Aligns legend text to the left
-        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
-
-```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2.
-The percentage of genes (n = 799) most closely related to any of the eight non-*Mixta* species.
-Closest relative was inferred using shortest genetic distance.
-Blue bars represent the nucleotide sequences and green bars represent the amino acid sequences.
-The lighter of each colour represents *M.
-calida* and the darker *M.
-gaviniae*.", fig.width=6.5}
-number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv",
-                             stringsAsFactors = FALSE) %>%
-  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis",
-                                              "E. cloacae", "P. syringae"), ordered = TRUE),
-         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
-
-ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")),
-                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")),
-                               expression(paste("AA and ", italic("M. calida"), sep = "")),
-                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
-  labs(y = "Percent of Genes",
-       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
-        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
-        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
-        legend.text.align = 0,                                                            # Aligns legend text to the left
-        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
-
-```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
-The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences.
-The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative.
-Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation.
-Nevertheless, this model may not be available in software programs or analysis methods within a software program.
-Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model.
-Orange bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the best available genetic distance model analysis in MEGAX.
-GTR and HKY models and invariant sites parameter were not available for this analysis.
+```{r Figure1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1.
+The percentage of genes (n = 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended model from model testing versus the best available) model for genetic distance analysis in MEGAX.
+The sequences were nucleotide sequences.
+The species with the shortest genetic distance from the *Mixta* species is considered to be the closest relative.
+The pink and orange bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the recommended model and the best available model, respectively.
 The lighter shades represent *M.
-calida*, and the darker shades represent *M.
-gaviniae*.", fig.width=6.5}
-number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv",
+calida* and the darker shades represent *M.
+gavinaie*.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeBestvsAvailableModelNT.csv", 
                              stringsAsFactors = FALSE) %>%
@@
   scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"),
-                    labels = c(expression(paste("Best Model and ", italic("M. calida"), sep = "")),
-                               expression(paste("Best Model and ", italic("M. gaviniae"), sep = "")),
+                    labels = c(expression(paste("Recommended Model and ", italic("M. calida"), sep = "")), 
+                               expression(paste("Recommended Model and ", italic("M. gaviniae"), sep = "")), 
                                expression(paste("Best Available Model and ", italic("M. calida"), sep = "")), 
@@
 
-```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
+<!-- # Old Results -->
+<!-- ```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE} -->
+<!-- PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE) -->
+
+<!-- kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.") -->
+<!-- ``` -->
+
+<!-- ```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE} -->
+<!-- PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_PmodelsAA.csv", stringsAsFactors = FALSE) -->
+
+<!-- kable(PM_AA, caption = "Table 3. According to model testing and the lowest BIC for amino acid sequences, the number of genes required each phylogenetic tree model.") -->
+<!-- ``` -->
+
+<!-- ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
+The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange).
+Models are for nucleotide sequences.", fig.width=6.5} -->
+<!-- models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE) -->
+
+<!-- ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   scale_fill_brewer(palette = "Dark2", -->
+<!--                     labels = c("Phylogenetic", "Genetic Distance")) + -->
+<!--   labs(y = "Percent of Genes \n Requiring Each Model", -->
+<!--        fill = "Analysis") + -->
+<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels -->
+<!--         legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner -->
+<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
+<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
+<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
+
+<!-- ```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (n = 799) most closely related to any of the eight non-*Mixta* species.
+Closest relative was inferred using shortest genetic distance.
+Blue bars represent the nucleotide sequences and green bars represent the amino acid sequences.
+The lighter of each colour represents *M.
+calida* and the darker *M.
+gaviniae*.", fig.width=6.5} -->
+<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", -->
+<!--                              stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
+<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
+<!--          Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE)) -->
+
+<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), -->
+<!--                                expression(paste("NT and ", italic("M. gaviniae"), sep = "")), -->
+<!--                                expression(paste("AA and ", italic("M. calida"), sep = "")), -->
+<!--                                expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) + -->
+<!--   labs(y = "Percent of Genes", -->
+<!--        fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) + -->
+<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
+<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
+<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
+<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
+<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
+
+<!-- ```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences.
+The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative.
+Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation.
+Nevertheless, this model may not be available in software programs or analysis methods within a software program.
+Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model.
+Orange bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the best available genetic distance model analysis in MEGAX.
+GTR and HKY models and invariant sites parameter were not available for this analysis.
+The lighter shades represent *M.
+calida*, and the darker shades represent *M.
+gaviniae*.", fig.width=6.5} -->
+<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv", -->
+<!--                              stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
+<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
+<!--          Type = factor(Type, levels = c("PD_calida", "PD_gaviniae", "GD_calida", "GD_gaviniae"), ordered = TRUE)) -->
+
+<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"), -->
+<!--                     labels = c(expression(paste("Best Model and ", italic("M. calida"), sep = "")), -->
+<!--                                expression(paste("Best Model and ", italic("M. gaviniae"), sep = "")), -->
+<!--                                expression(paste("Best Available Model and ", italic("M. calida"), sep = "")), -->
+<!--                                expression(paste("Best Available Model and ", italic("M. gaviniae"), sep = "")))) + -->
+<!--   labs(y = "Percent of Genes", -->
+<!--        fill = expression(paste("Model and ", italic("Mixta"), " species", sep = ""))) + -->
+<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
+<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
+<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
+<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
+<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
+
+<!-- ```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
 Closest relatives to *M.
@@
 Therefore, *P.
-septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5}
-FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
-  mutate(Closest_Relative = factor(Closest_Relative,
-                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
-                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
+septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5} -->
+<!-- FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
+<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
+<!--                                               "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE)) -->
 
-ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
-  coord_polar() +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringae"),
-                    na.value = "#ff7f00") +
-  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
-  labs(fill = "Closest Relative") +
-  theme_bw() +
-  theme(legend.text = element_text(face = "italic"),
-        axis.title = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
+<!-- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
+<!--   coord_polar() + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
+<!--                                "P. syringae"), -->
+<!--                     na.value = "#ff7f00") + -->
+<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
+<!--   labs(fill = "Closest Relative") + -->
+<!--   theme_bw() + -->
+<!--   theme(legend.text = element_text(face = "italic"), -->
+<!--         axis.title = element_blank(), -->
+<!--         axis.text.y = element_blank(), -->
+<!--         axis.ticks.y = element_blank(), -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
 
-```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.
+<!-- ```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.
 Closest relatives to *M.
@@
 Therefore, *P.
-septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5}
-FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
-  mutate(Closest_Relative = factor(Closest_Relative,
-                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
-                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
+septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5} -->
+<!-- FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
+<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
+<!--                                               "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE)) -->
 
-ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
-  coord_polar() +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringae"),
-                    na.value = "#ff7f00") +
-  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
-  labs(fill = "Closest Relative") +
-  theme_bw() +
-  theme(legend.text = element_text(face = "italic"),
-        axis.title = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
+<!-- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
+<!--   coord_polar() + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
+<!--                                "P. syringae"), -->
+<!--                     na.value = "#ff7f00") + -->
+<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
+<!--   labs(fill = "Closest Relative") + -->
+<!--   theme_bw() + -->
+<!--   theme(legend.text = element_text(face = "italic"), -->
+<!--         axis.title = element_blank(), -->
+<!--         axis.text.y = element_blank(), -->
+<!--         axis.ticks.y = element_blank(), -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
 
-```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5.
+<!-- ```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5.
 The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using percent site identities.
@@
 calida* sequences, and the dark green bars represent *M.
-gaviniae* sequences.", fig.width=6.5}
-number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5_FirstRelativeMixtaNTAAIdentity.csv",
-                             stringsAsFactors = FALSE) %>%
-  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis",
-                                              "E. cloacae", "P. syringae"), ordered = TRUE),
-         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
+gaviniae* sequences.", fig.width=6.5} -->
+<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5_FirstRelativeMixtaNTAAIdentity.csv", -->
+<!--                              stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
+<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
+<!--          Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE)) -->
 
-ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")),
-                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")),
-                               expression(paste("AA and ", italic("M. calida"), sep = "")),
-                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
-  labs(y = "Percent of Genes",
-       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
-        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
-        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
-        legend.text.align = 0,                                                            # Aligns legend text to the left
-        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
+<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), -->
+<!--                                expression(paste("NT and ", italic("M. gaviniae"), sep = "")), -->
+<!--                                expression(paste("AA and ", italic("M. calida"), sep = "")), -->
+<!--                                expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) + -->
+<!--   labs(y = "Percent of Genes", -->
+<!--        fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) + -->
+<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
+<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
+<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
+<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
+<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
 
-```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6.
+<!-- ```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6.
 The most similar taxa to *M.
@@
 calida* sequence.
-Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5}
-FI_calida_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6_CircularIdentities_NT.csv",
-                         stringsAsFactors = FALSE) %>%
-  mutate(Closest_Relative = factor(Closest_Relative,
-                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
-                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5} -->
+<!-- FI_calida_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6_CircularIdentities_NT.csv", -->
+<!--                          stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
+<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
+<!--                                               "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE)) -->
 
-ggplot(FI_calida_NT, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
-  coord_polar() +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringae")) +
-  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
-  labs(fill = "Most Similar Taxa") +
-  theme_bw() +
-  theme(legend.text = element_text(face = "italic"),
-        axis.title = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
+<!-- ggplot(FI_calida_NT, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
+<!--   coord_polar() + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
+<!--                                "P. syringae")) + -->
+<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
+<!--   labs(fill = "Most Similar Taxa") + -->
+<!--   theme_bw() + -->
+<!--   theme(legend.text = element_text(face = "italic"), -->
+<!--         axis.title = element_blank(), -->
+<!--         axis.text.y = element_blank(), -->
+<!--         axis.ticks.y = element_blank(), -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
 
-```{r Figure7, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 7.
+<!-- ```{r Figure7, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 7.
 The most similar taxa to *M.
@@
 calida* sequence.
-Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5}
-FI_calida_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_CircularIdentities_AA.csv",
-                         stringsAsFactors = FALSE) %>%
-  mutate(Closest_Relative = factor(Closest_Relative,
-                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
-                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))
+Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5} -->
+<!-- FI_calida_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_CircularIdentities_AA.csv", -->
+<!--                          stringsAsFactors = FALSE) %>% -->
+<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
+<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
+<!--                                               "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE)) -->
 
-ggplot(FI_calida_AA, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
-  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
-  coord_polar() +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringae")) +
-  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
-  labs(fill = "Most Similar Taxa") +
-  theme_bw() +
-  theme(legend.text = element_text(face = "italic"),
-        axis.title = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
+<!-- ggplot(FI_calida_AA, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
+<!--   coord_polar() + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
+<!--                                "P. syringae")) + -->
+<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
+<!--   labs(fill = "Most Similar Taxa") + -->
+<!--   theme_bw() + -->
+<!--   theme(legend.text = element_text(face = "italic"), -->
+<!--         axis.title = element_blank(), -->
+<!--         axis.text.y = element_blank(), -->
+<!--         axis.ticks.y = element_blank(), -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
 
-```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
-stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE)
+<!-- ```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE} -->
+<!-- stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE) -->
 
-same_spp <- data.frame(table(stats_on_all$Same_Species))
-same_spp <- same_spp[which(same_spp$Var1 != "Nope"), ]
+<!-- same_spp <- data.frame(table(stats_on_all$Same_Species)) -->
+<!-- same_spp <- same_spp[which(same_spp$Var1 != "Nope"), ] -->
 
-same_spp <- rbind(same_spp, data.frame(Var1 = "T_ptyseos", Freq = 0))
+<!-- same_spp <- rbind(same_spp, data.frame(Var1 = "T_ptyseos", Freq = 0)) -->
 
-colnames(same_spp) <- c("Species", "Number of Genes")
+<!-- colnames(same_spp) <- c("Species", "Number of Genes") -->
 
-same_spp <- mutate(same_spp,
-                   Species = case_when(Species == "E_amylovora" ~ "E. amylovora",
-                                       Species == "E_tasmaniensis" ~ "E. tasmaniensis",
-                                       Species == "P_agglomerans" ~ "P. agglomerans",
-                                       Species == "P_septica" ~ "P. septica",
-                                       Species == "T_ptyseos" ~ "T. ptyseos",
-                                       Species == "T_saanichensis" ~ "T. saanichensis",
-                                       Species == "E_cloacae" ~ "E. cloacae")) %>%
-  mutate(Species = factor(Species, levels = c("P. agglomerans", "P. septica", "E. amylovora", "E. tasmaniensis",
-                                              "T. ptyseos", "T. saanichensis", "E. cloacae"), ordered = TRUE))
+<!-- same_spp <- mutate(same_spp, -->
+<!--                    Species = case_when(Species == "E_amylovora" ~ "E. amylovora", -->
+<!--                                        Species == "E_tasmaniensis" ~ "E. tasmaniensis", -->
+<!--                                        Species == "P_agglomerans" ~ "P. agglomerans", -->
+<!--                                        Species == "P_septica" ~ "P. septica", -->
+<!--                                        Species == "T_ptyseos" ~ "T. ptyseos", -->
+<!--                                        Species == "T_saanichensis" ~ "T. saanichensis", -->
+<!--                                        Species == "E_cloacae" ~ "E. cloacae")) %>% -->
+<!--   mutate(Species = factor(Species, levels = c("P. agglomerans", "P. septica", "E. amylovora", "E. tasmaniensis", -->
+<!--                                               "T. ptyseos", "T. saanichensis", "E. cloacae"), ordered = TRUE)) -->
 
-same_spp <- arrange(same_spp, Species)
+<!-- same_spp <- arrange(same_spp, Species) -->
 
-kable(same_spp, caption = "Table 5. The number of genes that had each species as the closest or most similar relative for all four analyses.")
+<!-- kable(same_spp, caption = "Table 5. The number of genes that had each species as the closest or most similar relative for all four analyses.") -->
 
-# table(stats_on_all$Same_Genus)
-# table(stats_on_all$NT_Same)
-# table(stats_on_all$AA_Same)
-# table(stats_on_all$Rel_Same)
-# table(stats_on_all$Ide_Same)
-```
+<!-- # table(stats_on_all$Same_Genus) -->
+<!-- # table(stats_on_all$NT_Same) -->
+<!-- # table(stats_on_all$AA_Same) -->
+<!-- # table(stats_on_all$Rel_Same) -->
+<!-- # table(stats_on_all$Ide_Same) -->
+<!-- ``` -->
 
-```{r Figure8, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 8. Sample for first 100 genes.", fig.width=6.5, fig.height=3}
-all_results_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureN_AllResultsCalida.csv",
-                               stringsAsFactors = TRUE) %>%
-  pivot_longer(cols = c(Rel_NT:Ide_AA), names_to = "Analysis", values_to = "Closest_Similar") %>%
-  mutate(Closest_Similar = factor(Closest_Similar,
-                                  levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis",
-                                             "E_cloacae", "P_syringiae"),
-                                  ordered = TRUE),
-         Analysis = factor(Analysis, levels = c("Ide_AA", "Rel_AA", "Ide_NT", "Rel_NT"), ordered = TRUE),
-         Gene_Name = reorder(Gene_Name, ID))
+<!-- ```{r Figure8, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 8. Sample for first 100 genes.", fig.width=6.5, fig.height=3} -->
+<!-- all_results_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureN_AllResultsCalida.csv", -->
+<!--                                stringsAsFactors = TRUE) %>% -->
+<!--   pivot_longer(cols = c(Rel_NT:Ide_AA), names_to = "Analysis", values_to = "Closest_Similar") %>% -->
+<!--   mutate(Closest_Similar = factor(Closest_Similar, -->
+<!--                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", -->
+<!--                                              "E_cloacae", "P_syringiae"), -->
+<!--                                   ordered = TRUE), -->
+<!--          Analysis = factor(Analysis, levels = c("Ide_AA", "Rel_AA", "Ide_NT", "Rel_NT"), ordered = TRUE), -->
+<!--          Gene_Name = reorder(Gene_Name, ID)) -->
 
-ggplot(all_results_calida[1:400, ], aes(x = Gene_Name, y = Analysis, fill = Closest_Similar)) +
-  geom_tile() +
-  scale_fill_brewer(palette = "Paired",
-                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
-                               "P. syringae")) +
-  scale_y_discrete(labels = c("AA Identity", "AA Closest", "NT Identity", "NT Closest")) +
-  labs(fill = "Closest Relative or \n Most Similar Taxa") +
-  theme(legend.text = element_text(face = "italic"),
-        # legend.position = "bottom",
-        text = element_text(size = 12,  family = "Times New Roman"),
-        axis.text.x = element_text(angle = 45, hjust = 1))               # Rotates x axis labels
+<!-- ggplot(all_results_calida[1:400, ], aes(x = Gene_Name, y = Analysis, fill = Closest_Similar)) + -->
+<!--   geom_tile() + -->
+<!--   scale_fill_brewer(palette = "Paired", -->
+<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
+<!--                                "P. syringae")) + -->
+<!--   scale_y_discrete(labels = c("AA Identity", "AA Closest", "NT Identity", "NT Closest")) + -->
+<!--   labs(fill = "Closest Relative or \n Most Similar Taxa") + -->
+<!--   theme(legend.text = element_text(face = "italic"), -->
+<!--         # legend.position = "bottom", -->
+<!--         text = element_text(size = 12,  family = "Times New Roman"), -->
+<!--         axis.text.x = element_text(angle = 45, hjust = 1))               # Rotates x axis labels -->
 
-stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE)
+<!-- stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE) -->
 
-# table(stats_on_all$Same_Species)
-# table(stats_on_all$Same_Genus)
-# table(stats_on_all$NT_Same)
-# table(stats_on_all$AA_Same)
-# table(stats_on_all$Rel_Same)
-# table(stats_on_all$Ide_Same)
-```
+<!-- # table(stats_on_all$Same_Species) -->
+<!-- # table(stats_on_all$Same_Genus) -->
+<!-- # table(stats_on_all$NT_Same) -->
+<!-- # table(stats_on_all$AA_Same) -->
+<!-- # table(stats_on_all$Rel_Same) -->
+<!-- # table(stats_on_all$Ide_Same) -->
+<!-- ``` -->
 
-```{r Figure9, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 9. Maximum likelihood gene trees for ttuB for nucleotide (A) and amino acid sequences (B).", fig.width=3.25}
-knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_NT.png")
-knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_AA.png")
-```
+<!-- ```{r Figure9, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 9. Maximum likelihood gene trees for ttuB for nucleotide (A) and amino acid sequences (B).", fig.width=3.25} -->
+<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_NT.png") -->
+<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_AA.png") -->
+<!-- ``` -->
 
-```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10.
+<!-- ```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10.
 Maximum likelihood gene trees for secY for nucleotide (A) and amino acid sequences (B).
-Since bifucations could not be seen in radial trees, condensed trees were computed with a 70% cutoff for both the nucleotide sequnces (C) and amino acid sequences (D).", fig.width=3.25}
-knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_NT.png")
-knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_AA.png")
-knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_NT70.png")
-knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_AA70.png")
-```
+Since bifucations could not be seen in radial trees, condensed trees were computed with a 70% cutoff for both the nucleotide sequnces (C) and amino acid sequences (D).", fig.width=3.25} -->
+<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_NT.png") -->
+<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_AA.png") -->
+<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_NT70.png") -->
+<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_AA70.png") -->
+<!-- ``` -->
 

### 2021-04-19 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -71,7 +71,17 @@ kable(models,
       genetic distance analysis.")
 # ```
 
-# Figure 1. First Relative: Best vs Available Model -------------------------------------------------------------------------------------------------------
+models <- data.frame(Gene = DM_models_NT$Gene,
+                     DM = DM_models_NT$Code, 
+                     PM = phylo_models_NT$BM)                                       # To check if the models are the same
+
+models$Same_Model <- models$DM == models$PM
+
+table(models$Same_Model)
+# FALSE  TRUE 
+# 242    557 
+
+# Figure 1. Closest Relative: Best vs Available Model -----------------------------------------------------------------------------------------------------
 GD_first_relative <- read.csv("O8_Results_Ten_NT/Four_Relatives_Ext_Ten_NT.csv", 
                               stringsAsFactors = FALSE)                             # Read in the closest relative results from Genetic Distances
 
@@ -192,16 +202,79 @@ ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
         text = element_text(size = 12,  family = "Times New Roman"))
 # ```
 
+difCR_cal <- cbind(models, test_cal)                                                # Did differences in models change CR?
+unique(difCR_cal$Gene == difCR_cal$PD_Gene)                                         # Checks that gene order is the same
+
+dif_models <- subset(difCR_cal, Same_Model == FALSE)                                # When recommended and best available models were different
+table(dif_models$CR_test)                                                           # Number of genes that had the same CR despite different models
+# FALSE  TRUE 
+# 49     193 
+
+same_models <- subset(difCR_cal, Same_Model == TRUE)                                # When recommended and best available models were the same
+table(same_models$CR_test)                                                          # Number of genes that had the same CR with same models
+# FALSE  TRUE 
+# 111    446 
+
+# Figure 2. Closest Relative: NTS and AAS -----------------------------------------------------------------------------------------------------------------
+datasets <- data.frame(project = c("NT", "NT", "AA", "AA"),
+                       dataset = c("calida_NT", "gaviniae_NT", "calida_AA", 
[…]

### 2021-04-19 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 # Results
+After the two filters, 799 of 954 homologous genes remained.
+The genes included those encoding rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the M.
+calida and M.
+gaviniae genomes.
+
+## Best Model vs Best Available Model
+In MEGAX [@Kumar2018], the user can either calculate genetic distance or estimate a phylogenetic tree.
+However, although model testing assesses all possible models available in the software, all models are not available for genetic distance analysis.
+The GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models and the +I (invariant sites) parameter are available only for phylogenetic tree analysis.
+
+The genetic distances for the nucleotide sequences (NTS) were estimated using both the recommended (RM) and best available models (BAM) for distance analysis to determine if different results would arise when using a model that does not explain most of the genetic variance.
+The RM is the model with the lowest Bayesian Information Criterion (BIC) among the available options.
+First, phylogenetic gene trees were computed using the recommended model from model testing.
+The software suggested twelve unique models for this method (Table 2).
+Genetic distance matrices were then extracted from the trees.
+Second, distance matrices were calculated directly using the BAM for calculating pairwise distances.
+Only four models were recommended for genetic distance analysis (Table 2).
+Of the 799 NTS, 242 genes required a different model between the two analyses, and the mean BIC difference between the two models was 13.8 (range: 0.1, 122.6).
+
 ```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
@@
 
+The closest relatives to both of the *Mixta* species were extracted from RM and BAM distance matrices.
+When the RM was used.
+*P.
+septica* was the closest relative to 66.3 and 66.5% of the *M.
+calida* and *M.
+gaviniae* sequences, respectively (Figure 1).
+These percentages increase by 12.6 and 14.4% for the *M.
+calida* and *M.
+gaviniae* sequences, respectively, when the BAM was used (Figure 1).
+Instead of *P.
+septica* when the RM is used, more genes are more closely related to *P.
+agglomerans*, *E.
+tasmaniensis*, *E.
+amylovora*, *T.
+saanichensis*, and *E.
+cloacae* (Figure 1).
+Additionally, there was more disparity between *M.
+calida* and *M.
+gaviniae* when the BAM was used (Figure 1).
+
 ```{r Figure1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1.
@@
 The species with the shortest genetic distance from the *Mixta* species is considered to be the closest relative.
-The pink and orange bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the recommended model and the best available model, respectively.
+The pink and orange bars represent the percentage of *Mixta* genes that were most closely related to any of the non-*Mixta* species when using the recommended model and the best available model, respectively.
 The lighter shades represent *M.
@@
 
+The RM and BAM were different for approximately 30.3% (n = 242) of the genes.
+Of these 242 genes, 20.2% (n = 49) were said to have different closest relatives between the two analyses.
+However, when the RM and BAM were the same (n = 557), 19.9% (n = 111) had different closest relatives.
+
+## Closest Relative from Genetic Distance
+Genetic distance gives a numerical estimate of evolutionary change between species, and the taxa with which a species of interest has the shortest distance is regarded as the closest relative.
+Since the phylogenetic trees were calculated using the RM, the distance matrices exported from the trees were kept for future analyses.
+*P.
+septica* was consistently the closest relative for most of the *Mixta* sequences (Figure 2).
+However, the number of genes most closely related *P.
+septica* decreased from 66.3 and 66.5% (*M.
+calida* and *M.
+gaviniae*, respectively) with the NTS by 23.7% with the amino acid sequences (AAS; Figure 2).
+Instead, the *Mixta* AAS were most closely related to the other taxa within the dataset except for *P.
+syringae* (Figure 2).
+Interestingly, the number of genes most closely related to *E.
+cloacae* increased between the NTS and the AAS from 2.9% by 3.1% (Figure 2).
+
+```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (n = 799) that were most closely related to any of the eight non-*Mixta* species using the recommended models and genetic distances.
+The species with the shortest genetic distance from the *Mixta* species is considered to be the closest relative.
+The percentage of *Mixta* genes that were most closely related to any of the non-*Mixta* species are represented by the blue and green bars for the nucleotide and amino acid sequences, respectively.
+The lighter shades represent *M.
+calida* and the darker shades represent *M.
+gavinaie*.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", 
+                             stringsAsFactors = FALSE) %>%
+  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                              "E. cloacae", "P. syringae"), ordered = TRUE),
+         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
+
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
 <!-- # Old Results -->
@@
 
-<!-- ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
+```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
 The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange).
-Models are for nucleotide sequences.", fig.width=6.5} -->
-<!-- models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE) -->
+Models are for nucleotide sequences.", fig.width=6.5}
+models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE)
 
-<!-- ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
-<!--   scale_fill_brewer(palette = "Dark2", -->
-<!--                     labels = c("Phylogenetic", "Genetic Distance")) + -->
-<!--   labs(y = "Percent of Genes \n Requiring Each Model", -->
-<!--        fill = "Analysis") + -->
-<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels -->
-<!--         legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner -->
-<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
-<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
-<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Dark2",
+                    labels = c("Phylogenetic", "Genetic Distance")) +
+  labs(y = "Percent of Genes \n Requiring Each Model",
+       fill = "Analysis") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels
+        legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
 

### 2021-04-20 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -221,6 +221,7 @@ datasets <- data.frame(project = c("NT", "NT", "AA", "AA"),
                                    "gaviniae_AA"))                                  # The four different datasets: two Mixta and NTS/AAS
 
 number_relatives <- data.frame(matrix(ncol = 4, nrow = 0))                          # Initialize the combined dataset
+CR_four <- data.frame(matrix(ncol = 0, nrow = 799))
 
 for(row in 1:nrow(datasets)) {
   four_rel <- read.csv(paste("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/9_Results_", 
@@ -238,6 +239,13 @@ for(row in 1:nrow(datasets)) {
                     Type = datasets$dataset[row])                                   # Which dataset
   
   number_relatives <- rbind(number_relatives, num_rel)                              # Combine everything
+  
+  first_rel <- subset(four_rel, select = c(Gene, Closest_Relative))                 # Get the genes and list of closest relatives
+  
+  colnames(first_rel) <- c(paste(datasets$dataset[row], "_Gene", sep = ""), 
+                           paste(datasets$dataset[row], "_CR", sep = ""))           # Change the column names so that they're distinguishable
+  
+  CR_four <- cbind(CR_four, first_rel)
 }; rm(four_rel, num_rel, row)
 
 number_relatives <- mutate(number_relatives,
@@ -246,12 +254,36 @@ number_relatives <- mutate(number_relatives,
 
 write.csv(number_relatives, "10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", row.names = FALSE)
 
-# ```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2. The percentage of genes (n = 799) that were most closely related to 
-# any of the eight non-*Mixta* species using genetic distances. The species with the shortest genetic distance from both of the *Mixta* species is 
-# considered to be the closest relative. Blue bars are the nucleotide sequences of the genes with light blue bars representing the percentage of *M. 
-# calida* genes and dark blue bars representing *M. gaviniae*. Green bars are the corresponding amino acid sequences with light green representing *M. 
-# calida* genes and dark blue bars representing *M. gaviniae*. Approximately 66% of *M. calida* and *M. gaviniae* genes were most closely related to *P. 
[…]

### 2021-04-20 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

--- 2021-04-19 Manuscript/Manuscript_Paper.Rmd
@@
 # Results
+After the two filters, 799 of 954 homologous genes remained.
+The genes included those encoding rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the M.
+calida and M.
+gaviniae genomes.
+
+## Best Model vs Best Available Model
+
 ```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
@@
         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+## Closest Relative from Genetic Distance
+
+```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
+The percentage of genes (n = 799) that were most closely related to any of the eight non-*Mixta* species using the recommended models and genetic distances.
+The species with the shortest genetic distance from the *Mixta* species is considered to be the closest relative.
+The percentage of *Mixta* genes that were most closely related to any of the non-*Mixta* species are represented by the blue and green bars for the nucleotide and amino acid sequences, respectively.
+The lighter shades represent *M.
+calida* and the darker shades represent *M.
+gavinaie*.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", 
+                             stringsAsFactors = FALSE) %>%
+  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                              "E. cloacae", "P. syringae"), ordered = TRUE),
+         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
+
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+## Spatial Patterns in the Closest Relatives
+
+```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
+The closest relatives to *M.
+calida* genes for nucleotide sequences around the *M.
+calida* chromosome.
+The length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
+The *M.
+calida* genome has 4092 genes.", fig.width=6.5, fig.height=5}
+FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
+
+ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+```
+
+```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.
+The closest relatives to *M.
+calida* genes for amino acid sequences around the *M.
+calida* chromosome.
+The length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
+The *M.
+calida* genome has 4092 genes.", fig.width=6.5, fig.height=5}
+FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
+
+ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
         text = element_text(size = 12,  family = "Times New Roman"))
@@
 The *M.
-calida* genome has 4084 genes.
+calida* genome has 4092 genes.
 No distinct patterns emerged; however, genes most closely related to *P.

### 2021-04-20 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -7,6 +7,7 @@ library("tidyverse")
 library("msa")
 library("ggplot2")
 library("cowplot")
+library("ggpubr")
 library("RColorBrewer")
 library("extrafont")
 # loadfonts(device = "win")
@@ -308,7 +309,7 @@ ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
         text = element_text(size = 12,  family = "Times New Roman"))
 # ```
 
-# Figure 3. Circular Plot: Nucleotide and M. calida -------------------------------------------------------------------------------------------------------
+# Figure 3. Circular Plot: NTS+AAS and M. calida ----------------------------------------------------------------------------------------------------------
 FR_nucl_calida <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE) %>%
   mutate(Relative_Number = case_when(Closest_Relative == "P_septica" ~ 3, Closest_Relative == "P_agglomerans" ~ 4,
                                      Closest_Relative == "E_tasmaniensis" ~ 5, Closest_Relative == "E_amylovora" ~ 6,
@@ -323,31 +324,6 @@ FR_nucl_calida <- rbind(FR_nucl_calida, extra_genes_NT)                         
 
 write.csv(FR_nucl_calida, "10_Tables_and_Figures/Figure3_Circular_NT.csv", row.names = FALSE)
 
-# ```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3. The closest relatives to *M. calida* genes for nucleotide sequences around 
-# the *M. calida* chromosome. the length and colour of the bars represent the different species to which the gene is most closely related according to 
-# genetic distance. The *M. calida* genome has 4084 genes.", fig.width=6.5, fig.height=5}
-FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
-  mutate(Closest_Relative = factor(Closest_Relative, 
-                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
-                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
-
-ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
[…]

### 2021-04-20 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

--- 2021-04-19 Manuscript/Manuscript_Paper.Rmd
@@
 library("ggplot2")
+library("ggpubr")
 library("RColorBrewer")
@@
 Instead of *P.
-septica* when the RM is used, more genes are more closely related to *P.
+septica* when the RM is used, more genes were more closely related to *P.
 agglomerans*, *E.
@@
 cloacae* increased between the NTS and the AAS from 2.9% by 3.1% (Figure 2).
+Additionally, 55.1 and 55.2% of the *M.
+calida* and *M.
+gaviniae* sequences, respectively, had the same closest relatives between the NTS and the AAS analyses.
 
@@
 
+## Spatial Patterns in the Closest Relatives
+Circular plots were generated using the *Mixta* gene IDs to simulate their relative positions on the *Mixta* chromosomes and uncover positional and spatial patterns.
+The bars’ lengths and colours correspond to the species that was the closest relative to the *Mixta* gene at that position.
+The results between the two *Mixta* species for the closest relatives were the same except for one and two genes for the NTS and the AAS, respectively; therefore, only the results for *M.
+calida* are shown.
+
+Many of the genes most closely related to any of the other taxa in this study were evenly positioned around the *Mixta* chromosome in the NTS (Figure 3A).
+The exception is for the *Tatumella* species (pinks) and *E.
+amylovora* (dark green; Figure 3A).
+The former and most of the genes closely related to the latter were located on the right side of the chromosome (Figure 3A).
+Unexpectedly, genes closely related to *E.
+cloacae* (orange) were distributed throughout the *Mixta* genome.
+
+```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
+The closest relatives to *M.
+calida* genes for (**A**) nucleotide and (**B**) amino acid sequences around the *M.
+calida* chromosome.
+The length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
+The *M.
+calida* genome has 4092 genes.", fig.width=6.5, fig.height=3.9}
+FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
+
+FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
+
+cal_NT <- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+cal_AA <- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+ggarrange(cal_NT, cal_AA, 
+          labels = c("A", "B"),
+          font.label = list(family = "Times New Roman"),
+          label.x = 0.04,
+          label.y = 0.98,
+          common.legend = TRUE, 
+          legend = "bottom")
+```
+
+In the AAS analysis, the genes closely related to the *Tatumella* species were not exclusive to the right half (Figure 3B).
+Additionally, the genes closely related to the *Erwinia* species and *E.
+cloacae* were more evenly distributed across the genome (Figure 3B).
+Similar to the NTS analysis, no gene clusters related to a single taxon were evident (Figure 3B).
+
 <!-- # Old Results -->
@@
 
-```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
+<!-- ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
 The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange).
-Models are for nucleotide sequences.", fig.width=6.5}
-models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE)
+Models are for nucleotide sequences.", fig.width=6.5} -->
+<!-- models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE) -->
 
-ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  scale_fill_brewer(palette = "Dark2",
-                    labels = c("Phylogenetic", "Genetic Distance")) +
-  labs(y = "Percent of Genes \n Requiring Each Model",
-       fill = "Analysis") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels
-        legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner
-        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
-        legend.text.align = 0,                                                            # Aligns legend text to the left
-        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
-        text = element_text(size = 12,  family = "Times New Roman"))
-```
+<!-- ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   scale_fill_brewer(palette = "Dark2", -->
+<!--                     labels = c("Phylogenetic", "Genetic Distance")) + -->
+<!--   labs(y = "Percent of Genes \n Requiring Each Model", -->
+<!--        fill = "Analysis") + -->
+<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels -->
+<!--         legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner -->
+<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
+<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
+<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
+<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
+<!-- ``` -->
 

### 2021-04-22 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 calida* and the darker shades represent *M.
-gavinaie*.", fig.width=6.5}
+gaviniae*.", fig.width=6.5}
 number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeBestvsAvailableModelNT.csv", 
@@
 calida* and the darker shades represent *M.
-gavinaie*.", fig.width=6.5}
+gaviniae*.", fig.width=6.5}
 number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", 
@@
 calida* genome has 4092 genes.", fig.width=6.5, fig.height=3.9}
-FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
+FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3A_Circular_NT.csv", stringsAsFactors = FALSE) %>%
   mutate(Closest_Relative = factor(Closest_Relative, 
@@
 
-FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
+FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3B_Circular_AA.csv", stringsAsFactors = FALSE) %>%
   mutate(Closest_Relative = factor(Closest_Relative, 
@@
         axis.title = element_blank(),
+        panel.border = element_blank(),
         axis.text.y = element_blank(),
@@
 
-<!-- # Old Results -->
-<!-- ```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE} -->
-<!-- PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE) -->
+## Closest Relative from Average Site Identity
+Site identity is the amount of similarity between two sequences and accounts for the sequence length.
+The non-*Mixta* sequences were aligned to each of the *Mixta* homologs, so the percent identity was relative to *Mixta*.
+The closest relative to both *Mixta* species is defined as the taxon with the highest similarity to *Mixta*.
 
-<!-- kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.") -->
-<!-- ``` -->
+Similar to the closest relative according to genetic distances, *P.
+septica* was the closest relative for most *Mixta* sequences (Figure 4).
+With the NTS, *P.
+septica* was the closest relative for 79.3 and 80.6% of the *M.
+calida* and *M.
+gaviniae* sequences, respectively (Figure 4).
+However, these percentage decreased by 31.5 and 33.0%, respectively, with the AAS (Figure 4).
+The *Mixta* AAS were instead more closely related to other taxa except for *P.
+syringae* (Figure 4).
+Unexpectedly, 1.9 and 1.3% of the *M.
+calida* and *M.
+gaviniae* sequences, respectively, were most similar to the *E.
+cloacae* sequences (Figure 4).
+Additionally, these percentages increased by 3.1 and 2.5% for *M.
+calida* and *M.
+gaviniae*, respectively, for the AAS (Figure 4).
 
-<!-- ```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE} -->
-<!-- PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_PmodelsAA.csv", stringsAsFactors = FALSE) -->
+```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.
+The percentage of genes (n = 799) that were most closely related to any of the eight non-*Mixta* species using percent site identity.
+The species with the highest percent identity to the *Mixta* species is considered to be the closest relative.
+The percentage of *Mixta* genes that were most closely related to any of the non-*Mixta* species are represented by the blue and green bars for the nucleotide and amino acid sequences, respectively.
+The lighter shades represent *M.
+calida* and the darker shades represent *M.
+gaviniae*.", fig.width=6.5}
+number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_FirstRelativeNTAA_Identity.csv", 
+                             stringsAsFactors = FALSE) %>%
+  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
+                                              "E. cloacae", "P. syringae"), ordered = TRUE),
+         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))
 
-<!-- kable(PM_AA, caption = "Table 3. According to model testing and the lowest BIC for amino acid sequences, the number of genes required each phylogenetic tree model.") -->
-<!-- ``` -->
+ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), 
+                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")), 
+                               expression(paste("AA and ", italic("M. calida"), sep = "")), 
+                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
+  labs(y = "Percent of Genes", 
+       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
+        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
+        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
+        legend.text.align = 0,                                                            # Aligns legend text to the left
+        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
+        text = element_text(size = 12,  family = "Times New Roman"))
 
-<!-- ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1.
-The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange).
-Models are for nucleotide sequences.", fig.width=6.5} -->
-<!-- models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE) -->
+```
 
-<!-- ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
-<!--   scale_fill_brewer(palette = "Dark2", -->
-<!--                     labels = c("Phylogenetic", "Genetic Distance")) + -->
-<!--   labs(y = "Percent of Genes \n Requiring Each Model", -->
-<!--        fill = "Analysis") + -->
-<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels -->
-<!--         legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner -->
-<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
-<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
-<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+When comparing the site identity analysis results, 89.2% of the NTS and 85.2% of the AAS agreed on the same closest relative between the two *Mixta* species.
+In contrast, only 55.1 and 53.9% of the *M.
+calida* and *M.
+gaviniae* sequences, respectively, had the same results between the NTS and AAS analyses.
 
-<!-- ```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2.
-The percentage of genes (n = 799) most closely related to any of the eight non-*Mixta* species.
-Closest relative was inferred using shortest genetic distance.
-Blue bars represent the nucleotide sequences and green bars represent the amino acid sequences.
-The lighter of each colour represents *M.
-calida* and the darker *M.
-gaviniae*.", fig.width=6.5} -->
-<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", -->
-<!--                              stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
-<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
-<!--          Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE)) -->
+## Next Spatial
 
-<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), -->
-<!--                                expression(paste("NT and ", italic("M. gaviniae"), sep = "")), -->
-<!--                                expression(paste("AA and ", italic("M. calida"), sep = "")), -->
-<!--                                expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) + -->
-<!--   labs(y = "Percent of Genes", -->
-<!--        fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) + -->
-<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
-<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
-<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
-<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
-<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+```{r Figure5, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 5.
+The closest relatives to *M.
+calida* genes for (**A**) nucleotide and (**B**) amino acid sequences around the *M.
+calida* chromosome.
+The length and colour of the bars represent the different species to which the gene is most closely related according to site identity.
+The *M.
+calida* genome has 4092 genes.", fig.width=6.5, fig.height=3.9}
+FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5A_Circular_NTcalida_Identity.csv", 
+                           stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
 
-<!-- ```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2.
-The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences.
-The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative.
-Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation.
-Nevertheless, this model may not be available in software programs or analysis methods within a software program.
-Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model.
-Orange bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the best available genetic distance model analysis in MEGAX.
-GTR and HKY models and invariant sites parameter were not available for this analysis.
-The lighter shades represent *M.
-calida*, and the darker shades represent *M.
-gaviniae*.", fig.width=6.5} -->
-<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv", -->
-<!--                              stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
-<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
-<!--          Type = factor(Type, levels = c("PD_calida", "PD_gaviniae", "GD_calida", "GD_gaviniae"), ordered = TRUE)) -->
+FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5B_Circular_AAcalida_Identity.csv", 
+                         stringsAsFactors = FALSE) %>%
+  mutate(Closest_Relative = factor(Closest_Relative, 
+                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
+                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))
 
-<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
-<!--   scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"), -->
-<!--                     labels = c(expression(paste("Best Model and ", italic("M. calida"), sep = "")), -->
-<!--                                expression(paste("Best Model and ", italic("M. gaviniae"), sep = "")), -->
-<!--                                expression(paste("Best Available Model and ", italic("M. calida"), sep = "")), -->
-<!--                                expression(paste("Best Available Model and ", italic("M. gaviniae"), sep = "")))) + -->
-<!--   labs(y = "Percent of Genes", -->
-<!--        fill = expression(paste("Model and ", italic("Mixta"), " species", sep = ""))) + -->
-<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
-<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
-<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
-<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
-<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+cal_NT <- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        panel.border = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
 
-<!-- ```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3.
-Closest relatives to *M.
-calida* genes for nucleotide sequences around the *M.
-calida* genome.
-Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
-The *M.
-calida* genome has 4092 genes.
-No distinct patterns emerged; however, genes most closely related to *P.
-septica* are the most common and occur throughout the genome.
-Therefore, *P.
-septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5} -->
-<!-- FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
-<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
-<!--                                               "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE)) -->
+cal_AA <- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        panel.border = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
 
-<!-- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
-<!--   coord_polar() + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
-<!--                                "P. syringae"), -->
-<!--                     na.value = "#ff7f00") + -->
-<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
-<!--   labs(fill = "Closest Relative") + -->
-<!--   theme_bw() + -->
-<!--   theme(legend.text = element_text(face = "italic"), -->
-<!--         axis.title = element_blank(), -->
-<!--         axis.text.y = element_blank(), -->
-<!--         axis.ticks.y = element_blank(), -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+ggarrange(cal_NT, cal_AA, 
+          labels = c("A", "B"),
+          font.label = list(family = "Times New Roman"),
+          label.x = 0.04,
+          label.y = 0.98,
+          common.legend = TRUE, 
+          legend = "bottom")
+```
 
-<!-- ```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4.
-Closest relatives to *M.
-calida* genes for amino acid sequences around the *M.
-calida* genome.
-Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance.
-The *M.
-calida* genome has 4092 genes.
-No distinct patterns emerged; however, genes most closely related to *P.
-septica* are the most common and occur throughout the genome.
-Therefore, *P.
-septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5} -->
-<!-- FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
-<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
-<!--                                               "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE)) -->
-
-<!-- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
-<!--   coord_polar() + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
-<!--                                "P. syringae"), -->
-<!--                     na.value = "#ff7f00") + -->
-<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
-<!--   labs(fill = "Closest Relative") + -->
-<!--   theme_bw() + -->
-<!--   theme(legend.text = element_text(face = "italic"), -->
-<!--         axis.title = element_blank(), -->
-<!--         axis.text.y = element_blank(), -->
-<!--         axis.ticks.y = element_blank(), -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
-
-<!-- ```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5.
-The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using percent site identities.
-The species with the highest percent identity to each of the *Mixta* species is considered the most similar taxon for a single gene.
-Blue bars represent the nucleotide sequences' results; the light blue bars represent *M.
-calida* sequences, and the dark blue bars represent *M.
-gaviniae* sequences.
-Green bars represent the amino acid sequences' results; the light green bars represent *M.
-calida* sequences, and the dark green bars represent *M.
-gaviniae* sequences.", fig.width=6.5} -->
-<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5_FirstRelativeMixtaNTAAIdentity.csv", -->
-<!--                              stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
-<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
-<!--          Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE)) -->
-
-<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), -->
-<!--                                expression(paste("NT and ", italic("M. gaviniae"), sep = "")), -->
-<!--                                expression(paste("AA and ", italic("M. calida"), sep = "")), -->
-<!--                                expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) + -->
-<!--   labs(y = "Percent of Genes", -->
-<!--        fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) + -->
-<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
-<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
-<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
-<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
-<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
-
+# Old Results
 <!-- ```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6.

### 2021-04-22 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -174,12 +174,12 @@ number_relatives$Species <- gsub("_", ". ", number_relatives$Species)
 
 write.csv(number_relatives, "10_Tables_and_Figures/Figure1_FirstRelativeBestvsAvailableModelNT.csv", row.names = FALSE)
 
-# ```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2. The percentage of genes (n = 799) that were most closely related to any 
+# ```{r Figure1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1. The percentage of genes (n = 799) that were most closely related to any 
 # of the eight non-*Mixta* species when using the recommended model from model testing versus the best available) model for genetic distance analysis in 
 # MEGAX. The sequences were nucleotide sequences. The species with the shortest genetic distance from the *Mixta* species is considered to be the closest 
 # relative. The pink and orange bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using 
 # the recommended model and the best available model, respectively. The lighter shades represent *M. calida* and the darker shades represent 
-# *M. gavinaie*.", fig.width=6.5}
+# *M. gaviniae*.", fig.width=6.5}
 number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeBestvsAvailableModelNT.csv", 
                              stringsAsFactors = FALSE) %>%
   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
@@ -216,7 +216,7 @@ table(same_models$CR_test)                                                      
 # FALSE  TRUE 
 # 111    446 
 
-# Figure 2. Closest Relative: NTS and AAS -----------------------------------------------------------------------------------------------------------------
+# Figure 2. Closest Relative: Genetic Distance ------------------------------------------------------------------------------------------------------------
 datasets <- data.frame(project = c("NT", "NT", "AA", "AA"),
                        dataset = c("calida_NT", "gaviniae_NT", "calida_AA", 
[…]

### 2021-04-22 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
-## Spatial Patterns in the Closest Relatives
+## Spatial Patterns in the Closest Relatives from Genetic Distances
 Circular plots were generated using the *Mixta* gene IDs to simulate their relative positions on the *Mixta* chromosomes and uncover positional and spatial patterns.
@@
 
-## Next Spatial
+## Spatial Patterns from Average Site Identity
 

### 2021-04-23 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -156,6 +156,18 @@ table(test_gav$CR_test)
 # FALSE  TRUE 
 # 178     621 
 
+### HERE!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+test <- data.frame(Gene = test_cal$PD_Gene, 
+                   cal_PD = test_cal$PD_CR, 
+                   cal_GD = test_cal$GD_CR2, 
+                   gav_PD = test_gav$PD_CR, 
+                   gav_GD = test_gav$GD_CR2)
+test <- mutate(test,
+               cal = cal_PD == cal_GD,
+               gav = gav_PD == gav_GD,
+               PD = cal_PD == gav_PD,
+               GD = cal_GD == gav_GD)
+
 # Gather it all together
 number_relatives <- rbind(data.frame(Species = Spp,                                 # Counts the number of genes belonging to each taxa for analysis
                                      Number = count_relatives(GD_calida),           # Collects all four analyses into one (two each calida and gaviniae)
@@ -680,309 +692,140 @@ ggarrange(gav_NT, gav_AA,
           legend = "bottom")
 # ```
 
-# Next ---------------------------
+# Table 3. All Analyses Give Same Species ------------------------------------------------------------------------------------------------------------------
+CR_NT_cal_dist <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE) 
+CR_AA_cal_dist <- read.csv("9_Results_AA/four_relatives_calida_AA.csv", stringsAsFactors = FALSE)
+CR_NT_cal_iden <- read.csv("9_Results_NT/four_identities_calida_NT.csv", stringsAsFactors = FALSE) 
+CR_AA_cal_iden <- read.csv("9_Results_AA/four_identities_calida_AA.csv", stringsAsFactors = FALSE) 
+
+unique(CR_AA_cal_dist$Gene == CR_AA_cal_iden$Gene)
+unique(CR_AA_cal_dist$Gene == CR_NT_cal_dist$Gene)
+unique(CR_AA_cal_dist$Gene == CR_NT_cal_iden$Gene)
+
+CR_calida <- data.frame(Gene = CR_AA_cal_dist$Gene,
+                        ID = CR_AA_cal_dist$ID,
+                        NT_dist = CR_NT_cal_dist$Closest_Relative,
[…]

### 2021-04-23 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 theme_set(theme_classic())
+
+Spp <- c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", 
+         "E_cloacae", "P_syringae")
 ```
@@
 ## Spatial Patterns from Average Site Identity
+Positional effects were investigated in the site identity results.
+Only the results of *M.
+calida* are shown because the *M.
+gaviniae* results showed similar patterns.
+The genes most similar to either of the *Pantoea* species and *E.
+cloacae* were evenly dispersed across the *Mixta* chromosome for both the NTS and AAS (Figure 5A and 5B).
+Similar to the genetic distance results, genes that are labelled as being closely related to either of the *Tatumella* species are located only on right side of chromosome, and genes that are closely related to *E.
+amylovora* are predominantly on the right side as well (Figure 5A).
+Unexpectedly, the lower left quadrant of the chromosome contains primarily genes that are closely related to either of the *Pantoea* species with only a few exceptions (Figure 5A).
+	In the AAS analysis, genes closely related to any of the other taxa were more evenly distributed across the genome (Figure 5B).
 
@@
 
-# Old Results
-<!-- ```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6.
-The most similar taxa to *M.
-calida* genes for nucleotide sequences around its genome.
-The most similar taxon is that which sequence is most identical to the *M.
-calida* sequence.
-Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5} -->
-<!-- FI_calida_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6_CircularIdentities_NT.csv", -->
-<!--                          stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
-<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
-<!--                                               "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE)) -->
+## A Comparison of the Four Analyses
+The four above analyses (genetic distance and site identity, both using NTS and AAS) gave slightly different results for the *Mixta* sequences.
+When comparing the results, approximately 38.7% of both the *M.
+calida* and *M.
+gaviniae* sequences agreed upon a closest relative.
+However, the number of sequences closely related to any of the other taxa were slightly different between *M.
+calida* and *M.
+gaviniae* (Table 3).
+There were no genes for which all four analyses agreed on *T.
+ptyseos* being the closest relative (Table 3).
+Additionally, 7.5% of the *M.
+calida* and *M.
+gaviniae* homologs did not share the same closest relative.
 
-<!-- ggplot(FI_calida_NT, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
-<!--   coord_polar() + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
-<!--                                "P. syringae")) + -->
-<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
-<!--   labs(fill = "Most Similar Taxa") + -->
-<!--   theme_bw() + -->
-<!--   theme(legend.text = element_text(face = "italic"), -->
-<!--         axis.title = element_blank(), -->
-<!--         axis.text.y = element_blank(), -->
-<!--         axis.ticks.y = element_blank(), -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+```{r Table3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+CR_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_calida.csv", 
+                      stringsAsFactors = FALSE)
 
-<!-- ```{r Figure7, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 7.
-The most similar taxa to *M.
-calida* genes for amino acid sequences around its genome.
-The most similar taxon is that which sequence is most identical to the *M.
-calida* sequence.
-Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5} -->
-<!-- FI_calida_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_CircularIdentities_AA.csv", -->
-<!--                          stringsAsFactors = FALSE) %>% -->
-<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
-<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
-<!--                                               "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE)) -->
+# table(CR_calida$ALL)
+# FALSE  TRUE         # FALSE: 61.3%
+# 490    309          # TRUE: 38.7% of M. calida genes had the same closest relative in all four analyses
 
-<!-- ggplot(FI_calida_AA, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
-<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
-<!--   coord_polar() + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
-<!--                                "P. syringae")) + -->
-<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
-<!--   labs(fill = "Most Similar Taxa") + -->
-<!--   theme_bw() + -->
-<!--   theme(legend.text = element_text(face = "italic"), -->
-<!--         axis.title = element_blank(), -->
-<!--         axis.text.y = element_blank(), -->
-<!--         axis.ticks.y = element_blank(), -->
-<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
-<!-- ``` -->
+CR_gaviniae <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_gaviniae.csv", 
+                        stringsAsFactors = FALSE)
 
-<!-- ```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE} -->
-<!-- stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE) -->
+# table(CR_gaviniae$ALL)
+# FALSE  TRUE 
+# 490    309
 
-<!-- same_spp <- data.frame(table(stats_on_all$Same_Species)) -->
-<!-- same_spp <- same_spp[which(same_spp$Var1 != "Nope"), ] -->
+all_same_cal <- subset(CR_calida, ALL == TRUE); all_same_gav <- subset(CR_gaviniae, ALL == TRUE)
 
-<!-- same_spp <- rbind(same_spp, data.frame(Var1 = "T_ptyseos", Freq = 0)) -->
+all_same_species <- data.frame(table(all_same_cal$NT_dist),
+                               table(all_same_gav$NT_dist)) %>%
+  subset(select = -Var1.1) %>%
+  mutate(Var1 = factor(Var1, 
+                          levels = Spp, ordered = TRUE)) %>%
+  arrange(Var1)
 
-<!-- colnames(same_spp) <- c("Species", "Number of Genes") -->
+colnames(all_same_species) <- c("Species", "# of M. calida genes", "# of M. gaviniae genes")
 
-<!-- same_spp <- mutate(same_spp, -->
-<!--                    Species = case_when(Species == "E_amylovora" ~ "E. amylovora", -->
-<!--                                        Species == "E_tasmaniensis" ~ "E. tasmaniensis", -->
-<!--                                        Species == "P_agglomerans" ~ "P. agglomerans", -->
-<!--                                        Species == "P_septica" ~ "P. septica", -->
-<!--                                        Species == "T_ptyseos" ~ "T. ptyseos", -->
-<!--                                        Species == "T_saanichensis" ~ "T. saanichensis", -->
-<!--                                        Species == "E_cloacae" ~ "E. cloacae")) %>% -->
-<!--   mutate(Species = factor(Species, levels = c("P. agglomerans", "P. septica", "E. amylovora", "E. tasmaniensis", -->
-<!--                                               "T. ptyseos", "T. saanichensis", "E. cloacae"), ordered = TRUE)) -->
+all_same_species$Species <- gsub("_", ". ", all_same_species$Species)
 
-<!-- same_spp <- arrange(same_spp, Species) -->
+kable(all_same_species, 
+      caption = "Table 3. The number of *M. calida* and *M. gaviniae* genes wherein all analyses gave the same species as the closest relative.")
+```
 
-<!-- kable(same_spp, caption = "Table 5. The number of genes that had each species as the closest or most similar relative for all four analyses.") -->
+```{r Table4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+CR_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_calida.csv", 
+                      stringsAsFactors = FALSE)
 
-<!-- # table(stats_on_all$Same_Genus) -->
-<!-- # table(stats_on_all$NT_Same) -->
-<!-- # table(stats_on_all$AA_Same) -->
-<!-- # table(stats_on_all$Rel_Same) -->
-<!-- # table(stats_on_all$Ide_Same) -->
-<!-- ``` -->
+CR_gaviniae <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_AllResults_gaviniae.csv", 
+                        stringsAsFactors = FALSE)
 
-<!-- ```{r Figure8, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 8. Sample for first 100 genes.", fig.width=6.5, fig.height=3} -->
-<!-- all_results_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureN_AllResultsCalida.csv", -->
-<!--                                stringsAsFactors = TRUE) %>% -->
-<!--   pivot_longer(cols = c(Rel_NT:Ide_AA), names_to = "Analysis", values_to = "Closest_Similar") %>% -->
-<!--   mutate(Closest_Similar = factor(Closest_Similar, -->
-<!--                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", -->
-<!--                                              "E_cloacae", "P_syringiae"), -->
-<!--                                   ordered = TRUE), -->
-<!--          Analysis = factor(Analysis, levels = c("Ide_AA", "Rel_AA", "Ide_NT", "Rel_NT"), ordered = TRUE), -->
-<!--          Gene_Name = reorder(Gene_Name, ID)) -->
+cal_genus <- subset(CR_calida, Genus != "Nope")
+gav_genus <- subset(CR_gaviniae, Genus != "Nope")
 
-<!-- ggplot(all_results_calida[1:400, ], aes(x = Gene_Name, y = Analysis, fill = Closest_Similar)) + -->
-<!--   geom_tile() + -->
-<!--   scale_fill_brewer(palette = "Paired", -->
-<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
-<!--                                "P. syringae")) + -->
-<!--   scale_y_discrete(labels = c("AA Identity", "AA Closest", "NT Identity", "NT Closest")) + -->
-<!--   labs(fill = "Closest Relative or \n Most Similar Taxa") + -->
-<!--   theme(legend.text = element_text(face = "italic"), -->
-<!--         # legend.position = "bottom", -->
-<!--         text = element_text(size = 12,  family = "Times New Roman"), -->
-<!--         axis.text.x = element_text(angle = 45, hjust = 1))               # Rotates x axis labels -->
+all_same_genus <- data.frame(table(cal_genus$Genus),
+                             table(gav_genus$Genus))
 
-<!-- stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE) -->
+all_same_genus <- subset(all_same_genus, select = -Var1.1)
+colnames(all_same_genus) <- c("Genus", "# of M. calida genes", "# of M. gaviniae genes")
 
-<!-- # table(stats_on_all$Same_Species) -->
-<!-- # table(stats_on_all$Same_Genus) -->
-<!-- # table(stats_on_all$NT_Same) -->
-<!-- # table(stats_on_all$AA_Same) -->
-<!-- # table(stats_on_all$Rel_Same) -->
-<!-- # table(stats_on_all$Ide_Same) -->
-<!-- ``` -->
+all_same_genus <- all_same_genus %>%
+  mutate(Genus = factor(Genus, 
+                       levels = c("Pantoea", "Erwinia", "Tatumella", "Enterobacter"), ordered = TRUE)) %>%
+  arrange(Genus)
 
-<!-- ```{r Figure9, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 9. Maximum likelihood gene trees for ttuB for nucleotide (A) and amino acid sequences (B).", fig.width=3.25} -->
-<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_NT.png") -->
-<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_ttuB_AA.png") -->
-<!-- ``` -->
+test <- data.frame(Gene = CR_calida$Gene, cal = CR_calida$Genus, gav = CR_gaviniae$Genus)
+test$Same <- case_when(test$cal == test$gav ~ test$cal,
+                       TRUE ~ "Nah")
+test <- subset(test, Same == "Nah")
+# table(test$Same)
+# FALSE  TRUE 
+# 39     760
 
-<!-- ```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10.
-Maximum likelihood gene trees for secY for nucleotide (A) and amino acid sequences (B).
-Since bifucations could not be seen in radial trees, condensed trees were computed with a 70% cutoff for both the nucleotide sequnces (C) and amino acid sequences (D).", fig.width=3.25} -->
-<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_NT.png") -->
-<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10_secY_AA.png") -->
-<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_NT70.png") -->
-<!-- knitr::include_graphics("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10b_secY_AA70.png") -->
-<!-- ``` -->
+kable(all_same_genus, 
+      caption = "Table 4. The number of *M. calida* and *M. gaviniae* genes wherein all analyses gave the same genus as the closest relative.")
+```
 

### 2021-04-24 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -156,17 +156,34 @@ table(test_gav$CR_test)
 # FALSE  TRUE 
 # 178     621 
 
-### HERE!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-test <- data.frame(Gene = test_cal$PD_Gene, 
-                   cal_PD = test_cal$PD_CR, 
-                   cal_GD = test_cal$GD_CR2, 
-                   gav_PD = test_gav$PD_CR, 
-                   gav_GD = test_gav$GD_CR2)
-test <- mutate(test,
-               cal = cal_PD == cal_GD,
-               gav = gav_PD == gav_GD,
-               PD = cal_PD == gav_PD,
-               GD = cal_GD == gav_GD)
+#
+all <- data.frame(Gene = GD_calida$gene,
+                  GD_cal = GD_calida$Closest_Relative,
+                  GD_gav = GD_gaviniae$Closest_Relative,
+                  PD_cal = PD_calida$Closest_Relative,
+                  PD_gav = PD_gaviniae$Closest_Relative)
+
+all <- mutate(all,
+              GD_cal = case_when(GD_cal == "Pantoea_sept" ~ "P_septica",
+                                 GD_cal == "Erwinia_tasm" ~ "E_tasmaniensis",
+                                 GD_cal == "Pantoea_aggl" ~ "P_agglomerans",
+                                 GD_cal == "Erwinia_amyl" ~ "E_amylovora",
+                                 GD_cal == "Tatumella_sa" ~ "T_saanichensis",
+                                 GD_cal == "Tatumella_pt" ~ "T_ptyseos",
+                                 GD_cal == "Enterobacter" ~ "E_cloacae"),
+              GD_gav = case_when(GD_gav == "Pantoea_sept" ~ "P_septica",
+                                 GD_gav == "Erwinia_tasm" ~ "E_tasmaniensis",
+                                 GD_gav == "Pantoea_aggl" ~ "P_agglomerans",
+                                 GD_gav == "Erwinia_amyl" ~ "E_amylovora",
+                                 GD_gav == "Tatumella_sa" ~ "T_saanichensis",
+                                 GD_gav == "Tatumella_pt" ~ "T_ptyseos",
+                                 GD_gav == "Enterobacter" ~ "E_cloacae"),)
+
+all <- mutate(all,
+              cal = GD_cal == PD_cal,
[…]

### 2021-04-24 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 The bars’ lengths and colours correspond to the species that was the closest relative to the *Mixta* gene at that position.
-The results between the two *Mixta* species for the closest relatives were the same except for one and two genes for the NTS and the AAS, respectively; therefore, only the results for *M.
+The results between the two *Mixta* species for the closest relatives were the same except for one (*ttuB*) and two (*ttuB* and *secY*) genes for the NTS and the AAS, respectively; therefore, only the results for *M.
 calida* are shown.

### 2021-04-25 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -864,42 +864,35 @@ all <- data.frame(Gene = NT_dist$Gene,
                   Distance_NTS = NT_dist$Closest_Relative,
                   Distance_AAS = AA_dist$Closest_Relative,
                   Identity_NTS = NT_iden$Closest_Relative,
-                  Identity_AAS = AA_iden$Closest_Relative); # rm(AA_dist, AA_iden, NT_dist, NT_iden)
+                  Identity_AAS = AA_iden$Closest_Relative); rm(AA_dist, AA_iden, NT_dist, NT_iden)
 
 all <- all %>%
   mutate(Agreement = case_when(Distance_NTS == Distance_AAS &
                                  Distance_NTS == Identity_NTS &
-                                 Distance_NTS == Identity_AAS ~ "Yes",
+                                 Distance_NTS == Identity_AAS ~ Distance_NTS,
                                TRUE ~ "No")) %>%
-  pivot_longer(cols = c(Distance_NTS:Agreement),
-               names_to = "Analysis",
-               values_to = "Closest_Relative") %>%
-  mutate(Gene = substring(Gene, first = 7),
-         Analysis = factor(Analysis, levels = c("Identity_AAS", "Identity_NTS", "Distance_AAS", "Distance_NTS", "Agreement"), ordered = TRUE),
-         Closest_Relative = factor(Closest_Relative, levels = c(Spp, "Yes", "No"), ordered = TRUE))
+  subset(Agreement != "No") %>%
+  mutate(Gene_Name = substring(Gene, first = 7),
+         Agreement = factor(Agreement, levels = Spp, ordered = TRUE),
+         ID = as.numeric(ID))
+
+all$Gene <- reorder(all$Gene, all$ID)
 
 all <- all[order(all$ID),]
 
-my_colours <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F",
-                "#000000", "#FFFFFF")
-
-my_labels <- c(expression(italic("P. septica")),
-               expression(italic("P. agglomerans")),
-               expression(italic("E. tasmaniensis")),
-               expression(italic("E. amylovora")),
-               expression(italic("T. ptyseos")),
-               expression(italic("T. saanichensis")),
-               expression(italic("E. cloacae")),
-               "Yes", "No")
-
-expression( italic(p~value) == 0.01 )
-
-ggplot(all, aes(x = Gene, y = Analysis, fill = Closest_Relative)) +
[…]

### 2021-04-25 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
+```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 5.
+The closest relatives to *M.
+calida* (**A**) and *M.
+gaviniae (**B**) genes that were agreed upon by all four analyses that were conducted in this study.
+The circular plot simulates the circular chromosome of the respective species.
+The colour of the bars represent the different species to which the gene is most closely related.
+The *M.
+calida* and *M.
+gaviniae* genomes have 4092 and 4242 genes, respectively.", fig.width=6.5, fig.height=3.9}
+all_cal <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6A_AllAgreement_calida.csv", stringsAsFactors = FALSE) %>%
+  mutate(Agreement = factor(Agreement, levels = Spp, ordered = TRUE))
+
+all_gav <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6B_AllAgreement_gaviniae.csv", stringsAsFactors = FALSE) %>%
+  mutate(Agreement = factor(Agreement, levels = Spp, ordered = TRUE))
+
+calida <- ggplot(all_cal, aes(x = ID, y = 1, fill = Agreement)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 5) + # 
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 1), breaks = c(0, 1)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        panel.border = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+gaviniae <- ggplot(all_gav, aes(x = ID, y = 1, fill = Agreement)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 5) + # 
+  coord_polar() +
+  scale_fill_brewer(palette = "Paired", 
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
+                               "P. syringae")) +
+  scale_y_continuous(limits = c(0, 1), breaks = c(0, 1)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        panel.border = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+ggarrange(calida, gaviniae, 
+          labels = c("A", "B"),
+          font.label = list(family = "Times New Roman"),
+          label.x = 0.04,
+          label.y = 0.98,
+          common.legend = TRUE, 
+          legend = "bottom")
+```
+
 # Discussion

### 2021-04-25 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -841,7 +841,7 @@ kable(all_same_species,
       caption = "Table 4. The number of *M. calida* and *M. gaviniae* genes wherein all analyses gave the same genus as the closest relative.")
 # ```
 
-# Figure 6. All Results: M. calida -------------------------------------------------------------------------------------------------------------------------
+# Figure 6. All Results: M. calida and M. gaviniae ---------------------------------------------------------------------------------------------------------
 NT_dist <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE)
 AA_dist <- read.csv("9_Results_AA/four_relatives_calida_AA.csv", stringsAsFactors = FALSE)
 NT_iden <- read.csv("9_Results_NT/four_identities_calida_NT.csv", stringsAsFactors = FALSE) %>%
@@ -859,14 +859,14 @@ AA_iden <- read.csv("9_Results_AA/four_identities_calida_AA.csv", stringsAsFacto
          End = as.numeric(End),
          ID = as.numeric(ID))
 
-all <- data.frame(Gene = NT_dist$Gene,
-                  ID = NT_dist$ID,
-                  Distance_NTS = NT_dist$Closest_Relative,
-                  Distance_AAS = AA_dist$Closest_Relative,
-                  Identity_NTS = NT_iden$Closest_Relative,
-                  Identity_AAS = AA_iden$Closest_Relative); rm(AA_dist, AA_iden, NT_dist, NT_iden)
+all_cal <- data.frame(Gene = NT_dist$Gene,
+                      ID = NT_dist$ID,
+                      Distance_NTS = NT_dist$Closest_Relative,
+                      Distance_AAS = AA_dist$Closest_Relative,
+                      Identity_NTS = NT_iden$Closest_Relative,
+                      Identity_AAS = AA_iden$Closest_Relative); rm(AA_dist, AA_iden, NT_dist, NT_iden)
 
-all <- all %>%
+all_cal <- all_cal %>%
   mutate(Agreement = case_when(Distance_NTS == Distance_AAS &
                                  Distance_NTS == Identity_NTS &
                                  Distance_NTS == Identity_AAS ~ Distance_NTS,
@@ -876,17 +876,114 @@ all <- all %>%
[…]

### 2021-04-26 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

--- 2021-04-24 Manuscript/Manuscript_Paper.Rmd
@@
 
-**Keywords:**
+**Keywords:** *Mixta*, phylogenetic analyses, average nucleotide identity, average amino acid identity
 
@@
                                "P. syringae")) +
+  xlim(-19, 4092) +
   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
@@
 The four above analyses (genetic distance and site identity, both using NTS and AAS) gave slightly different results for the *Mixta* sequences.
-When comparing the results, approximately 38.7% of both the *M.
+The four analyses agreed upon a closest relative for approximately 38.7% of each of the *M.
 calida* and *M.
-gaviniae* sequences agreed upon a closest relative.
-However, the number of sequences closely related to any of the other taxa were slightly different between *M.
+gaviniae* sequences.
+The number of sequences that were closely related to one of the other taxa differed between the two *Mixta* species (Table 3).
+Additionally, although 38.7% of both *Mixta* species’ sequences had a consistent closest relative, 30 genes differed between them.
+For these 30 genes, there was a consensus for only one of the *Mixta* species and not the other.
+Nevertheless, 34.9% of the *M.
 calida* and *M.
-gaviniae* (Table 3).
-There were no genes for which all four analyses agreed on *T.
-ptyseos* being the closest relative (Table 3).
-Additionally, 7.5% of the *M.
-calida* and *M.
-gaviniae* homologs did not share the same closest relative.
+gaviniae* homologs agreed on the same species.
 
@@
 # table(CR_calida$ALL)
-# FALSE  TRUE         # FALSE: 61.3%
-# 490    309          # TRUE: 38.7% of M. calida genes had the same closest relative in all four analyses
+# E_amylovora      E_cloacae E_tasmaniensis             No  P_agglomerans      P_septica T_saanichensis 
+#           8              6             17            490             22            255              1 
+# FALSE  TRUE 
+# 490    309
 
@@
 # table(CR_gaviniae$ALL)
+# E_amylovora      E_cloacae E_tasmaniensis             No  P_agglomerans      P_septica T_saanichensis 
+#           7              7             20            490             22            252              1 
 # FALSE  TRUE 
@@
 
-all_same_cal <- subset(CR_calida, ALL == TRUE); all_same_gav <- subset(CR_gaviniae, ALL == TRUE)
+all_same_cal <- subset(CR_calida, ALL != "No"); all_same_gav <- subset(CR_gaviniae, ALL != "No")
 
-all_same_species <- data.frame(table(all_same_cal$NT_dist),
-                               table(all_same_gav$NT_dist)) %>%
+all_same_species <- data.frame(table(all_same_cal$ALL),
+                               table(all_same_gav$ALL)) %>%
   subset(select = -Var1.1) %>%
-  mutate(Var1 = factor(Var1, 
-                          levels = Spp, ordered = TRUE)) %>%
+  mutate(Var1 = factor(Var1, levels = Spp, ordered = TRUE)) %>%
   arrange(Var1)
@@
 ```
+
+Displaying these results in a circular plot showed that genes closely related to any of the taxa were dispersed across the *M.
+calida* and *M.
+gaviniae* chromosomes (Figures 6A and 6B, respectively).
+Given that similar patterns were found in the four analyses, genes with a consistent closest relative are randomly distributed across the genome and the position within the genome does not have an effect on whether the analyses agree (Figures 6A and 6B).
+
+```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6.
+The closest relatives to the *M.
+calida* (**A**) and *M.
+gaviniae (**B**) genes that were agreed upon by all four analyses that were conducted in this study.
+The circular plot simulates the circular chromosome of the respective species.
+The length and colour of the bars represent the different species to which the gene is most closely related.
+The *M.
+calida* and *M.
+gaviniae* genomes have 4092 and 4242 genes, respectively.", fig.width=6.5, fig.height=3.9}
+all_cal <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6A_AllAgreement_calida.csv", stringsAsFactors = FALSE) %>%
+  mutate(Agreement = factor(Agreement, levels = Spp, ordered = TRUE))
+
+all_gav <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6B_AllAgreement_gaviniae.csv", stringsAsFactors = FALSE) %>%
+  mutate(Agreement = factor(Agreement, levels = Spp, ordered = TRUE))
+
+my_colours <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", 
+                "#E31A1C", "#FDBF6F") # For P_septica, P_agglomerans, E_tasmaniensis, E_amylovora, T_saanichensis, and E_cloacae
+
+calida <- ggplot(all_cal, aes(x = ID, y = CR_Number, fill = Agreement)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_manual(values = my_colours,
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. saanichensis", "E. cloacae")) +
+  xlim(-19, 4092) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        panel.border = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+gaviniae <- ggplot(all_gav, aes(x = ID, y = CR_Number, fill = Agreement)) +
+  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
+  coord_polar() +
+  scale_fill_manual(values = my_colours,
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. saanichensis", "E. cloacae")) +
+  xlim(-19, 4242) +
+  scale_y_continuous(limits = c(0, 9), breaks = c(0, 9)) +
+  labs(fill = "Closest Relative") +
+  theme_bw() +
+  theme(legend.text = element_text(face = "italic"),
+        axis.title = element_blank(),
+        panel.border = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+ggarrange(calida, gaviniae, 
+          labels = c("A", "B"),
+          font.label = list(family = "Times New Roman"),
+          label.x = 0.04,
+          label.y = 0.98,
+          common.legend = TRUE, 
+          legend = "bottom")
+```
+
+Interestingly, the four analyses agreed on a genus being the closest relative for a total of 63.4 and 63.2% of the *M.
+calida* and *M.
+gaviniae* sequences, respectively (Table 4).
+Unexpectedly, only 39 of the *M.
+calida* and the *M.
+gaviniae* homologs did not agree on which genus was the closest relative.
+Of these genes, there was a consensus for one of *M.
+calida* or *M.
+gaviniae* and not the other for 38 homologs.
+For only one gene (*ttuB*) did the analyses agree on *Tatumella* for *M.
+calida* and *Pantoea* for *M.
+gaviniae.
 
@@
 
+## Phylogenetic Trees
+
 # Discussion

### 2021-04-26 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -345,12 +345,6 @@ FR_nucl_calida <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsA
                                      Closest_Relative == "T_ptyseos" ~ 7, Closest_Relative == "T_saanichensis" ~ 8,
                                      Closest_Relative == "E_cloacae" ~ 9, Closest_Relative == "P_syringae" ~ 0))
 
-extra_genes_NT <- data.frame(Gene = NA_character_, First = NA_character_, Second = NA_character_, Third = NA_character_, Fourth = NA_character_,
-                             Mixta_check = NA, Closest_Relative = "P_syringae", Gene_Length = NA_real_, Beg = NA_real_, End = NA_real_, 
-                             ID = as.numeric((max(FR_nucl_calida$ID) + 1):4092), Relative_Number = 0)
-
-FR_nucl_calida <- rbind(FR_nucl_calida, extra_genes_NT)                                   # Combines M_calida with the extra genes
-
 write.csv(FR_nucl_calida, "10_Tables_and_Figures/Figure3A_Circular_NT.csv", row.names = FALSE)
 
 FR_aa_calida <- read.csv("9_Results_AA/four_relatives_calida_AA.csv", stringsAsFactors = FALSE) %>%
@@ -358,12 +352,6 @@ FR_aa_calida <- read.csv("9_Results_AA/four_relatives_calida_AA.csv", stringsAsF
                                      Closest_Relative == "E_tasmaniensis" ~ 5, Closest_Relative == "E_amylovora" ~ 6,
                                      Closest_Relative == "T_ptyseos" ~ 7, Closest_Relative == "T_saanichensis" ~ 8,
                                      Closest_Relative == "E_cloacae" ~ 9, Closest_Relative == "P_syringae" ~ 0))
-
-extra_genes_AA <- data.frame(Gene = NA_character_, First = NA_character_, Second = NA_character_, Third = NA_character_, Fourth = NA_character_,
-                             Mixta_check = NA, Closest_Relative = "P_syringae", Gene_Length = NA_real_, Beg = NA_real_, End = NA_real_, 
-                             ID = as.numeric((max(FR_aa_calida$ID) + 1):4092), Relative_Number = 0)
-
[…]

### 2021-05-02 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -973,3 +973,72 @@ ggarrange(calida, gaviniae,
           common.legend = TRUE, 
           legend = "bottom")
 # ```
+
+# Phylogenetic Trees: Four Genes (MLSA) ------------------------------------------------------------------------------------------------------------------
+NT_dist <- read.csv("9_Results_NT/four_relatives_calida_NT.csv", stringsAsFactors = FALSE)
+AA_dist <- read.csv("9_Results_AA/four_relatives_calida_AA.csv", stringsAsFactors = FALSE)
+NT_iden <- read.csv("9_Results_NT/four_identities_calida_NT.csv", stringsAsFactors = FALSE) %>%
+  separate(col = Mixta, into = c("Species", "Gene_Length", "Beg", "End", "ID"), sep = "\\|") %>%
+  subset(select = -c(Species)) %>%
+  mutate(Gene_Length = as.numeric(Gene_Length),
+         Beg = as.numeric(Beg),
+         End = as.numeric(End),
+         ID = as.numeric(ID))
+AA_iden <- read.csv("9_Results_AA/four_identities_calida_AA.csv", stringsAsFactors = FALSE) %>%
+  separate(col = Mixta, into = c("Species", "Gene_Length", "Beg", "End", "ID"), sep = "\\|") %>%
+  subset(select = -c(Species)) %>%
+  mutate(Gene_Length = as.numeric(Gene_Length),
+         Beg = as.numeric(Beg),
+         End = as.numeric(End),
+         ID = as.numeric(ID))
+
+all_cal <- data.frame(Gene = NT_dist$Gene,
+                      ID = NT_dist$ID,
+                      Distance_NTS = NT_dist$Closest_Relative,
+                      Distance_AAS = AA_dist$Closest_Relative,
+                      Identity_NTS = NT_iden$Closest_Relative,
+                      Identity_AAS = AA_iden$Closest_Relative); rm(AA_dist, AA_iden, NT_dist, NT_iden)
+
+all_cal <- all_cal %>%
+  mutate(Agreement = case_when(Distance_NTS == Distance_AAS &
+                                 Distance_NTS == Identity_NTS &
+                                 Distance_NTS == Identity_AAS ~ Distance_NTS,
+                               TRUE ~ "No"))
+
+NT_dist_g <- read.csv("9_Results_NT/four_relatives_gaviniae_NT.csv", stringsAsFactors = FALSE)
[…]

### 2021-05-02 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 Alignments were performed in R as described above, and the aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@Wickham2019].
-A function was created using base `R` code to calculate the number of identical sites between each sequence and the *Mixta* homologs.
+A function was created using base `R` code to calculate the number of identical sites between each sequence and the *Mixta* homologues.
 Values that equal or are close to 100% are more identical to the *Mixta* sequences.
@@
 Site identity is the amount of similarity between two sequences and accounts for the sequence length.
-The non-*Mixta* sequences were aligned to each of the *Mixta* homologs, so the percent identity was relative to *Mixta*.
+The non-*Mixta* sequences were aligned to each of the *Mixta* homologues, so the percent identity was relative to *Mixta*.
 The closest relative to both *Mixta* species is defined as the taxon with the highest similarity to *Mixta*.
@@
 cloacae* were evenly dispersed across the *Mixta* chromosome for both the NTS and AAS (Figure 5A and 5B).
-Similar to the genetic distance results, genes that are labelled as being closely related to either of the *Tatumella* species are located only on right side of chromosome, and genes that are closely related to *E.
+Similar to the genetic distance results, genes that are labelled as closely related to either of the *Tatumella* species are located only on the right side of the chromosome, and genes closely related to *E.
 amylovora* are predominantly on the right side as well (Figure 5A).
-Unexpectedly, the lower left quadrant of the chromosome contains primarily genes that are closely related to either of the *Pantoea* species with only a few exceptions (Figure 5A).
+Unexpectedly, the lower left quadrant of the chromosome primarily contains genes closely related to either of the *Pantoea* species, with only a few exceptions (Figure 5A).
 In the AAS analysis, genes closely related to any of the other taxa were more evenly distributed across the genome (Figure 5B).
@@
 The four above analyses (genetic distance and site identity, both using NTS and AAS) gave slightly different results for the *Mixta* sequences.
-The four analyses agreed upon a closest relative for approximately 38.7% of each of the *M.
+The four analyses agreed upon the closest relative for approximately 38.7% of each of the *M.
 calida* and *M.
 gaviniae* sequences.
-The number of sequences that were closely related to one of the other taxa differed between the two *Mixta* species (Table 3).
+The number of sequences closely related to one of the other taxa differed between the two *Mixta* species (Table 3).
 Additionally, although 38.7% of both *Mixta* species’ sequences had a consistent closest relative, 30 genes differed between them.
@@
 calida* and *M.
-gaviniae* homologs agreed on the same species.
+gaviniae* homologues agreed on the same species.
 
@@
 gaviniae* chromosomes (Figures 6A and 6B, respectively).
-Given that similar patterns were found in the four analyses, genes with a consistent closest relative are randomly distributed across the genome and the position within the genome does not have an effect on whether the analyses agree (Figures 6A and 6B).
+Given that similar patterns were found in the four analyses, genes with a consistent closest relative are randomly distributed across the genome.
+The position within the genome does not affect whether the analyses agree (Figures 6A and 6B).
 
@@
 
-my_colours <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", 
-                "#E31A1C", "#FDBF6F") # For P_septica, P_agglomerans, E_tasmaniensis, E_amylovora, T_saanichensis, and E_cloacae
+my_colours <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", # For P_septica, P_agglomerans, E_tasmaniensis, E_amylovora,
+                "#E31A1C", "#FDBF6F") # T_saanichensis, and E_cloacae
 
@@
 calida* and the *M.
-gaviniae* homologs did not agree on which genus was the closest relative.
+gaviniae* homologues did not agree on which genus was the closest relative.
 Of these genes, there was a consensus for one of *M.
 calida* or *M.
-gaviniae* and not the other for 38 homologs.
-For only one gene (*ttuB*) did the analyses agree on *Tatumella* for *M.
+gaviniae* and not the other for 38 homologues.
+Only one gene (*ttuB*) did the analyses agree on *Tatumella* for *M.
 calida* and *Pantoea* for *M.
@@
 
-## Phylogenetic Trees
+## Phylogenetic Gene Trees for *secY* and *ttuB*
+In the genetic distance analysis, only one (*ttuB*) and two (*secY* and *ttuB*) genes differed between the two *Mixta* species for the NTS and AAS, respectively.
+The phylogenetic trees for the *secY* gene show that the gene is well-conserved across the *Erwiniaceae* taxa (Figures 7A and 7B).
+When using the NTS, the *Mixta* species are grouped together and fall outside of the other three *Erwiniaceae* taxa (Figure 7A).
+When using the AAS, the *Mixta* species do not group together (Figure 7B).
+Additionally, *M.
+calida* has a slightly shorter branch length to the *Erwinia* species, whereas *M.
+gaviniae* has a somewhat shorter branch length to *P.
+septica* (Figure 7B).
+However, the nodes separating the *Mixta* species from the other taxa are not well-supported (Figures 7A and 7B).
+Unexpectedly, *P.
+septica*’s homologue of *secY* is most closely related to *E.
+cloacae* for the AAS (Figure 7B) whereas *E.
+cloacae* falls outside of the *Erwiniaceae* for the NTS (Figure 7A).
+
+![Figure 7.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringiae* as outgroups.
+The trees were reconstructed from the nucleotide (**A**) and amino acid (**B**) sequences of the gene *secY*.
+HKY+G and LG+G models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_secY_Trees.png)
+
+The phylogenetic trees for the *ttuB* gene show a clearer separation between the two *Mixta* homologues (Figures 8A and 8B).
+For both the NTS and AAS, *M.
+calida* was more closely related to *Tatumella* followed by *E.
+cloacae* and *M.
+gaviniae* was more closely related to *Pantoea* followed by *Erwinia* (Figures 8A and 8B).
+Unlike with *secY*, many of the nodes are well-supported (Figures 8A and 8B).
+
+![Figure 8.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringiae* as outgroups.
+The trees were reconstructed from the nucleotide (**A**) and amino acid (**B**) sequences of the gene *ttuB*.
+GTR+G and LG+G models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure8_ttuB_Tree.png)
+
+## Phylogenetic Trees from MLSA Genes
+Phylogenetic analyses often use several concatenated essential gene sequences to infer relatedness, namely *atpD*, *gyrB*, *infB*, and *rpoB*.
+However, *atpD* and *gyrB* did not meet the length requirements during the filtering process.
+Therefore, genes encoding the other subunits of ATP synthase (*atp*), translation initiation factor IF (*inf*), and DNA-directed RNA polymerase (*rpo*) were concatenated to create phylogenetic trees.
+The resulting trees show that *Mixta* are more closely related to *Pantoea* followed by *Erwinia* (Figures 9A and 9B).
+With the NTS, a well-supported bifurcation occurs between *Mixta* and *Pantoea* (Figure 9A).
+However, this does not occur with the AAS (Figure 9B).
+Instead, a somewhat-supported bifurcation occurs first between *Pantoea* and *Erwinia* and then a little-supported bifurcation between these two and *Mixta* (Figure 9B).
+
+![Figure 9.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringiae* as outgroups.
+The trees were reconstructed from the concantenated nucleotide (**A**) and amino acid (**B**) sequences of the genes *atpA*, *atpB*, *atpC*, *atpE*, *atpF*, *atpG*, *atpH*, *infA*, *infB*, *rpoA*, *rpoB*, *rpoC*, *rpoD*, and *rpoH*.
+GTR+G+I and LG+G+F models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+The nucleotide model allowed for some sites to be evolutionary invariable.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure9_Four_Tree.png)
+
+The four analyses conducted in this study agreed on a closest relative for only four of these genes – *atpA*, atpG*, *rpoB*, and *rpoC* – for both *M.
+calida* and *M.
+gaviniae*.
+This closest relative was *P.
+septica* for these genes.
+The other genes were more closely related to *P.
+agglomerans*, *E.
+tasmaniensis*, *E.
+amylovora*, and *E.
+cloacae* and these varied amongst the analyses.
+The gene *rpoD* was the only gene that was said to be closely related to *E.
+cloacae* by one of the analyses.
+The NTS genetic distance analysis labelled both the *Mixta* species as being closely related to *E.
+cloacae* and the remaining three analyses labelled this gene as being closely related to *P.
+septica*.
+
+## Phylogenetic Trees Showing Other Results
 

### 2021-05-02 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -1042,3 +1042,8 @@ four_cal <- rbind(filter(all_cal, grepl("atp", Gene)),
 four_gav <- rbind(filter(all_gav, grepl("atp", Gene)),
                   filter(all_gav, grepl("inf", Gene)),
                   filter(all_gav, grepl("rpo", Gene)))
+
+cal <- pivot_longer(four_cal, cols = c(Distance_NTS:Identity_AAS), names_to = "Analysis", values_to = "Species")
+
+ggplot(cal, aes(Analysis, Gene)) +
+  geom_tile(aes(fill = z), colour = "grey50")

### 2021-05-02 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
-The four analyses conducted in this study agreed on a closest relative for only four of these genes – *atpA*, atpG*, *rpoB*, and *rpoC* – for both *M.
+The four analyses conducted in this study agreed on a closest relative for only four of these genes – *atpA*, *atpG*, *rpoB*, and *rpoC* – for both *M.
 calida* and *M.
-gaviniae*.
+gaviniae* (Figures 10A and 10B).
 This closest relative was *P.
@@
 amylovora*, and *E.
-cloacae* and these varied amongst the analyses.
-The gene *rpoD* was the only gene that was said to be closely related to *E.
-cloacae* by one of the analyses.
-The NTS genetic distance analysis labelled both the *Mixta* species as being closely related to *E.
+cloacae* and these varied amongst the analyses (Figures 10A and 10B).
+One only of the genes, *rpoD*, was said to be closely related to *E.
+cloacae* by one of the analyses; the NTS genetic distance analysis labelled both the *Mixta* species as being closely related to *E.
 cloacae* and the remaining three analyses labelled this gene as being closely related to *P.
-septica*.
+septica* (Figures 10A and 10B).
+
+```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10.
+The closest relatives to the *M.
+calida* (**A**) and *M.
+gaviniae (**B**) MLSA genes for each of the four analyses.
+The colour of the tiles represent the different species to which the gene is most closely related.", fig.width=6.5}
+four_cal <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10A_PhyloTreeFour_calida.csv", 
+                     stringsAsFactors = FALSE) %>%
+  pivot_longer(cols = Distance_NTS:Identity_AAS, names_to = "Analysis", values_to = "Species") %>%
+  mutate(Gene = substring(Gene, first = 7),
+         Analysis = factor(Analysis, levels = c("Distance_NTS", "Distance_AAS", "Identity_NTS", "Identity_AAS"), ordered = TRUE),
+         Species = factor(Species, levels = Spp, ordered = TRUE))
+
+four_gav <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure10B_PhyloTreeFour_gaviniae.csv", 
+                     stringsAsFactors = FALSE) %>%
+  pivot_longer(cols = Distance_NTS:Identity_AAS, names_to = "Analysis", values_to = "Species") %>%
+  mutate(Gene = substring(Gene, first = 7),
+         Analysis = factor(Analysis, levels = c("Distance_NTS", "Distance_AAS", "Identity_NTS", "Identity_AAS"), ordered = TRUE),
+         Species = factor(Species, levels = Spp, ordered = TRUE))
+
+my_colours2 <- c("#A6CEE3", "#1F78B4",  # P. septica and P. agglomerans
+                 "#B2DF8A", "#33A02C",  # E. tasmaniensis and E. amylovora
+                 "#FDBF6F")             # E. cloacae
+
+cal <- ggplot(four_cal, aes(x = Analysis, y = Gene, fill = Species)) +
+  geom_tile() +
+  scale_fill_manual(values = my_colours2,
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "E. cloacae")) +
+  labs(fill = "Closest Relative") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.text.y = element_text(face = "italic"),
+        legend.text = element_text(face = "italic"),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+gav <- ggplot(four_gav, aes(x = Analysis, y = Gene, fill = Species)) +
+  geom_tile() +
+  scale_fill_manual(values = my_colours2,
+                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "E. cloacae")) +
+  labs(fill = "Closest Relative") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.text.y = element_text(face = "italic"),
+        legend.text = element_text(face = "italic"),
+        text = element_text(size = 12,  family = "Times New Roman"))
+
+ggarrange(cal, gav, 
+          labels = c("A", "B"),
+          font.label = list(family = "Times New Roman"),
+          common.legend = TRUE, 
+          legend = "right")
+```
 

### 2021-05-02 6_Results_Analysis.R
Update 6_Results_Analysis.R

@@ -25,6 +25,9 @@ count_relatives <- function(data) {
 
 Spp <- c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", 
          "E_cloacae", "P_syringae")
+
+colours <- data.frame(col = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00"),
+                      spp = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", "E_cloacae", "P_syringae"))
 #
 # Table 1. Genomes ----------------------------------------------------------------------------------------------------------------------------------------
 strain <- data.frame(Species = c("Mixta calida", "Mixta gaviniae", "Pantoea agglomerans", "Pantoea septica", "Erwinia amylovora",
@@ -1038,12 +1041,59 @@ all_gav <- all_gav %>%
 four_cal <- rbind(filter(all_cal, grepl("atp", Gene)),
                   filter(all_cal, grepl("inf", Gene)),
                   filter(all_cal, grepl("rpo", Gene)))
+
+write.csv(four_cal, "10_Tables_and_Figures/Figure10A_PhyloTreeFour_calida.csv", row.names = FALSE)
   
 four_gav <- rbind(filter(all_gav, grepl("atp", Gene)),
                   filter(all_gav, grepl("inf", Gene)),
                   filter(all_gav, grepl("rpo", Gene)))
 
-cal <- pivot_longer(four_cal, cols = c(Distance_NTS:Identity_AAS), names_to = "Analysis", values_to = "Species")
+write.csv(four_gav, "10_Tables_and_Figures/Figure10B_PhyloTreeFour_gaviniae.csv", row.names = FALSE)
 
-ggplot(cal, aes(Analysis, Gene)) +
-  geom_tile(aes(fill = z), colour = "grey50")
+# ```{r Figure10, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 10. The closest relatives to the *M. calida* (**A**) and *M. gaviniae (**B**) 
+# MLSA genes for each of the four analyses. The colour of the tiles represent the different species to which the gene is most closely related.", 
+# fig.width=6.5}
[…]

### 2021-05-03 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
-<!-- \newpage -->
+\newpage
 # Abstract
-<!-- *Mixta* is a recently described genus within the bacterial family *Erwiniaceae* and is closely related to *Pantoea*, *Erwinia*, and *Tatumella*.
+*Mixta* is a recently described genus within the bacterial family *Erwiniaceae* and is closely related to *Pantoea*, *Erwinia*, and *Tatumella*.
 Depending on the genes and method used for phylogenetic analysis, the genus with which *Mixta* shares the most recent ancestor differs.
-This study aimed to determine the cause behind these contentious results and identified genes that may have resulted from horizontal gene transfer.
+This study aimed to determine the cause behind these contentious results.
 Ten complete or partial genomes – two *Mixta* species, six species from other *Erwiniaceae* genera, and two outgroups – were retrieved from NCBI and annotated.
 Homologous genes were extracted yielding a dataset of 799 genes.
-The genes were aligned with the ClustalW algorithm, and MEGA-CC was used for model testing and computing phylogenetic trees for each gene for both nucleotide and amino acid sequences.
-Distance matrices were then extracted from these trees.
+The genes were aligned with the ClustalW algorithm, and MEGA-CC was used to calculate the most appropriate model, distance matrices, and phylogenetic trees for both nucleotide and amino acid sequences.
 Nucleotide and amino acid identity analyses were also done using the programming language `R`.
@@
 Furthermore, future research should consider that different phylogenetic analyses may provide different results.
--->
 
-# Intro and MM
-<!-- \newpage -->
-<!-- # Introduction -->
-<!-- The bacterial family *Erwiniaceae* includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms, including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, and *Kalamiella* [REF CRAIG].
+\newpage
+# Introduction
+The bacterial family *Erwiniaceae* includes diverse Gram-negative [@Winslow1917; @Prakash2015] organisms, including the genera *Mixta*, *Pantoea*, *Erwinia*, *Tatumella*, *Buchnera*, *Wigglesworthia*, *Phaseolibacter*, and *Kalamiella* [REF CRAIG].
 *Mixta* is a relatively newer genus first identified by Palmer et al.
@@
 The first two were previously named as *Pantoea* species.
--->
 
-<!-- *Mixta*'s position relative to other members of the *Erwiniaceae*, in particular to *Pantoea*, *Erwinia*, *Tatumella*, differs vastly between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
+*Mixta*'s position relative to other members of the *Erwiniaceae*, in particular to *Pantoea*, *Erwinia*, *Tatumella*, differs vastly between studies [@Prakash2015; @Palmer2017; @Palmer2018; @Adeolu2016; @Brady2011; @Rezzonico2016; @Brady2012; @Gueule2015].
 Prakash et al.
@@
 However, the ribosomal MLSA gene tree proposed *Pantoea* and *Tatumella* were sister taxa followed by *Erwinia* and then by *Mixta*.
--->
 
-<!-- What remains unclear is the relative contribution of disaparate evolutionary histories of *Mixta* genes, tree-building methods, and sequence evolution on the phylogenetic position of the genus.
+What remains unclear is the relative contribution of disaparate evolutionary histories of *Mixta* genes, tree-building methods, and sequence evolution on the phylogenetic position of the genus.
 This paper aims to determine the cause behind the *Mixta* species' contentious phylogenetic results.
@@
 Phylogenetic trees were constructed for combinations of different gene sequences.
--->
 
-<!-- # Materials and Methods -->
-<!-- ## Genomes and Gene Sets -->
-<!-- Ten publicly available full or partial genomes of the species' type strains were retrieved from the National Center for Biotechnology Information (NCBI).
+# Materials and Methods
+## Genomes and Gene Sets
+Ten publicly available full or partial genomes of the species' type strains were retrieved from the National Center for Biotechnology Information (NCBI).
 Two representative genomes were used for each of the four *Erwiniaceae* genera while *Enterobacter cloacae* (*Enterobacterales*) and *Pseudomonas syringae* (*Pseudomonaceae*) were used as outgroups (Table 1).
@@
 GET_HOMOLOGUES extracted 954 homologous genes between the ten species.
--->
 
-<!-- ```{r Table1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE} -->
-<!-- strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table1_Genomes.csv", stringsAsFactors = FALSE) -->
+```{r Table1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
+strain <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table1_Genomes.csv", stringsAsFactors = FALSE)
 
-<!-- colnames(strain)[3] <- "GenBank Accession No." -->
+colnames(strain)[3] <- "GenBank Accession No."
 
-<!-- kable(strain, caption = "Table 1. Genomes used in this study.") -->
-<!-- ``` -->
+kable(strain, caption = "Table 1. Genomes used in this study.")
+```
 
-<!-- In `R` [@RCore2019], two filter functions were written for the nucleotide sequence gene files.
+In `R` [@RCore2019], two filter functions were written for the nucleotide sequence gene files.
 The first filter kept genes files that contained ten, single-copy sequences that were approximately the same length.
@@
 Of the 954 homologous gene sets, 38 multi-copy genes were excluded, and another 175 were excluded for not meeting the length requirements.
--->
 
-<!-- A second filter was applied to these 175 genes to identify any genes wherein both representatives of each genera were approximately the same length.
+A second filter was applied to these 175 genes to identify any genes wherein both representatives of each genera were approximately the same length.
 The filter split the genes into two groups: sequences that met the original 90% length cutoff and sequences that were between 80 and 90% the length of the longest sequence.
@@
 The amino acid sequences for these genes were subsequently downloaded.
--->
 
-<!-- ## Gene Alignment -->
-<!-- Gene alignments were done in `R` using a ClustalW algorithm.
+## Gene Alignment
+Gene alignments were done in `R` using a ClustalW algorithm.
 The fasta files were read into the session using the `readDNAStringSet()` function (package: Biostrings) [@Pages2020].
 The function called to do the alignments was the `msaClustalW()` function (package: msa) [@Bodenhofer2015] using default parameters and 100 iterations.
-The aligned sequences were then converted into a comma-separated values (CSV)-compatible format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007] -->
+The aligned sequences were then converted into a comma-separated values (CSV)-compatible format using `msaConvert()` (package: msa)[@Bodenhofer2015] and written into a new fasta file using the `write.fasta()` function (package: seqinr) [@Charif2007]
 
-<!-- ## Model Selection -->
-<!-- The best model of sequence evolution was identified for each nucleotide and amino acid alignment using Model Selection as implemented in MEGA-CC [@Kumar2012].
+## Model Selection
+The best model of sequence evolution was identified for each nucleotide and amino acid alignment using Model Selection as implemented in MEGA-CC [@Kumar2012].
 Default parameters were used for model testing.
@@
 These CSV files were read into `R`, and the model for phylogenetic analysis was extracted based on the lowest Bayesian Information Criterion (BIC).
--->
 
-<!-- ## Phylogenetic Analysis -->
-<!-- Phylogenetic analysis was performed in MEGA-CC using maximum likelihood with 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
+## Phylogenetic Analysis
+Phylogenetic analysis was performed in MEGA-CC using maximum likelihood with 500 bootstrap replications, the model, and rate patterns to create a phylogenetic tree for each gene.
 If a model required a Gamma distribution, the number of discrete Gamma categories was set to 5, which is the default.
--->
 
-<!-- Distance matrices were extracted from each of the phylogenetic trees and saved as Excel files.
+Distance matrices were extracted from each of the phylogenetic trees and saved as Excel files.
 These files were then read into the `R` session.
@@
 Tables were created in `R Markdown` using the `kable()` function (package: knitr) [@Xie2020] and figures were created using the `ggplot()` function (package: ggplot2) [@Wickham2016].
--->
 
-<!-- ## Nucleotide and Amino Acid Identity -->
-<!-- Each of the non-*Mixta* sequences were aligned to one of the corresponding *Mixta* sequences in turn.
+## Nucleotide and Amino Acid Identity
+Each of the non-*Mixta* sequences were aligned to one of the corresponding *Mixta* sequences in turn.
 Alignments were performed in R as described above, and the aligned sequences were then split into the individual sites using the `str_split()` function (package: stringr) [@Wickham2019].
@@
 Values that equal or are close to 100% are more identical to the *Mixta* sequences.
--->
 
@@
 In the genetic distance analysis, only one (*ttuB*) and two (*secY* and *ttuB*) genes differed between the two *Mixta* species for the NTS and AAS, respectively.
-The phylogenetic trees for the *secY* gene show that the gene is well-conserved across the *Erwiniaceae* taxa (Figures 7A and 7B).
-When using the NTS, the *Mixta* species are grouped together and fall outside of the other three *Erwiniaceae* taxa (Figure 7A).
-When using the AAS, the *Mixta* species do not group together (Figure 7B).
+The phylogenetic trees for the *secY* gene showed that the gene is well-conserved across the *Erwiniaceae* taxa (Figures 7A and 7B).
+When using the NTS, the *Mixta* species grouped together and fell outside of the other three *Erwiniaceae* taxa (Figure 7A).
+When using the AAS, the *Mixta* species did not group together (Figure 7B).
 Additionally, *M.
-calida* has a slightly shorter branch length to the *Erwinia* species, whereas *M.
-gaviniae* has a somewhat shorter branch length to *P.
+calida* had a slightly shorter branch length to the *Erwinia* species, whereas *M.
+gaviniae* had a somewhat shorter branch length to *P.
 septica* (Figure 7B).
-However, the nodes separating the *Mixta* species from the other taxa are not well-supported (Figures 7A and 7B).
+However, the nodes separating the *Mixta* species from the other taxa were not well-supported (Figures 7A and 7B).
 Unexpectedly, *P.
-septica*’s homologue of *secY* is most closely related to *E.
+septica*’s homologue of *secY* was most closely related to *E.
 cloacae* for the AAS (Figure 7B) whereas *E.
-cloacae* falls outside of the *Erwiniaceae* for the NTS (Figure 7A).
+cloacae* fell outside of the *Erwiniaceae* for the NTS (Figure 7A).
 
@@
 cloacae* and *P.
-syringiae* as outgroups.
+syringae* as outgroups.
 The trees were reconstructed from the nucleotide (**A**) and amino acid (**B**) sequences of the gene *secY*.
@@
 
-The phylogenetic trees for the *ttuB* gene show a clearer separation between the two *Mixta* homologues (Figures 8A and 8B).
+The phylogenetic trees for the *ttuB* gene showed a more apparent separation between the two *Mixta* homologues (Figures 8A and 8B).
 For both the NTS and AAS, *M.
@@
 gaviniae* was more closely related to *Pantoea* followed by *Erwinia* (Figures 8A and 8B).
-Unlike with *secY*, many of the nodes are well-supported (Figures 8A and 8B).
+Unlike with *secY*, many of the nodes were well-supported (Figures 8A and 8B).
 
@@
 cloacae* and *P.
-syringiae* as outgroups.
+syringae* as outgroups.
 The trees were reconstructed from the nucleotide (**A**) and amino acid (**B**) sequences of the gene *ttuB*.
@@
 Therefore, genes encoding the other subunits of ATP synthase (*atp*), translation initiation factor IF (*inf*), and DNA-directed RNA polymerase (*rpo*) were concatenated to create phylogenetic trees.
-The resulting trees show that *Mixta* are more closely related to *Pantoea* followed by *Erwinia* (Figures 9A and 9B).
-With the NTS, a well-supported bifurcation occurs between *Mixta* and *Pantoea* (Figure 9A).
-However, this does not occur with the AAS (Figure 9B).
-Instead, a somewhat-supported bifurcation occurs first between *Pantoea* and *Erwinia* and then a little-supported bifurcation between these two and *Mixta* (Figure 9B).
+The resulting trees showed that *Mixta* are more closely related to *Pantoea* followed by *Erwinia* (Figures 9A and 9B).
+With the NTS, a well-supported bifurcation occured between *Mixta* and *Pantoea* (Figure 9A).
+However, this did not happen with the AAS (Figure 9B).
+Instead, a somewhat-supported bifurcation occured first between *Pantoea* and *Erwinia* and then a weakly-supported bifurcation between these two and *Mixta* (Figure 9B).
 
@@
 cloacae* and *P.
-syringiae* as outgroups.
+syringae* as outgroups.
 The trees were reconstructed from the concantenated nucleotide (**A**) and amino acid (**B**) sequences of the genes *atpA*, *atpB*, *atpC*, *atpE*, *atpF*, *atpG*, *atpH*, *infA*, *infB*, *rpoA*, *rpoB*, *rpoC*, *rpoD*, and *rpoH*.
@@
 
-The four analyses conducted in this study agreed on a closest relative for only four of these genes – *atpA*, *atpG*, *rpoB*, and *rpoC* – for both *M.
+The four analyses conducted in this study agreed on the closest relative for only four of these genes – *atpA*, *atpG*, *rpoB*, and *rpoC* – for both *M.
 calida* and *M.
 gaviniae* (Figures 10A and 10B).
-This closest relative was *P.
-septica* for these genes.
+This closest relative to these genes were *P.
+septica*.
 The other genes were more closely related to *P.
@@
 cloacae* and these varied amongst the analyses (Figures 10A and 10B).
-One only of the genes, *rpoD*, was said to be closely related to *E.
+One only of the genes, *rpoD*, was closely related to *E.
 cloacae* by one of the analyses; the NTS genetic distance analysis labelled both the *Mixta* species as being closely related to *E.
@@
 ## Phylogenetic Trees Showing Other Results
+Several genes were found to be more closely related to taxa other than *P.
+septica* or *P.
+agglomerans*.
+Thus, gene sequences wherein the four analyses agreed upon each of the other taxa were concatenated to compute phylogenetic trees.
+There was a consensus for 32 genes that the closest relative was not either of the *Pantoea* species; 17 were more closely related to *E.
+tasmaniensis*, 8 to *E.
+amylovora*, 1 to *T.
+saanichensis*, and 6 to *E.
+cloacae*.
+
+The genes most closely related to *E.
+tasmaniensis* were *fabG*, *folK*, *greB*, *hemE*, *hscA_2*, *ibaG*, *proA*, *rppH*, *sbp*, *tusA*, *tusE*, *uppS*, *ycfH*, *yffB*, and 3 hypothetical proteins.
+The phylogenetic trees resulting from the concatenated NTS (Figure 11A) and AAS (Figure 11B) showed strong support for the *Mixta* species being more closely related to *Erwinia*.
+Interestingly, *E.
+cloacae* fell outside of the *Erwiniaceae* for only the AAS (Figure 11B).
+However, the bifurcations between *E.
+cloacae* and the other taxa are not well-supported for either sequence type (Figures 11A and 11B).
+
+![Figure 11.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringae* as outgroups.
+The trees were reconstructed from the concantenated nucleotide (**A**) and amino acid (**B**) sequences of the genes *fabG*, *folK*, *greB*, *hemE*, *hscA_2*, *ibaG*, *proA*, *rppH*, *sbp*, *tusA*, *tusE*, *uppS*, *ycfH*, *yffB*, and 3 hypothetical proteins.
+The four analyses in this study agreed on *E.
+tasmaniensis* as being the closest relative for these 17 genes.
+GTR+G+I and LG+G+F models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+The nucleotide model allowed for some sites to be evolutionary invariable.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure11_Etasmaniensis.png)
+
+The eight genes most closely related to *E.
+amylovora* were *acuI*, *coaBC*, *dapB*, *grpE*, *mtnC*, *rlmJ*, *ssuE*, and a hypothetical protein.
+Similar to the results for *E.
+tasmaniensis*, the bifurcations between *Mixta* and *Erwinia* are well-supported for both the NTS (Figure 12A) and AAS (Figure 12B).
+Additionally, *E.
+cloacae*’s position relative to *Mixta* and *Erwinia* is well-supported for the NTS, but its position is ambiguous for AAS (Figures 12A and 12B).
+
+![Figure 12.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringae* as outgroups.
+The trees were reconstructed from the concantenated nucleotide (**A**) and amino acid (**B**) sequences of the genes *acuI*, *coaBC*, *dapB*, *grpE*, *mtnC*, *rlmJ*, *ssuE*, and a hypothetical protein.
+The four analyses in this study agreed on *E.
+amylovora* as being the closest relative for these 8 genes.
+GTR+G+I and LG+G+F models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+The nucleotide model allowed for some sites to be evolutionary invariable.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure12_Eamylovora.png)
+
+Only one gene, *tsf*, was most closely related to *T.
+saanichensis*.
+Nevertheless, this relationship is well-supported for both the NTS and AAS (Figures 13A and 13B).
+Unexpectedly, the two *Tatumella* species were not grouped for the AAS (Figure 13B).
+
+![Figure 13.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringae* as outgroups.
+The trees were reconstructed from the concantenated nucleotide (**A**) and amino acid (**B**) sequences of the gene *tsf*.
+The four analyses in this study agreed on *T.
+saanichensis* as being the closest relative for this one gene.
+K2+G and LG+G models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure13_Tsaanichensis.png)
+
+The six genes most closely related to the outgroup species *E.
+cloacae* were *bioD1*, *cueR*, *hemN_2*, *iscS*, *rpsM*, and *yqjC*.
+The relationship between *Mixta* and *E.
+cloacae* is well-supported in the phylogenetic trees computed from the concatenated NTS (Figure 14A) and AAS (Figure 14B).
+Both of the trees showed similar results as the *Tatumella* species were the next closest relative to *Mixta* with the same main difference being weaker branch support for the AAS (Figures 14A and 14B).
+Additionally, *Pantoea* and *Erwinia* fell into their own group (Figures 14A and 14B).
+
+![Figure 14.
+Maximum-likelihood phylogenetic trees for members of the *Erwiniaceae* with *E.
+cloacae* and *P.
+syringae* as outgroups.
+The trees were reconstructed from the concantenated nucleotide (**A**) and amino acid (**B**) sequences of the genes *bioD1*, *cueR*, *hemN_2*, *iscS*, *rpsM*, and *yqjC*.
+The four analyses in this study agreed on *E.
+cloacae* as being the closest relative for these 6 genes.
+GTR+G+I and LG+G+F models were used for the nucleotide and amino acid sequences, respectively.
+Both trees required a discrete Gamma distribution (5 categories) to model the evolutionary rate differences among sites.
+The nucleotide model allowed for some sites to be evolutionary invariable.
+Branch support was inferred from 500 bootstrap replicates; the percentage of trees in which the associated taxa clustered together are adjacent to the branches.
+The tree is drawn to scale, with branch lengths measured in the number of substitutions per site.
+Evolutionary analyses were conducted in MEGAX.](C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure14_Ecloacae.png)
 

### 2021-05-04 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
-Kimberly N. Hinz, John Stavrinides^1^
+Kim N. Hinz, John Stavrinides^1^
 
@@
 # Discussion
+Phylogenetic software programs vary on which models and parameters they have available, and these may vary within a program depending on the chosen method.
+MEGAX [@Kumar2018], for example, offers a wide range of models and additional parameters for the building of phylogenetic trees; however, the options are reduced for genetic distance calculation.
+A vital practise of selecting the most accurate model for your data is by conducting model testing.
+Model testing provides a list of models with additional parameters in the order of which model explains the most variation in the sequences’ differences.
+Selecting the most accurate model is always recommended; however, it is not always possible.
+My results show that slight differences in the taxa’s relationships occur when the more precise model is not used.
+For some genes, the phylogenetic signal of which taxon was the closest relative to *Mixta* was strong enough despite inaccuracies in the model.
+Nonetheless, there is a risk that the wrong species is labelled as the closest relative, and this poses problems if that species belongs to a separate genus or order.
+The species of interest may be wrongfully defined as belonging to a specific genus or order, or the genes may be inappropriately labelled as results from horizontal gene transfer (HGT).
 
-# Acknowledgements
+Additionally, the chosen method may contribute to contentious results.
+In order to use the most accurate model, I had to first run a phylogenetic analysis and then extract the genetic distance matrices.
+For the best available model for genetic distance, I ran the genetic distance analysis directly and thus, forwent the phylogenetic tree step.
+Even when the most accurate and the best available models were the same, the results for the closest relative were different.
+This suggests that the two methods calculate taxa relatedness differently despite using the same models and parameters.
+One explanation is the Gamma parameter (+G).
+The genetic distance method uses a continuous distribution in which 1.00 is the default, while the phylogenetic tree method uses a discrete Gamma parameter set to 5 categories.
+Therefore, diligence must be given when selecting software programs, methods, and models.
+
+Phylogenetic analysis is an accepted method for understanding the evolutionary relatedness [@Boundless2021] while site identity is used to confirm identification and taxon boundaries [@Ciufo2018].
+These two methods were conducted in this study as both nucleotide (NTS) and amino acid (AAS) sequences.
+*P.
+septica* was the closest relative to the majority of *Mixta* homologues in all four analyses.
+This lends strong support that *Mixta* and *P.
+septica* share a recent common ancestor, followed by other *Pantoea* species, *Erwinia*, and then *Tatumella*.
+This agrees with *Mixta* being previously known as part *Pantoea*.
+Additionally, there were no *Mixta* sequences that were closely related to *P.
+syringae* which demonstrates that this species is a suitable outgroup.
+
+For both methods, fewer genes were labelled as being closely related to *P.
+septica* when using the AAS versus the NTS.
+This difference may be to a loss of phylogenetic signal when using amino acid sequences caused by codon degeneracy.
+Nevertheless, sequence type, as well as different evolutionary models between NTS and AAS, affects the results.
+AAS may be better suited for uncovering taxonomic relationships at the genus or higher level and NTS may be more appropriate for the species level.
+
+When considering the spatial effects of the closest relatives, the uniform pattern of the genes related to *P.
+septica* further lends support that this species and *Mixta* share a recent common ancestor.
+Additionally, the other taxa followed the same pattern.
+If the genes that were closely related to non-*Pantoea* species are a result of HGT, then this suggests that HGT may be a common occurrence within the *Erwiniaceae*.
+A second explanation for the even distribution patterns in the closest relatives is that the genes are highly conserved.
+Two or more species may have been tied for closest relative and the R program may have selected one based another characteristic, such as alphabetical order.
+Nevertheless, further research into these genes and the frequency of HGT events with *Mixta* and other genera within the *Erwiniaceae* is needed.
 

### 2021-07-22 Extra.R
Create Extra.R
Extra code to answer John's questions

@@ -1,0 +1,294 @@
+### M. calida Distance/Identity #############################################################################################################################
+library("readxl")
+library("tidyverse")
+
+setwd("C:/Users/Kim/OneDrive/Undergraduate_Research/Biology_396/")
+
+distance <- read.csv("9_Results_NT/distance_calida_NT.csv")
+
+distance$GeneCode = substring(distance$Gene, first = 0, last = 6)
+
+genes <- unique(distance$GeneCode)
+spp <- c("M_calida", "M_gaviniae", "P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", "T_saanichensis", "E_cloacae",
+         "P_syringae")
+
+distance <- mutate(distance,
+                   GeneCode = factor(GeneCode, levels = genes, ordered = TRUE),
+                   Species = factor(Species, levels = spp, ordered = TRUE))
+
+distance <- arrange(distance, Gene, Species)
+
+
+identity <- read.csv("identity_calida_NT.csv")
+identity$GeneCode = substring(identity$Gene, first = 0, last = 6)
+
+identity <- mutate(identity,
+                   GeneCode = factor(GeneCode, levels = genes, ordered = TRUE),
+                   Species = factor(Species, levels = spp, ordered = TRUE))
+
+identity <- arrange(identity, Gene, Species)
+
+test <- cbind(distance, identity)
+colnames(test) <- c("Spp_d", "Distance", "Gene_d", "Code_d", "Gene_i", "Mixta", "Spp_i", "Percent", "Code_i")
+test <- mutate(test,
+               Gene_Check = Code_d == Code_i,
+               Spp_Check = Spp_d == Spp_i)
+
+M_calida_nt <- data.frame(Mixta = "M_calida",
+                          Gene = test$Gene_i,
+                          Species = test$Spp_d,
+                          Genetic_Distance = test$Distance, 
+                          Percent_Identity = test$Percent)
+
+write.csv(M_calida_nt, "9_Results_NT/Mcalida_results.csv", row.names = FALSE)
+
+### M. gaviniae Distance/Identity ###########################################################################################################################
+distance <- read.csv("9_Results_NT/distance_gaviniae_NT.csv")
+
+distance$GeneCode = substring(distance$Gene, first = 0, last = 6)
+
+genes <- unique(distance$GeneCode)
[…]

### 2021-07-22 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

@@
 
-\newpage
 # Introduction
@@
 [@Prakash2015], and Gueule et al.
-[@Gueule2015] – placed the *Mixta* species within the *Pantoea* genera.
+[@Gueule2015] – placed the *Mixta* species within the *Pantoea* genus.
 In contrast, the analysis in Palmer et al.
@@
 [@Rezzonico2016], and Brady et al.
-[@Brady2012] also ran analyses with the four MLSA genes as amino acid sequences.
-*Mixta* was found to be related closest to *Tatumella* by Rezzonico et al.
-[@Rezzonico2016] and *Pantoea* by Palmer et al.
+[@Brady2012] also ran analyses with the four MLSA (multi-locus sequence analysis) genes as amino acid sequences.
+*Mixta* was found to be related closest to *Pantoea* by Palmer et al.
 [@Palmer2018].
-Brady et al.
+Both Rezzonico et al.
+[@Rezzonico2016] and Brady et al.
 [@Brady2012] suggested that *Mixta* was a part of the *Pantoea* genus and not separate.
@@
 The number of sequences closely related to one of the other taxa differed between the two *Mixta* species (Table 3).
-Additionally, although 38.7% of both *Mixta* species’ sequences had a consistent closest relative, 30 genes differed between them.
+Additionally, although these sequences had a consistent closest relative, 30 genes differed between them.
 For these 30 genes, there was a consensus for only one of the *Mixta* species and not the other.
@@
 Thus, gene sequences wherein the four analyses agreed upon each of the other taxa were concatenated to compute phylogenetic trees.
-There was a consensus for 32 genes that the closest relative was not either of the *Pantoea* species; 17 were more closely related to *E.
+There was a consensus for 32 genes that the closest relative was not either of the *Pantoea* species: 17 were more closely related to *E.
 tasmaniensis*, 8 to *E.
@@
 
-Phylogenetic analysis is an accepted method for understanding the evolutionary relatedness [@Boundless2021] while site identity is used to confirm identification and taxon boundaries [@Ciufo2018].
+Phylogenetic analysis is an accepted method for understanding the evolutionary relatedness [@Boundless2021], while site identity is used to confirm identification and taxon boundaries [@Ciufo2018].
 These two methods were conducted in this study as both nucleotide (NTS) and amino acid (AAS) sequences.
@@
 septica* share a recent common ancestor, followed by other *Pantoea* species, *Erwinia*, and then *Tatumella*.
-This agrees with *Mixta* being previously known as part *Pantoea*.
+This result agrees with *Mixta* being previously known as part *Pantoea*.
 Additionally, there were no *Mixta* sequences that were closely related to *P.
-syringae* which demonstrates that this species is a suitable outgroup.
+syringae*, which demonstrates that this species is a suitable outgroup.
 
@@
 This difference may be to a loss of phylogenetic signal when using amino acid sequences caused by codon degeneracy.
-Nevertheless, sequence type, as well as different evolutionary models between NTS and AAS, affects the results.
-AAS may be better suited for uncovering taxonomic relationships at the genus or higher level and NTS may be more appropriate for the species level.
+Nevertheless, sequence type and different evolutionary models between NTS and AAS affect the results.
+AAS may be better suited for uncovering taxonomic relationships at the genus or higher level, and NTS may be more appropriate for the species level.
 
@@
 Additionally, the other taxa followed the same pattern.
-If the genes that were closely related to non-*Pantoea* species are a result of HGT, then this suggests that HGT may be a common occurrence within the *Erwiniaceae*.
-A second explanation for the even distribution patterns in the closest relatives is that the genes are highly conserved.
-Two or more species may have been tied for closest relative and the R program may have selected one based another characteristic, such as alphabetical order.
+If the genes closely related to non-*Pantoea* species are a result of HGT, this suggests that HGT may be a common occurrence within the *Erwiniaceae*.
+A second explanation for the consistent distribution patterns in the closest relatives is that the genes are highly conserved.
+Two or more species may have been tied for the closest relative, and the R program may have selected one based on another characteristic, such as alphabetical order.
 Nevertheless, further research into these genes and the frequency of HGT events with *Mixta* and other genera within the *Erwiniaceae* is needed.
+Furthermore, for 279 (34.9%) homologues, the closest relative was consistent across all four analyses and between the two *Mixta* species.
+These results are more likely to be accurate because there was a strong phylogenetic signal and conserved site identity between the NTS and AAS.
+An additional 30 genes differed between *M.
+calida* and *M.
+gaviniae* on the closest species relative where there was a consensus for only one of the *Mixta* species.
+Codon degeneracy, codon bias, and HGT to one of the *Mixta* species may explain contentious results between the analyses and between the *Mixta* species.
 
-# Contributions to the Field Statement
+The closest genus relative was also considered to account for sequence conservation and similar identity in recently diverged species.
+There was an agreement on the closest genus relative across the analyses for approximately 63% of the *Mixta* sequences.
+Of these, only 39 homologues differed between the *Mixta* species, with only one gene, *ttuB*, having a consensus on the closest genus relative for both of the *Mixta* species.
+A gene tree for *ttuB* reveals a possible HGT event wherein *M.
+calida* received a homologue from a *Tatumella* species.
+Because *ttuB* is a single-copy gene, the original version may have been deleted.
+Additionally, this explains why the genetic distance results for *M.
+calida* and *M.
+gaviniae* differed for both the NTS and AAS.
+A second gene, *secY*, also gave contentious results between the *Mixta* species in the AAS genetic distance analysis; however, this is likely due to its highly conservative nature given the short branch lengths in its gene tree.
+
+Many previous studies into the *Erwiniaceae* have conducted phylogenetic analysis using four core genes: *atpD*, *gyrB*, *infB*, and *rpoB* [@Prakash2015; @Palmer2017; @Rezzonico2016; @Gueule2015].
+Two genes, *atpD* and *gyrB*, were removed from this study’s dataset during the filtering process; however, *infB*, *rpoB*, and twelve other related subunits were available.
+The concatenated NTS sequences produced a similar topology tree to Brady et al.’s [@Brady2011] maximum-likelihood (ML) tree based on 16S rRNA gene sequences, and none of the trees based on the four genes.
+The differences may be caused by different software programs, phylogenetic models, and the addition of genes in my study.
+Additionally, the topology of the AAS sequences of the 14 core genes was not found by any previous studies [@Prakash2015; @Palmer2017; @Palmer2018; @Rezzonico2016; @Gueule2015; @Brady2011; @Brady2012].
+When the four genes were concatenated as AAS, Palmer et al.
+[@Palmer2018] found *Pantoea* and *Tatumella* formed sister taxa followed by *Mixta*.
+My analysis identified *Pantoea* and *Erwinia* as sister taxa followed by *Mixta*.
+Nonetheless, the lack of branch support leaves the true relationships ambiguous for the AAS.
+
+Interestingly, phylogenetic trees showing a close relationship between *Mixta* and a non-*Pantoea* species were highly supported for both the NTS and AAS.
+Genes wherein there was a consensus between the four analyses were concatenated to create phylogenetic trees for each of *E.
+tasmaniensis*, *E.
+amylovora*, *T.
+saanichensis*, and *E.
+cloacae*.
+The strong branch support and long branch lengths suggest that these genes are likely candidates for HGT.
+Furthermore, the inclusion of these genes into future phylogenetic studies of the *Erwiniaceae* may result in contentious results and add a phylogenetic signal that is not due to vertical evolutionary history.
+
+In conclusion, in phylogenetic analyses of the *Erwiniaceae*, diligence must be given to the software program, evolutionary model, sequence type, and included genes.
+Different programs may offer other methods, models, and parameters.
+Additionally, model testing is a valuable tool to ensure the most accurate estimation of evolutionary history is calculated.
+The sequence type – nucleotides versus amino acids – further play a role due to the difference in sequence lengths and codon degeneracy.
+The selected genes may also affect phylogenetic results as conserved genes may not conclusively separate species, and HGT genes may skew the taxa’s relationships.
+Traditional phylogenetic methods (i.e., genetic distance and phylogenetic trees) versus site identity may also give contentious results.
+Phylogenetic analyses estimate evolutionary history, and some models penalise specific mutations.
+However, site identity analyses provide the similarity between sequences and operate on the assumption that members of a clade should be genetically similar.
+Ultimately, it is best practice for future studies to be transparent in how they conducted their phylogenetic analyses to ensure reproducibility.
+
+<!-- # Author Contributions (please edit if I accidentally missed something) -->
+<!-- KH carried out conceptualisation, methodology, investigation, formal analysis, and writing (original draft preparation).
+JS carried out conceptualisation, methodology, resources (study materials and computing resources), writing (review and editing), and supervision.
+-->
 
